### 2024-04-10 draft.tex
start q-boson paper

\documentclass[a4paper,11pt]{amsart}

\usepackage{/home/thiago/.config/dot-files/latex/globaldef}
\input{/home/thiago/.config/dot-files/latex/definitions.tex}

\bibliography{bib-database.bib}

\begin{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\title{Comments on \(Q\)-Bosons, deformed fermions and their relations
    with integrable hierarchies}

\author{Thiago Araujo}

\address{\noindent 
Instituto de Fíısica Teórica, UNESP-Universidade Estadual Paulista
R. Dr. Bento T. Ferraz 271, Bl. II, Sao Paulo 01140-070, SP, Brazil\\
and
Instituto de Física\\ Universidade de S\~ao Paulo\\ 
Rua do Matão Travessa 1371, 05508-090\\ São Paulo, SP. Brazil
}

\email{\texttt{\href{tharaujo@if.usp.br}{tharaujo@if.usp.br}}}

\keywords{integrability, spin-chains}
\subjclass[2020]{82B20, 82B23}
\date{\today}

\begin{abstract}
  In this work, we study aspects of the \(Q\)-boson model, and 
its relations with integrable hierarchies. 
% \vspace{4pt}
 \noindent \textbf{Keywords:} Integrability, Q-bosons, Schur
\end{abstract}

\maketitle

\setcounter{tocdepth}{1}
\tableofcontents

%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*
\section{q-Bosons and the Phase Model}
%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*

This section intruduces the \(q\)-boson and phase
models~\cite{Bogoliubov:1992, Bogoliubov:1997soj, Bogoliubov2005,
  Bogoliubov:1997soj, Tsilevich:2006}. We mainly follow the notation
of~\cite{Wheeler:2010vmq}.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Phase model}
Consider the \((M+1)\) set of operators \(\{\phi_i,
\phi^\dagger,\mathcal{N}_i\}_{i=0}^M\) such that
\begin{equation}
 [\mathcal{N}_i, \phi_j] = - \phi_i \delta_{i,j} \quad
 [\mathcal{N}_i, \phi_j^\dagger] =  \phi_i^\dagger \delta_{i,j}  \quad 
 [\phi_i, \phi_j^\dagger] =  \pi_i \delta_{i,j}  
\end{equation}
where \(\pi_i =(\ket{0} \bra{0})_i\) is the vacuum projection operator.
These operators can be written as
\begin{equation}
\phi = \sum_{n\geq 0}\ket{n}\bra{n+1}\quad 
\phi^\dagger = \sum_{n\geq 0}\ket{n+1}\bra{n} \quad 
N = \sum_{n\geq 0}n\ket{n}\bra{n}\; ,
\end{equation}
and it is easy to see that \(\phi^\dagger \phi = \bm{1} - \ket{0}
\ket{0}\) and \(\phi\phi^\dagger = \bm{1}\).

The Hamiltonian is given by
\begin{equation}
  H = - \frac{1}{2} \sum_{n =0}^M \left(\phi_n^\dagger \phi_{n+1}
  + \phi_n \phi_{n+1}^\dagger \right) + \bar{\mathcal{N}}\; .
\end{equation}
where \(\bar{\mathcal{N}} = \sum_{i=0}^M \mathcal{N}_i\), and we also
impose periodic boundary conditions \(\phi_{M+1} = \phi_0\) and
\(\phi_{M+1}^\dagger = \phi_0^\dagger\).

These operators appear in the context of quantum optics, and for
this reason it is referred as the \emph{phase model}. It 
corresponds to the strongly correlated \(q\)-bosons
model~\cite{Bogoliubov:1997soj} the we will define soon.

\subsubsection{Partitions representation}
Let us define the i\(^{th}\)-vacuum \(\ket{0}_i\) by the the conditions 
\(\phi_i\ket{0}_i\). The representation of this Hilbert space, denoted by
\(\mathcal{F}_i\), is given by \(\ket{n_i}_i = \phi_i^\dagger \ket{0}_i\).

Given the vacuum \(\ket{\bm{0}} = \ket{0}_0\otimes \ket{0}_1
\otimes \cdots \otimes  \ket{0}_M\),
the Fock space is defined as 
\begin{equation}
  \mathcal{F} = \bigotimes_{i=0}^M \mathcal{F}_i = 
  \left\{\ket{\vec{n}} = \ket{n_0}\otimes \ket{n_1} \otimes \cdots
  \otimes \ket{n_M} \ | \ n_i \in \mathbb{N} \right\}\; .
\end{equation}
where the states \(\ket{\vec{n}}\) are defined as 
\begin{equation}
  \ket{\vec{n}} = \ket{n_0}\otimes \ket{n_1} \otimes \cdots \otimes \ket{n_M} 
 : =  (\phi_0^\dagger)^{n_0} (\phi_1^\dagger)^{n_1} \cdots  (\phi_M^\dagger)^{n_M} \ket{\bm{0}} \; ,
\end{equation}
where we write \(\phi_i^\dagger \equiv \bm{1} \otimes  \cdots \otimes
\phi_i^\dagger \otimes \cdots \otimes \bm{1}\).
We will assume that these states are normalized, that is
\(\bracket{\vec{n}}{\vec{m}}=\delta_{\vec{n}, \vec{m}}\).

Finally, the actions of the operators \(\mathcal{N}_i\) and \(\pi_i\) are
\begin{equation}
    \pi_i\ket{\vec{n}}  = \delta_{n_i, 0} \ket{\vec{n}} \qquad 
    \mathcal{N}_i\ket{\vec{n}} = n_i \ket{\vec{n}}\; .
\end{equation}

Given a state \(\ket{\vec{n}} = \ket{n_0, n_1, \dots, n_M}\), we
associate a partition \( \lambda = (0^{n_0} 1^{n_1} 2^{n_2} \cdots
M^{n_M})\). Observe that this correspondence is not unique, since the
partition \(\lambda\) ignores the number \(n_0\) of particles in the
site \(0\). From the total number of particles, we can, on the other hand,
determine the number of particles \(n_0 = N - \ell(\lambda)\), where
\(\ell(\lambda)\) is the number of rows in the Young diagram defined
by \(\lambda\). This is very important, since in our consideations 
below, the \(N\) particle sector is fixed, thanks to the integrability of 
the model, therefore, the number \(n_0\) is known once we define the 
partition \(\lambda\).

In fact, from the correspondence above, Wheeler~\cite{Wheeler:2010vmq}
defines a map \(\mathcal{M}_\psi: \mathcal{F}\to
\mathcal{F}^{(0)}_\psi\), where \(\mathcal{F}^{(0)}_\psi\) is the
space Fock space of charged free fermions built upon the neutral
(fermionic) vacuum. We will discuss a map from the bosonic space to
the space of symmetric functions.

\subsubsection{Bethe Ansatz}
The \(L\)-matrix is given by 
\begin{equation}
  L_{an} = 
\begin{pmatrix}
x^{ - 1/2} & \phi_n^\dagger \\ \phi_n & x^{1/2}
\end{pmatrix}_a\; , 
\end{equation}
from where we can build the monodromy matrix 
\begin{equation}
  T_a(x) = L_{aM}(x) \cdots L_{a0}(x) = 
\begin{pmatrix}
A(x) & B(x) \\ C(x) & D(x)
\end{pmatrix}_a\; .
\end{equation}

With these expressions, one can finally build the Bethe states and their dual 
\begin{equation}
  \ket{\Psi(y_1, \dots, y_N)} = \prod_{j=1}^N B(y_j) \ket{\bm{0}}\qquad 
  \bra{\Psi(y_1, \dots, y_N)} = \bra{\bm{0}}\prod_{j=1}^N C(y_j) \; .
\end{equation}
When the set \(\{ y_j \ | \ j =1, \dots , N\}\) satisfies the Bethe equations 
\begin{equation}
  y^{N + M}_i = (-1)^{N-1} \prod_{\substack{j = 1 \\ j \neq i}}^N y_j\; , \qquad i = 1, \dots, N\; , 
\end{equation}
we says that the Bethe states are \emph{on-sheel}. In what follows, we will 
only consider \emph{off-shell} states.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{q-Bosons}
The Hamiltonian for the q-boson model is
given by
\begin{equation}
  \mathcal{H} = -\frac{1}{2} \sum_{n=0}^M
  \left(b_n^\dagger b_{n+1} + b_n b_{n+1}^\dagger \right) + \bar{\mathcal{N}}\; ,
\end{equation}
where \(\bar{\mathcal{N}} = \sum_{i=0}^M \mathcal{N}_i\), and as in
the case of the phase mode, we impose periodic boundary conditions
\(b_{M+1} = b_0\) and \(b_{M+1}^\dagger = b_0^\dagger\). The set of
operators \(\{b_i, b^\dagger,\mathcal{N}_i\}_{i=0}^M\) form the
\(M+1\) copies of independent q-boson algebras:
\begin{equation}
[\mathcal{N}_i, b_j^\dagger]=\delta_{i,j} b_i^\dagger\; , \quad 
[\mathcal{N}_i, b_j]=-\delta_{i,j}b_i\; , \quad
[B_i, B_j^\dagger]= \delta_{i,j} q^{-2\mathcal{N}_i}  \equiv \delta_{i,j} Q^{\mathcal{N}_i}\; , 
\end{equation}
where we write the deformation parameter as \(Q = q^{-2}\). 
In the limit \(Q\to 1\), the q-bosons are ordinary bosons. 
In the limit \(q\to \infty\), we recover the phase model we discussed above.

\subsubsection{Representation}

\subsubsection{Bethe Ansatz}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Bethe Ansatz for the q-boson}

The \(L\)-operator for the q-boson is given by
\begin{equation}
  L_n(\lambda) =
  \begin{pmatrix}
    e^\lambda & \chi B_n^\dagger \\ \chi B_n & e^{-\lambda}
  \end{pmatrix}
\end{equation}
where \(\lambda \in \mathbb{C}\) is the spectral parameter and \(\chi
= \sqrt{1 - q^{-2}}\). From this expression, we have the monodromy matrix
\begin{equation}
  T(\lambda) = \overleftarrow{\prod_{n = 1}^M} L_n(\lambda) =
  \begin{pmatrix}
    A(\lambda) & B(\lambda) \\ C(\lambda) & D(\lambda)
  \end{pmatrix}
\end{equation}
The eigenstates of the Hamiltonian are of the form
\begin{equation}
\ket{ \psi_N(\lambda_1, \dots, \lambda_N)} = \prod_{j=1}^N B(\lambda_j) \ket{0}\; , 
\end{equation}
where the parameters \(\{\lambda_i\}_{i=1}^N\) satisfies the Bethe
Algebraic Equation
\begin{equation}
  e^{2M\lambda_j} = \prod_{k\neq j}^N \frac{f(\lambda_k, \lambda_j)}{f(\lambda_j, \lambda_k)}\quad \textrm{with} \quad
  f(\lambda, \mu) = \frac{\sinh(\lambda-\mu+\gamma)}{ \sinh(\lambda- \mu)}
  \quad j = 1,\dots \ldots , N\; .
\end{equation}
where

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Bethe Ansatz for the phase model}

Taking the limit \(\gamma\to \infty\) we obtain the ABA for the phase model.
In particular, we find
\begin{equation}
  L_n(\lambda) =
  \begin{pmatrix}
    e^{ip/2} & \phi_n^\dagger \\ \phi_n & e^{-ip/2}
  \end{pmatrix}
\end{equation}
where we write \(\lambda = i p/2\). The Algebraic Bethe Equations are
\begin{equation}
\exp \left(i (M + N) p_j \right) = (-1)^{N-1} \exp \left( i \sum_{k=1}^N p_k \right)\quad j = 1, \dots, N\; ,
\end{equation}
that can be solved explicitly as
\begin{equation}
  \label{}
p_j = \frac{ 2 \pi I_j + \sum_{k=1}^N p_k}{M+N}\; , 
\end{equation}
where \(i_j \in \mathbb{z}\) ( \( + 1/2\)) if \(n\) is odd (even).

In what follows, we consider the transformation \(x^{-1} = e^{ip}\), see~\cite{Wheeler:2010vmq},
in such a way that the \(L\)-operator becomes
\begin{equation}
  L_{an}(\lambda) =
  \begin{pmatrix}
    x^{-1/2} & \phi_n^\dagger \\ \phi_n & x^{1/2} 
  \end{pmatrix}_a
\end{equation}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsubsection{Bethe states}

Now we also consider the Bethe states (and their dual) 
\begin{equation}
\ket{\Psi(\bm{x})} = \prod_{j=1}^N B(x_j) \ket{\bm{0}}\quad \; ,
\bra{\Psi(\bm{x})} = \bra{\bm{0}}\prod_{j=1}^N C(x_j) \; ,
\end{equation}
where we have written \(\bm{x} = (x_1, \dots, x_N)\), and the monodromy matrix
is given by
\begin{equation}
  T_a(x) = \prod_{j=1}^M L_{a j}(x) = 
  \begin{pmatrix}
    A(x) & B(x) \\
    C(x) & D(x)
  \end{pmatrix}_a \; .
\end{equation}
Folowing~\cite{Wheeler:2010vmq}, we define \(\mathbb{B} = x^{M/2} B(x)\) and 
\(\mathbb{C} = x^{M/2} C(x^{-1})\).

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Scalar product}

Bogoliubov~\cite{Bogoliubov2005}, in the notation~\cite{Wheeler:2010vmq},
has shown that the the scalar product of two vectors in the
\(N\)-particles sector in a chain of length \(M + 1\) is
\begin{equation}
\begin{split}
\label{eq:scalar}
  S(N, M|\bm{x}, \bm{y}) & =
  \bra{\bm{0}} \prod_{j=1}^N \mathbb{C}(x_j) \prod_{j=1}^N \mathbb{B}(y_j) \ket{\bm{0}} \\ 
 & = \frac{ \det H(\bm{x},\bm{y})}{ \prod_{i<j}(x_i - x_j)(y_i - y_j)} \; ,
\end{split}
\end{equation}
where \(H\) is an \(N\times N\) matrix with components
\begin{equation}
  H_{ij} = H(x_i, y_j) 
  =\frac{1 - (x_i y_j)^{ M + N}}{1 - x_i y_j }\; .
\end{equation}

This scalar product is related to the counting of plane
partitions, with
\begin{equation}
 S(N, M| \bm{x}, \bm{y}) 
 = \sum_{\lambda \subseteq (N, N, M)} s_{\lambda}(\bm{x}) s_{\lambda} (\bm{y})\; .
\end{equation}
In the limit \(M, N \to \infty\), one can use the Cauchy identity
\begin{equation}
  \sum_{\lambda } s_{\lambda}(\bm{x}) s_{\lambda}(\bm{y})
  = \prod_{i,j>0} \frac{1}{1- x_iy_j}\; ,
\end{equation}
and specialize this expression to \( x_i = y_i = q^{i-1/2} \) to find
the MacMahon function
\begin{equation}
  \sum_{\lambda } s_{\lambda}(\bm{q}_\star) s_{\lambda}(\bm{q}_\star) = \prod_{i,j>0} \frac{1}{1- q^{i +j -1}}\; ,
\end{equation}
where \(\bm{q}_{\star} = (q^{1/2}, q^{3/2}, \dots)\).

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{KP and Toda tau functions}

Wheeler has shown in~\cite{Wheeler:2010vmq} that the the scalar
product~(\ref{eq:scalar}) is a KP tau function with respect to the
Miwa coordinates \(t_m = \frac{1}{m}\sum_j x_j^m\) (\(t'_m =
\frac{1}{m}\sum_j y_j^m\)), where the Schur functions
\(s_{\lambda}(\bm{y}) = c_\lambda\) ( \(s_{\lambda}(\bm{x}) =
c'_\lambda\)) become Plücker coordinates, and vice-versa.
The proof for this statement is relies on the map \(\mathcal{M}_\psi\)
between the space \(\mathcal{V}\) os Bethe states and Fock
space \(\mathcal{F}_\psi\) of free fermions.

In fact, there is an easier way to see this fact. First one may see that 
\begin{equation}
 S(N,M,\bm{y}|\{\bm{t}\}) 
 = \sum_{\lambda} s_\lambda(\bm{y})  s_\lambda(\bm{t}) 
 \equiv \sum_{\lambda} c_\lambda(\bm{y})  s_\lambda(\bm{t}) 
\end{equation}
is a tau function of the KP hierarchy, since
\(c_\lambda(\bm{y}) = \det (h_{\lambda_i-i +j}(\bm{y}))\) are Plücker
coordinates~\cite{Miwa2000, Foda:2009zz}. Therefore, the truncation
\begin{equation}
 S(N,M,\bm{y}|\{\bm{t}\}) 
 = \sum_{\lambda \subseteq [N, \infty)} s_\lambda(\bm{y})  s_\lambda(\bm{t}) 
\end{equation}
is also a tau function. According to~\cite{Alexandrov:2012tr,
  Kharchev:1991gd, Zabrodin:2010ii}, the truncation of the tau
function corresponds to the inclusion of a projection operator
\(\mathrm{P}^+\) in the expectation value of the tau function.
% See section 3.3.2 of~\cite{Alexandrov:2012tr},

\subsubsection{Scalar product}
We also argue that this is also a tau function of the 2D toda hierarchy. 
We can write the Cauchy's identity as
\begin{subequations}
\begin{equation}
  \sum_{\lambda } s_{\lambda}(\bm{t}) s_{\lambda}(\bm{t}')
    = \exp \left( \sum_{m\geq1} m t_m t'_m \right) \; .
\end{equation}
where \(\bm{t}\) and \(\bm{t}'\) are two independent sets of Miwa coordinates.
It is a trivial tau function of the 2D Toda hierarchy, since one can write it as
\begin{equation}
\sum_{\lambda } s_{\lambda}(\bm{t}) s_{\lambda}(\bm{t}')
 = \bra{\bm{0}} e^{\bm{J}_+(\bm{t})} e^{-\bm{J}_-(\bm{t}')} \ket{\bm{0}}
\end{equation}
and from the free-fermions representation of the tau functions,
see~\cite{Alexandrov:2012tr}, we have that it is a tau function of the
Toda hierarchy with trivial element \(1 \in GL(\infty)\).
\end{subequations}

Therefore, the truncation 
\begin{equation}
 S(N,M|\{\bm{t}\},\{\bm{t}'\}) = 
  \sum_{\lambda \subseteq [N,N,M]} s_{\lambda}(\bm{t}) s_{\lambda}(\bm{t}')
\end{equation}
is a also a restricted tau function. These proofs follow from simple
properties of tau functions, and it should be contrated with the rich
formalism in~\cite{Wheeler:2010vmq}. Moreover, from~\cite{Zabrodin:2010ii}
we have that the this tau function can be written as the matrix integral
\begin{equation}
  S(N,M|\{\bm{t}\}, \{\bm{t}'\}) =
  \frac{1}{N!} \prod_{\ell=1}^N \oint_{\Gamma_\ell} \frac{dz_\ell}{2 \pi i z_\ell}
  e^{\xi_+(\bm{t}_+, z_\ell) - \xi_-(\bm{t}_-, z_{\ell}^{-1})} \Delta(z)\Delta(z^{-1})
\end{equation}

When the Bethe equations are satisfied, we have that \(x_j = y_j = e^{ip_j}\),
therefore \(t_{-m} = - t_m^\star\), and the the phase model is equivalent to
an ensemble of \(N\) 2D Coulomb particles on a circle.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsubsection{Correlation functions}
Bogoliubov has also shown that the correlation functions
\begin{subequations}
\begin{equation}
  P(m|\bm{x}, \bm{y}) = \bra{0} \prod_{j=1}^N C(x) \prod_{j=1}^{N-1} B(y) \phi_m^\dagger \ket{0}
\end{equation}
can be written as
\begin{equation}
  P(m|\bm{x}, \bm{y}) = \frac{(-1)^{N-1}}{(\sqrt{y_N}^{N-1}}
  \left( \prod_{t<N} \frac{x_N - y_t}{y_t} \right) \frac{\det Q}{\det H} S(N,M|\bm{x}, \bm{y})
\end{equation}
\end{subequations}
where the \(N\times N\) matrix \(Q\) has components 
\begin{equation}
 Q_{jN} = x_j^{(M + N - 1- 2m)/2} \quad  \textrm{and} \quad 
 Q_{jk} = H_{jk} \; , 
\end{equation}
where \(H_{jk} = H(x_j, y_k)\) are the components of the matrix \(H\). 
Since the components \(Q_{jN}\) are independent of the coordinates \(y\), 
we cannot write the above expression in terms of a Toda hierarchy tau
function.

We already know that 
\begin{equation}
   S(N,M|\bm{x}, \bm{y}) =  \Delta(x) \Delta(y) \det H \; ,
\end{equation}
therefore we define
\begin{equation}
  \mathbb{P}(m|\bm{x}, \bm{y}) =\frac{\det Q}{\Delta(x)}\; ,
\end{equation}
and we are going to treat the coordinates \(\{ \bm{y} \}\) as
a set of \(N\) fixed constants, in a such a way that we write \(N\)
functions \(\bm{F}(z) = (F_1, \dots, F_N)\) defined as
\begin{equation}
  F_j (z) = H(z, y_j) \quad \textrm{if} \ j \neq N\; , \qquad \textrm{and}\qquad 
   F_N (z)  = z^{(M + N - 1 - 2m)/2} 
\end{equation}
Moreover, we have that
\begin{equation}
  H(z, y_j) = \frac{1 - (y_j z)^{M + N}}{1 - y_j z} = \sum_{n=0}^{N + M - 1} (y_j z)^n\; , 
\end{equation}
therefore, we write
\begin{equation}
  F_j(z) =  \sum_{n=0}^{M + N -1} f_{j, n} z^n
\end{equation}
where we will also assume the \(M + N \in 2 \mathbb{Z}\). With these definitions,
we conclude that
\begin{equation}
  \mathbb{P}(m|\bm{x}, \bm{y}) =\frac{\det_{jk} F_j(x_k)}{\Delta(x)}\; ,
\end{equation}
and we write~\cite{Takasaki:2010qm}, 
\begin{equation}
  \det_{jk} F_j(x_k) = \det_{jk} \left(  \sum_{n=0}^{M + N -1} f_{j, n} x_k^n \right) \; .
\end{equation}
With the aid of the Cauchy-Binet formula, we write  
\begin{equation}
  \mathbb{P}(m|\bm{x}, \bm{y}) = \sum_{0\leq \ell_N\leq \dots \leq \ell_1 \leq N+M}
  \frac{\det_{jk}(f_{j, \ell_k}) \det_{jk}(x_k^{\ell_j})}{\Delta(x)}\; ,
\end{equation}
and from the definition of the Schur polynomials and the
Jacobi-Trudi expression of Plücker coordinates, we can write this last
expression as 
\begin{equation}
  \mathbb{P}(m|\bm{x}, \bm{y}) = \sum_{\lambda} c_\lambda(\bm{y}) s_\lambda(\bm{x}) \; ,
\end{equation}
where \(c_\lambda(f_{j, \ell_k})\),  \(\ell_k = \lambda_k - k +N\), define point
in the finite dimensional Grassmannian manifold \(\mathrm{Gr}(N, N+M)\).

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsubsection{Not everything is a tau-function}
It is also worth noticing that not everything is a tau function. Consider the state
\begin{equation}
  \ket{\mathcal{Y}} =
 \sum_{\vec{n}} \prod_{j=0}^M \ket{n_j}\quad \sum_j n_j = N\; .
\end{equation}
We have~\cite{Bogoliubov2005} the correlation function 
\begin{equation}
  \prod_{j=1}^N x^{M/2} \bra{\mathcal{Y}} \prod_{j=1}^N B(\bm{x}) \ket{\bm{0}} = 
  \sum_{\lambda\subseteq (N,N,M)} s_\lambda(\bm{x})\; . 
\end{equation}
This sum of Schur polynomials are not tau functions of the KP hierarchy.

We can easily see this fact, if we consider the the Fermionic
construction. Since
\begin{equation}
\tau(\bm{t}) = \bra{\bm{0}} e^{\bm{J}_+(\bm{t})} G \ket{\bm{0}}
\end{equation}
where \(G\in GL(\infty)\), and 
\begin{equation}
  \bra{\bm{0}} e^{\bm{J}_+(\bm{t})} = \sum_\lambda \bra{\mu} s_\lambda(\bm{t})\; , 
\end{equation}
therefore
\begin{equation}
\tau(\bm{t}) = \sum_{\lambda} \bra{\lambda} G \ket{\bm{0}} s_\lambda(\bm{t})
\end{equation}
and we need \(1 = \bra{\lambda} G \ket{\bm{0}} \). Then, 
\begin{equation}
  \ket{\Psi_0} = G\ket{\bm{0}} = \sum_\lambda \ket{\lambda}\; ,
\end{equation}
is a coherent state, and we conclude that
\begin{equation}
G = 1 +  \sum_{d\geq 1} \prod_{j=1}^d \psi^\star_{a_j} \psi_{-b_j}\ \notin \ GL(\infty) \;. 
\end{equation}
It is interesting to notice that this state appears as the ground
state of the crystal melting problem in~\cite{Dijkgraaf:2008ua}.

Therefore, the restriction for diagrams with \(\ell(\lambda_1) \leq N\) is
not a tau function either.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Scalar products for the Q-bosons}

Here I would like to start considering the q-boson problem. It has
been shown~\cite{Wheeler:2010vmq, Sulkowski:2008mx, Tsilevich:2006}
that the scalar product of two off-shell Bethe states in the q-boson
model is given by
\begin{equation}
S_Q(N,M | \bm{x}, \bm{y}) = \bra{\bm{0}} \prod_{j=1}^N C(x_j, Q)
\prod_{k=1}^N B(y_k, Q) \ket{\bm{0}}
= \sum_{\lambda \subseteq
  (N,N,M)} b_\lambda(Q) P_{\lambda}(\bm{x}, Q) P_{\lambda}(\bm{y}, Q)
\end{equation}
where \(P_\lambda\) denote the Hall-Littlewood polynomials.

It turns out that this expansion is  also related to the Cauchy identity for 
Hall-Littlewood polynomials~\cite{Macdonald:1998}
\begin{equation}
\sum_{\lambda} b_\lambda(Q) P_{\lambda}(\bm{x}, Q) P_{\lambda}(\bm{y}, Q)
= \prod_{j, k=1}^\infty \frac{1-Q x_j y_k}{1 - x_j y_k}\; ,
\end{equation}
and using the results in~\cite[see chap.3, sec. 4, eq. (4.7)]{Macdonald:1998}, 
we can expand this identity as 
\begin{equation}
\label{eq:cachy-hl}
 \prod_{j, k=1}^\infty \frac{1-Q x_j y_k}{1 - x_j y_k} = 
\sum_{\lambda} S_{\lambda}(\bm{x}, Q) s_{\lambda}(\bm{y}) =
\sum_{\lambda} S_{\lambda}(\bm{y}, Q) s_{\lambda}(\bm{x}) \; ,
\end{equation}
where the polynomials, the big-Schur~\cite{Wheeler:2018}, are given by 
a Jacobi-Trudi formula
\begin{equation}
  S_{\lambda} (\bm{y}, Q) = \det(q_{\lambda_i -i + j}(\bm{y}, Q))\; , 
\end{equation}
and the coefficients \(q_m\) can be obtained from the following expression
\begin{equation}
 \sum_{m} q_m(\bm{y}, Q) z^m =
 \prod_{j} \frac{1-Q y_j z}{1 - y_j z}
\end{equation}
where \(z\) is a formal variable. It has been argued
in~\cite{Foda:2008hn} that if we interpret the \(Q\)-Schur functions
as Plücker coordiantes, the identity~(\ref{eq:cachy-hl}) is a KP-tau
function.

Therefore, we see that the scalar product 
\begin{equation}
  S_Q(N,M | \bm{x}, \bm{y})
  = \sum_{\lambda \subseteq (N,N,M)} S_{\lambda}(\bm{x}, Q) s_{\lambda}(\bm{y})
  = \sum_{\lambda \subseteq (N,N,M)} S_{\lambda}(\bm{y}, Q) s_{\lambda}(\bm{x})\; ,
\end{equation}
is a restricted KP tau function with respect to both set of coordinates,
that is \(\bm{x}\) and \(\bm{y}\).

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Supersymmetric Schur polynomials}

Let us also decompose 
\begin{equation}
 \sum_{m} q_m(\bm{y}, Q) z^m =
 \prod_j (1 + y_j (-Q z)) \prod_k (1 - y_k z)^{-1} = 
 \sum_{j, k} e_j(\bm{y}) h_k(\bm{y}) (- Q)^j z^{j+k}\; .
\end{equation}
And reorganizing this sum, we conclude that 
\begin{equation}
  q_m(\bm{y}, Q)  = \sum_{j=0}^m e_j(\bm{y}) h_{m-j}(\bm{y}) (- Q)^j =
 \sum_{j=0}^m e_j(-Q\bm{y}) h_{m-j}(\bm{y}) \; ,
\end{equation}
where in the last equality we have used the homogeneity of the elementary symmetric polynomials,
and \( -Q\bm{y} = (-Qy_1, -Qy_2, \dots)\).

%One can also define the set of coordinates \(\bm{\xi} =
%\{\xi_i(\bm{x}, Q)\}\), in such a way that
%\begin{equation}
% \sum_{m} q_m(\bm{x}, Q) z^m := \sum_{m} h_m(\bm{\xi}) z^m =
% \prod_k \frac{1}{(1 - \xi_k z)} \;.
%\end{equation}
%In this sense, the \(Q\)-Schur function is an ordinary Schur function
%with respect the variables \(\bm{\xi}\).

It is easier to consider the following expansions 
\begin{subequations}
\begin{equation}
  \begin{split}
    \prod_{j=1}^N (1 - (Q y_i)z) & = \prod_{j=1}^N e^{ \ln  (1 - (Q y_i)z)} = 
    \prod_{j=1}^N \exp \left( - \sum_{n>0} \frac{Q^n y^n_j}{n} z^n \right) \\ 
    & = \exp \left( - \sum_{n>0} \sum_{j=1}^N \frac{Q^n y^n_j}{n} z^n \right)  =
    \exp \left( - \sum_{n>0} t^{(Q)}_n z^n \right)  
  \end{split}
\end{equation}
where \(t^{(Q)}_n = \frac{Q^n}{n} \sum_j y_j^n = Q^n t'_n\), and 
\begin{equation}
    \prod_{j=1}^N \frac{1}{(1 - y_iz)} = \prod_{j=1}^N e^{- \ln  (1 - y_iz)} = 
    \prod_{j=1}^N \exp \left( \sum_{n>0} \frac{y_j}{n} z^n \right)=
    \exp \left(\sum_{n>0} t_n z^n \right) \; . 
\end{equation}
\end{subequations}

Therefore 
\begin{equation}
 \sum_{m} q_m(\bm{y}, Q) z^m =
 \prod_j \frac{1 + y_j (-Q z)}{1 - y_k z}
 = \exp \left( \sum_{n\geq 1} (t'_n - t_n^{(Q)})z^n \right)
 \equiv \exp \left( \sum_{n\geq 1} T_n z^n \right)\; ,
\end{equation}
with \(T_n = (1 - Q^n) t'_n\). Then, we conclude that \(q_m(\bm{y}, Q)\)
are homogeneous polynomials with respect the Miwa coordinates
\(\bm{T} = (T_1, T_2, \dots)\). All in all, we conclude that 
\begin{equation}
  S_\lambda(\bm{t}', Q) = \det \left(h_{\lambda_i - i +j}(\bm{T})\right) = s_\lambda(\bm{T})\; .
\end{equation}

Supersymmetric Schur functions, see~\cite{Zinnjustin2009}, \(s_\lambda(\bm{\alpha}/\bm{\beta})\)
are defined as ordinary schur functions evaluated at Miwa coodinates
of the form
\begin{equation}
  T_n = \frac{1}{n} \left( \sum_{i=1}^{\dim(\alpha)} \alpha_i^n
       - \sum_{i=1}^{\dim(\beta)} (-\beta_i)^n\right)\; ,
\end{equation}
and comparing with the results above, we can see that the big Schur functions
\(S_\lambda(\bm{t}', Q)\)\; , is the supersymmetric Schur functions for \(\bm{\alpha} = \bm{y}\) 
and \(\bm{\beta} = - Q \bm{y}\), that is 
\begin{equation}
  S_\lambda(\bm{t}', Q) = s_\lambda[\bm{y}/(- Q\bm{y})]\; .
\end{equation}

Putting these facts together, we immediately conclude that since 
\begin{equation}
  \sum_\lambda s_\lambda(\bm{T}) s_\lambda (\bm{t}) \; ,
\end{equation}
is a Toda hierarchy tau function, we obviously have that 
\begin{equation}
  S_Q(N,M| \bm{x}, \bm{y}) \equiv S_Q(N,M| \bm{t}, \bm{T})
  = \sum_{\lambda \subseteq [N,N,M]} s_\lambda(\bm{T}) s_\lambda(\bm{t})\; , 
\end{equation}
is also a restricted tau function of the Toda hierarchy.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Toda hierarchy}

One of the greatest advantages of the expressions above is that we do not 
need to consider the t-Schur fermions.

It would be very interesting to have some basic understanding of what
is happenning in the fermionic representation of this tau function.
Let us expand these polynomials in a Schur polynomials basis, that is
\begin{equation}
P_\lambda(\bm{x}, Q) = \sum_{\mu} K^{-1}_{\lambda \mu}(Q) s_\lambda(\bm{x})\; , 
\end{equation}
where \(K^{-1}_{\mu\nu}(Q) \in \mathbb{Z}[Q]\) are inverse
Kostka–Foulkes coefficients, and \(K^{-1}_{\lambda\mu} = 0 \) if
\(|\lambda|\neq |\mu|\)~\cite{Macdonald:1998, Wheeler:2018}.

Expanding now the \(Q\)-Schur functions as
\begin{equation}
  S_\lambda(\bm{x}, Q) = \sum_{\mu} c_{\lambda\mu}(Q) s_\mu(\bm{x})\; ,
\end{equation}
we can fix the coefficients \(c_{\lambda\mu}\) in terms of Kostka-Foulkes
polynomials as follows. Since~\cite{Macdonald:1998} 
\begin{equation}
  \frac{1}{b_\mu}\langle P_\mu(\bm{x}, Q), P_\mu(\bm{x}, Q)\rangle = \langle
  S_\mu(\bm{x}, Q), s_\mu(\bm{x})\rangle = \delta_{\mu\nu}\; ,  
\end{equation}
we have
\begin{equation}
  \begin{split}
\delta_{\mu\nu} & = \sum_\lambda c_{\mu\lambda}  \langle s_\lambda(\bm{x}), s_\mu(\bm{x})\rangle\\
& = \sum_\lambda \sum_{\pi_1, \pi_2} c_{\mu\lambda}  K_{\lambda \pi_1} K_{\mu \pi_2}
\langle P_{\pi_1}(\bm{x}, Q), P_{\pi_2}(\bm{x}, Q)\rangle \\ 
& = \sum_\lambda \sum_{\pi} c_{\mu\lambda} b_{\mu}^{-1}  K_{\lambda \pi} K_{\mu \pi}\; ,
  \end{split}
\end{equation}
and we conclude that 
\begin{equation}
\label{eq:indices}
c_{\mu \nu} = \sum_\pi K_{\mu\pi}^{-1} b_\mu (K_{\nu\pi}^{-1})^T\; .
\end{equation}
\marginpar[left]{\tiny Fix these \\ indices later.}

Hence, we conclude that 
\begin{equation}
\sum_{\mu} S_{\mu}(\bm{x},Q) s_{\nu}(\bm{y}) = 
\sum_{\mu , \nu} c_{\mu\nu} s_{\mu}(\bm{x}) s_{\nu}(\bm{y}) 
\end{equation}
and everything boils down to the analysis of the
indices~(\ref{eq:indices}).

Do they satisfy Plücker relations? We already know that the
big Schur functions themselves are KP-tau
functions, see also~\cite{Necoechea:2019wbg}, therefore, the coefficients
\((c_\lambda)_\mu \equiv c_{\lambda\mu}\) must satisfy Plücker relations
for the KP hierarchy.

From the above considerations, we can define 2 set of functions \(F_i(x)\) and
\(G_i(y)\) such that
\begin{equation}
 F_i(x) = \sum_{n} F_{i,n} x^n \qquad 
 G_i(y) = \sum_{n} G_{i,n} y^n\; ,
\end{equation}
in such a way that 
\begin{equation}
  c_{\mu\nu} = \det(F_{i, m_j} | G_{i, n_k})\; ,
\end{equation}
and we conclude that the scalar product is, indeed, a Toda tau function. 
\marginpar[left]{\tiny Improve this\\ discussion}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{!! Fermionic construction}

Let us now consider the fermionic construction of this tau function. We write 
the limit \(N,M\to \infty\) of the scalar product
\begin{subequations}
\begin{equation}
  S_Q(\bm{T}, \bm{t}') = \sum_{\mu} s_\mu(\bm{T}) s_\mu(\bm{t}')
\end{equation}
that is quite quite trivial from the viewpoint of the free fermion construction, 
\begin{equation}
  S_Q(\bm{T}, \bm{t}') = \bra{\bm{0}} e^{\bm{J}_+(\bm{T})} e^{\bm{J}_-(\bm{t}')} \ket{\bm{0}}\; .
\end{equation}

Moreover, we also know that
\begin{equation}
  S_Q(\bm{t}, \bm{t}') = \sum_{\mu\nu} c_{\mu\nu} s_\mu(\bm{t}) s_\mu(\bm{t}')
\end{equation}
where \(n t = \sum_j x^n_j\) and \(n t' = \sum_j y^n_j\) are Miwa coordinates, then  
\begin{equation}
  S_Q(\bm{t}, \bm{t}') = \bra{\bm{0}} e^{J_+(\bm{t})} G_Q  e^{J_-(\bm{t}')} \ket{\bm{0}}
\end{equation}
and \(G_Q \in GL(\infty)\) that depends on \(Q\). We could also think of \(G_Q\)
as a twist
\(G_Q = e^{Q H} G e^{-Q H}\), with the action of \(H\in \mathfrak{gl}(\infty)\)
on ordinary fermions generating the Q-fermions. 
\marginpar[left]{\tiny Investigate this\\ twists}
\end{subequations}

It seems that we have some different representations for this tau function, 
and it is worth investigating what is happening here. That is what we want 
to do now.

%\input{./sections/fermions.tex}
%
%\input{./sections/schurpol.tex}
%
%\input{./sections/transfer.tex}
%
%\input{./sections/t-fermions.tex}
%
%\input{./sections/top-strings.tex}

%\addcontentsline{toc}{section}{References}
\printbibliography

\end{document}

### 2024-04-10 sections/fermions.tex
start q-boson paper

\section{Charged fermions and the Fermion-Boson Correspondence}

\subsection{Representation}

Here I use the notation of~\cite{Dijkgraaf:2008ua, Justin2008} with
some small changes in~\cite{Alexandrov:2012tr, Marino:2005sj}. Other useful references
are \cite{Jimbo:1983if, Babelon2003, Borot2015}. The idea we want to
explore is the relation between Young diagrams and free fermions. We
start with the definition of the generating functions
\begin{equation}
 \psi(z) = \sum_{r \in \mathbb{Z} + 1/2} \psi_{-r} z^{r - 1/2}\qquad 
 \psi^\star(z) = \sum_{r \in \mathbb{Z} + 1/2} \psi_{r}^\star z^{r - 1/2}\; . 
\end{equation}
where the operators \(\psi_r\) and \(\psi_s^\ast\), satisfy the canonical
anticommutation relations
\begin{subequations}
\begin{equation}
\label{anticom}
\{\psi_r, \psi_s\} = \{\psi_r^\star, \psi_s^\star\} = 0
\qquad \{\psi_r, \psi_s^\star\} =\delta_{rs} \quad r, s \in \mathbb{Z}+1/2\; . 
\end{equation}
We can also define the occupancy and vacancy operators at the position
\(r\) 
\begin{equation}
\eta_r = \psi_r^\star \psi_r \, \quad
\zeta_r = 1- \psi_r^\star \psi_r\; .
\end{equation}
\end{subequations}
This is a simple model of charged free fermions, where the particles
are created by the operator \(\psi^\star_r\) and annihilated by
\(\psi_r\) at the \(r^{th}\) site. Equivalently, holes at the site
\(r\) are created by \(\psi_r\) and are annihilated by
\(\psi_r^\star\).

The vacuum configuration \(\ket{\bm{0}}\) is defined by the dirac Sea
filled up to the location \(r=0\), that is
\begin{center}
    \begin{tikzpicture}
    \node at (-4,0) { \(|\bm{0}\rangle=\)};
    \node at (-3,0) {\(\cdots\)};
    \draw (-2.5,0) -- (-2,0);
    \node at (-2,0) {\(\bullet\)};
    \draw (-2,0) -- (-1,0);
    \node at (-2,-.5) {\(-\tfrac{3}{2}\)};
    \node at (-1,0) {\(\bullet\)};
    \draw (-1,0) -- (-.06,0);
    \node at (-1,-.5) {\(-\tfrac{1}{2}\)};
    \node at (0,0) {\(\circ\)};
    \draw (.06,0) -- (.95,0);
    \node at (0,-.5) {\(\tfrac{1}{2}\)};
    \node at (1,0) {\(\circ\)};
    \draw (1.05,0) -- (1.5,0);
    \node at (1,-.5) {\(\tfrac{3}{2}\)};
    \node at (2,0) {\(\cdots\)};
  \end{tikzpicture}
\end{center}	
We assume that this state exists. Alternatively, one might define the empty state
\(\ket{\bm{\infty}}\) as \(\psi_r\ket{\bm{\infty}} = 0 \), \(\forall \ r \), therefore
\(\ket{\bm{0}} = \prod_{r < 0} \psi^\star_r \ket{\bm{\infty}}\). Regardless the case,
these definitions give equivalent results.

We then have
\begin{equation}
\psi_{r}^\star\ket{\bm{0}}=0\ \ \ r< 0\; , \qquad  
\psi_{r} \ket{\bm{0}} = 0 \ \ \ r> 0
\end{equation}
The dual vacuum is defined as
\begin{equation}
\bra{\bm{0}} \psi_{r}=0\ \ \ r< 0\; , \qquad  
\bra{\bm{0}} \psi_{r}^\star = 0 \ \ \ r> 0\; ,
\end{equation}
therefore, we have the relation \((\psi_r)^\dagger\equiv \psi_r^\star\).

We can also build the shifted vacuum \(\ket{\ell}\) with \(\ell \in
\mathbb{Z}\) as
\begin{equation}
  \label{eq:charged-states}
  \ket{\ell} = 
  \left\{
  \begin{array}{ll}
    \psi^\star_{\ell - 1/2} \psi^\star_{\ell - 3/2} \cdots \psi^\star_{3/2} \psi^\star_{1/2} \ket{\bm{0}} & \ell > 0 \\
    \psi_{\ell + 1/2} \psi_{\ell + 3/2} \cdots \psi^\star_{-3/2} \psi^\star_{-1/2} \ket{\bm{0}} & \ell < 0
  \end{array}
  \right.
\end{equation}
These states satisfy
\begin{equation}
\psi_r\ket{\ell} = 0 \quad r > \ell \quad , \quad  
\psi_r^\star\ket{\ell} = 0  \quad r < \ell \; .
\end{equation}
Let us denote the Fock spaces labeled by these vacua as
\(\mathcal{F}_\ell\), for a complete discussion on the
structure of these spaces, see~\cite{Wheeler:2010vmq}. 
In particular, we have
\begin{equation}
\ket{\ell + 1} = \psi^\ast_{\ell + 1/2} \ket{\ell} \qquad
\ket{\ell} = \psi_{\ell + 1/2} \ket{\ell + 1} \; .
\end{equation}

Many of the results and identities we use have been explained in the
review~~\cite{Alexandrov:2012tr}. In order to pass from their notation
to ours, we can simply consider the transformation
\begin{equation}
\psi_k \equiv (\psi_{AZ})^\star_{k - 1/2} \qquad
\psi_k^\star \equiv (\psi_{AZ})_{k - 1/2} \; .
\end{equation}
where \textrm{AZ} stands for \emph{Alexandrov-Zabrodin.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{States and Partitions}

A partition is a sequence of weakly decreasing non-negative integers,
\(\mu_i \geq \mu_{i+1}\) \(i = 1, \dots, n\). We can think of
partition as Young diagrams, where the \(i^{th}\) line has \(\mu_i\)
boxes. Given the partition \(\mu = (\mu_1, \mu_2, \dots, \mu_r)\),
the Frobenius coordinates \(\mu = (\alpha_j , \beta_j)\) are defined as
\begin{subequations}
\begin{equation}
\alpha_j=\mu_j - j \; , \quad 
\beta_j=\mu^T_j - j \; .
\end{equation}
It is convenient to write 
\begin{equation}
a_j=\alpha_j +\frac{1}{2}	\; , \quad 
b_j=\beta_j+\frac{1}{2}\; 
\end{equation}
and hereafter, these are the Frobenius coordinates we work with. For
further details, see~\cite{Alexandrov:2012tr}.
\end{subequations}
Moreover, we denote the number of
boxes in a partition \(\mu\) as
\begin{equation}
\#\texttt{Boxes}\equiv |\mu| = \sum_{j=1}^d (a_j+b_j)
\end{equation}

Given a partition \(\mu = (\mu_1, \mu_2, \cdots, \mu_n)\), one
can associate a fermionic state \cite{Dijkgraaf:2008ua, Okounkov:2003sp}
\begin{equation}
\label{eq:part-states}
\ket{\mu} = (-1)^{b(\lambda)} \prod_{j=1}^d \psi_{a_j}^\star \psi_{-b_j}\ket{\bm{0}}\; , \qquad
\langle \mu | = (-1)^{b(\lambda)}  \bra{\bm{0}}\prod_{j=1}^d  \psi_{-b_j}^\star \psi_{a_j}\;
,\quad a_j, b_j>0\ \ \forall j\; ,
\end{equation}
where we have added the factor \(b(\lambda) = \sum_{j = 1}^d (b_j -
1/2)\) for convenience. In~\cite{Alexandrov:2012tr}, the authors
define the same factor in the Schur polynomials expansion. We have
moved this factor to the states, as in~\cite{Marino:2005sj}, so that
the Schur polynomials expansio does not have any negative sign moving
around. It is worth noticing that if we have different partitions,
\(\ket{\lambda} \neq \ket{\nu}\), then there is a mismatch in the
number of pairs \(\psi^\star \psi\), and the states are orthonormal,
\(\langle \mu |\nu \rangle=\delta_{\mu\nu}\).

It is just a combinatorial exercice to show that the partition
states~\ref{eq:part-states}, and more generally the charged states,
can be equivalently written as
\begin{equation}
  \label{eq:part-states:2}
  \ket{\mu, \ell} = \psi^\star_{\ell + \mu_1 - 1/2} \dots
  \psi^\star_{\ell + \mu_n - n + 1/2} \ket{\ell - n}\; .
\end{equation}
See~\cite{Alexandrov:2012tr} for a discussion on both representations.

The normal ordering (with respect to the vacuum \(\ket{\bm{0}}\) is
defined as
\begin{equation}
:\psi_s \psi_r^\star: = \psi_s \psi_r^\star - \bra{\bm{0}} \psi_s
  \psi_r^\star \ket{\bm{0}} \; .
\end{equation}
That means that we need to move all creation operators to the left, and all
annihilation operators to the left, taking into account the negative factors
comming from the anticommutation relations. That is
\begin{equation}
:\psi_s \psi_r^\star:   = - :\psi_r^\star \psi_s: = 
\left\{
\begin{array}{rll}
-\psi_r^\star \psi_s && r>0 \\
\psi_s \psi_r^\star && r<0 
\end{array}	
\right.
\end{equation}

Using these relations, one can see that the number of boxes in
a generic partition \(\ket{\mu}\) is the eigenvalue of the following
operator \cite{Marino:2005sj}
\begin{equation}
L_0 = \sum_{r \in \mathbb{Z} + 1/2} r :\psi_{-r} \psi_r^\star:
\end{equation}
Therefore, we have
\begin{equation}
q^{L_0}|\mu \rangle = q^{|\mu|}|\mu \rangle \; .
\end{equation}

% CHECK THESE THINGS LATER
% One interpretation to this system is given in terms of the free fermion
% \begin{subequations}
% \begin{equation}
% H= \sum_r r  \psi_r^\star \psi_r \; .
% \end{equation}
% It is easy to see that the state \(\ket{\bm{0}}\) is, indeed,
% a vacuum of the free Hamiltonian
% \begin{equation}
% H=\sum_r r :\psi^\star_r \psi_r: \; .
% \end{equation} 
% Explicitly, we have
% \begin{equation}
% \begin{split} 
%   H \ket{\bm{0}} & = \left(  \sum_{r<0} r  :\psi^\star_r \psi_r:  + \sum_{r>0} r 
%    :\psi^\star_r \psi_r:  \right) \ket{\bm{0}}  \\
%   & = \left(  \sum_{r<0} r  :\psi^\star_r \psi_r:  + \sum_{r>0} r  :\psi^\star_r 
%    \psi_r:  \right) \ket{\bm{0}}  \\
%   & = \left( - \sum_{r<0} r \psi_r \psi^\star_r  +
%        \sum_{r>0} r \psi^\star_r \psi_r \right) \ket{\bm{0}} = 0	
% \end{split}
% \end{equation}
% The Hamiltonian above is equivalent to the continuous system
% \begin{equation}
% S\sim \int d z\ \psi^\star(z) \partial \psi(z)\ \ \Rightarrow \ \  \sum_r r  \psi_r^\star \psi_r  
% \end{equation}
% \end{subequations}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{\(\mathfrak{gl}(\infty)\) and \(\widehat{\mathfrak{gl}(1)}\) Algebras}

The free fermion formulation allows us to rewrite the partition
problem above in terms of algebraic problems. Bilinears
\(\psi_r\psi_s^\star\) are closed under commutation
\begin{equation}
[\psi_r\psi_s^\star, \psi_p\psi_q^\star]= \delta_{sp} \psi_r\psi_q^\star -  \delta_{sq} \psi_p\psi_s^\star\; .
\end{equation}
We can define the objects
\begin{equation}
X = \sum_{r,s} M_{rs}   \psi_r \psi^\star_{s}   \; ,
\end{equation}
and it is easy to see that the normal ordering introduces a central
extension in the commutation relation above.  All in all, we define
the algebra \(\widehat{\mathfrak{gl}}(\infty)\) \cite{Babelon2003}
\begin{equation}
\mathfrak{gl}(\infty): = \{X=M_{rs} \psi_r \psi^\star_{s} \ |
\ M_{rs}=0\ \forall\ |r+s|\geq N>> 0 \}\oplus \mathbb{C}\; .
\end{equation}

The Heisenberg algebra \(\hat{\mathfrak{u}}(1) \equiv
\widehat{\mathfrak{gl}}(1) \subset \widehat{\mathfrak{gl}}(\infty) \)
is defined as
\begin{equation}
[J_m , J_n]  =m \delta_{m+n,0}\; ,
\end{equation} 
where the generators are given by
\begin{equation}
J_m := \sum_r  :\psi_{r-m}^\star \psi_r :\; , 
\qquad 
J_m^\dagger := \sum_r :\psi_r^\star \psi_{r-m}:\;. 
\end{equation} 
It is easy to verify that \(J^\dagger_m = J_{-m}\). Moreover, one may easily check
that the normal ordering is unecessary for the cases where \(m\neq 0\), since
\begin{equation}
\begin{split}
J_m := &\sum_r :\psi_{r-m}^\star \psi_r: \\
= & - \sum_{r<m} \psi_r \psi_{r-m}^\star + \sum_{r>m} \psi_{r-m}^\star \psi_r  \qquad m\geq 0\; .
\end{split}
\end{equation} 	
Finally, it is easy to check that
\begin{equation}
J_m \ket{\bm{0}}= \bra{\bm{0}} J_{-m}=0 \qquad \forall \ m\geq 0\; .
\end{equation}

Since the positive and negative modes commute among themselves, we can
define the following operators
\begin{equation}
\begin{split}
\bm{J}_+({\bf t}) & :=\sum_{m>0}t_m  J_m\; ,\quad  \bm{t}=(t_1, t_2, \dots )\\
\bm{J}_-({\bf t}) & :=\sum_{m<0}t_m  J_m\; ,\quad  \bm{t}=(t_{-1}, t_{-2}, \dots )\; ,
\end{split}
\end{equation}
where one mat think of these parameters as an infinite set of
times. Finally, these two objects satisfy the identity
\begin{equation}
\label{v-ope}
\boxed{
  e^{\bm{J}_{+}(\bm{t})}e^{\bm{J}_{-}(\bm{t}')}
  = \exp\left( \sum_{n>0} n t_n t'_{-n} \right) e^{\bm{J}_{-}(\bm{t}')} e^{\bm{J}_{+}(\bm{t})}}
\end{equation}

Observe that \(\bm{J}_+(\bm{t}) \ket{\bm{0}}=0 \) and \(
\bra{\bm{0}} \bm{J}_-(\bm{t})=0 \) so that
\begin{equation}
e^{\bm{J}_+(\bm{t})} \ket{\bm{0}}= \ket{\bm{0}} \; , \qquad
 \bra{\bm{0}} e^{\bm{J}_-(\bm{t})} = \bra{\bm{0}}\; .
\end{equation}
Furthermore, the states \(e^{\bm{J}_-(\bm{t})}\ket{\bm{0}}\) can be
expanded in terms of basis states with the Schur polynomials as their
coefficients~\cite{Alexandrov:2012tr}. In fact, here we assume that
the definitions of the Schur polynomials are given in terms of the
free fermions we have been working so far.

Let us define the bosonic states
\begin{equation}
|\vec{k}\rangle \equiv |k_1, k_2, \dots \rangle = \prod_{j=1}^{\infty}
(J_{-j})^{k_j}|0\rangle\; .
\end{equation}

The two bases \(|\vec{k}\rangle\) and
\(|\lambda\rangle\) are related by~\cite{Marino:2005sj}
\begin{subequations}
\begin{equation}
|\lambda\rangle = \sum_{\vec{k}}
\frac{\chi_\lambda[C(\vec{k})]}{z_{\vec{k}}} |\vec{k}\rangle\; ,
\end{equation}
and inverse 
\begin{equation}
|\vec{k}\rangle = \sum_{\lambda} \chi_\lambda[C(\vec{k})]
|\lambda\rangle \; ,
\end{equation}
\end{subequations}
where the coefficients \(z_{\vec{k}} = \prod_{j=1}^\infty k_j!
j^{k_j}\), and \(\chi_\lambda[C(\vec{k})]\) is the character of the
symmetric group \(\mathfrak{S}_n\) evaluated in the
\(\lambda\)-representation and conjugacy class
\(C(\vec{k})\). Moreover, the sum is performed over the partitions
\(\lambda\) with number of boxes the level of the state
\(|\vec{k}\rangle\), that is
\begin{equation}
|\lambda|= \sum_{j\geq 1} j k_j\; . 
\end{equation}
In fact, one can also write the bosonic states themselves in a
partition-like notation. First, we observe that \(J_m J_n = J_n J_m\),
for \(m\) and \(n\) negative. Therefore, we organize the state
in decreasing order, that is \(J_m^{k_m} J_n^{k_n}\), \(m<n\)
\begin{equation}
  |\vec{k}\rangle = \prod_{j>0}(J_{-n_j})^{k_j}\ket{\bm{0}} =
J_{-1}^{k_1} J_{-2}^{k_2} \cdots J_{-n}^{k_n}\cdots \ket{\bm{0}} 
\end{equation}
Now, each operator \(J_{-n}\) corresponds to a new row, and its index
\(n\) gives the length of the corresponding row. One useful feature is
that the bosonic and fermionic spaces are decomposed into subspaces
with fixed number os boxes.
Given the state \(\vec{k}\), one can write
the corresponding Young diagram with the Python's codes described in
the appendix~\ref{app:snippets}.

\subsubsection{Frobenius character formula}
The explicit formulas for the characters can be obtained from the
following expression~\cite{Fulton:2004uyc} give the charactes
\begin{subequations}
\begin{equation}
\label{eq:frobenius}
\chi_{\lambda}[C(\vec{k})] = \left[ 
\Delta(\vec{x}) \prod_{j=1}^{r} P_j(\vec{x})^{k_j} 
\right]_{(\ell_1, \dots, \ell_m)}\; ,
\end{equation}
with
\begin{equation}
\Delta(\vec{x}) = \prod_{i<j}(x_i-x_j)\; , \quad P_j(\vec{x})=\sum_{i=1}^mx_i^j\; .
\end{equation}
\end{subequations}
The vector \(\vec{k}\) have a finite number of nonzero
components, \(\vec{k} = (k_1, \dots, k_r)\).Additionally, we have
the auxiliary coordinates \(\vec{x} = (x_1, \dots, x_m)\) and \(m\) is the
number rows of the Young diagram \(\lambda = (\lambda_1, \dots ,
\lambda_m)\). Furthermore, we also need the following coefficients
\begin{equation}
\ell_j = \lambda_j + m -j\; ,\ \ j=1,\dots, m\; .
\end{equation}
Given a generic polynomial \(f(\vec{x})\), we have used the following notation
 \begin{equation}
\left. f(\vec{x})\right|_{(\ell_1, \dots, \ell_m)}:=\text{coeff. of}
\ \ x_1^{\ell_1}\cdots x_m^{\ell_m}\; .
\end{equation}

The advantage of this construction is that now we have
all ingredients to translate bosonic to fermionic partition
states. For example, we have the following characters
\[
\begin{array}{cr} 
    |\lambda|  = 1 &
    (1,0,\dots)\\
    \hline \hline
    \chi_{(1)} & 1
\end{array}
\]
\[
\begin{array}{crr} 
    |\lambda| = 2 & (0,1,0,\dots) & (2,0,\dots) \\
    \hline\hline
    \chi_{(1,1)} & -1 & 1 \\ 
    \chi_{(2)} & 1 & 1 \\ 
\end{array}
\]
\[
\begin{array}{crrr} 
   |\lambda| = 3 & (0,0,1,\dots) &  (1,1,0,\dots) & (3,0,0,\dots) \\
  \hline \hline
    \chi_{(1,1,1)} & 1 & -1 & 1 \\ 
    \chi_{(2,1)} & -1 & 0 & 2 \\ 
    \chi_{(3)} & 1 & 1 & 1 \\ 
\end{array}
\]
For the calculation of any character, one can use the code in 
the appendix~\ref{app:snippets}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Schur Functions}

We define the Schur functions to be the coefficients in the expansions
\begin{equation}
\label{evol}
\begin{split}
    e^{\bm{J}_-(\bm{t})} \ket{\bm{0}}
    & = \sum_\lambda s_{\lambda}(\bm{t}) \ket{\lambda} \\
    \bra{\bm{0}} e^{\bm{J}_+(\bm{t})} 
    &= \sum_\lambda s_{\lambda}(\bm{t})\bra{\lambda}  
\end{split}
\end{equation}
Using the orthogonality of the vector states \(|\mu\rangle\), we have
\begin{equation}
\boxed{
s_{\lambda}(\bm{t}) = 
\bra{\lambda} e^{\bm{J}_-(\bm{t})} \ket{\bm{0}}}
\end{equation}

More generally, one can consider two Young diagrams
\(\mu \subseteq \lambda\), then
\begin{equation}
  \label{eq:skschur}
\begin{split}
e^{\bm{J}_-(\bm{t})} \ket{\mu}
& = \sum_\lambda s_{\lambda/\mu}(\bm{t})\ket{\lambda} \\
\bra{\mu} e^{\bm{J}_+(\mathbf{t})} 
&= \sum_\lambda s_{\lambda/\mu}(\bm{t}) \bra{\lambda}\; .
\end{split}
\end{equation}
Consequently
\begin{equation}
\boxed{
s_{\lambda/\mu}(\bm{t}) = 
 \bra{\lambda}e^{\bm{J}_-(\bm{t})}|\mu\rangle}\;. 
\end{equation}
These functions are called \emph{Skew Schur functions}. 
In the particular case where \(\mu =\bm{0}\), we recover the
expressions for the ordinary Schur functions.

Let us consider the first coefficients in the expansion
\begin{equation}
e^{\bm{J}_-(\bm{t})} \ket{\bm{0}}=\left(
1 + t_{-1}J_{-1} + t_{-2}J_{-2} + \cdots 
\frac{1}{2}t_{-1}t_{-1}J_{-1} J_{-1} + \cdots\; .
\right)
\ket{\bm{0}}
\end{equation}

Diagrammatically, we have
\begin{center}
    \begin{tikzpicture}
    \node at (-2.3,.5) {\(e^{\bm{J}_-(\bm{t})} \ket{\bm{0}} =\)};
    \draw[line width=0.5mm][rotate=45]   (0,0) -- (0,1.7);
    \draw[line width=0.5mm][rotate=45]   (0,0) -- (1.7,0);
    \end{tikzpicture}
    \begin{tikzpicture}
    \node at (-1.4,.5) {\(+ s_{(1)}\)};
    \draw [line width=0.3mm][rotate=45][fill=lightgray]  (0,0) rectangle (.5,.5);
    \draw[line width=0.5mm][rotate=45]   (0,0) -- (0,1.7);
    \draw[line width=0.5mm][rotate=45]   (0,0) -- (1.7,0);
    \end{tikzpicture}		
    \begin{tikzpicture}
    \node at (-1.4,.5) {\(+ s_{(2)}\)};
    \draw [line width=0.3mm][rotate=45][fill=lightgray] (0,0) rectangle (1,.5);
    \draw[line width=0.3mm][rotate=45]   (0.5,0) -- (0.5,.5);
    \draw[line width=0.5mm][rotate=45]   (0,0) -- (0,1.7);
    \draw[line width=0.5mm][rotate=45]   (0,0) -- (1.7,0);
    \end{tikzpicture}		
    \begin{tikzpicture}
    \node at (-1.5,.5) {\(- s_{(1,1)}\)};
    \draw [line width=0.3mm][rotate=45][fill=lightgray] (0,0) rectangle (.5,1);
    \draw[line width=0.3mm][rotate=45]   (0,.5) -- (.5,.5);
    \draw[line width=0.5mm][rotate=45]   (0,0) -- (0,1.7);
    \draw[line width=0.5mm][rotate=45]   (0,0) -- (1.7,0);
    \end{tikzpicture}
    \begin{tikzpicture}
    \node at (-1.4,.5) {\(+ s_{(3)}\)};
    \draw [line width=0.3mm][rotate=45][fill=lightgray] (0,0) rectangle (1.5,0.5);
    \draw[line width=0.3mm][rotate=45]   (.5,0) -- (.5,.5);
    \draw[line width=0.3mm][rotate=45]   (1,0) -- (1,.5);
    \draw[line width=0.5mm][rotate=45]   (0,0) -- (0,1.7);
    \draw[line width=0.5mm][rotate=45]   (0,0) -- (1.7,0);
    \end{tikzpicture}
    \begin{tikzpicture}
    \node at (-1.6,.5) {\(+ s_{(1,1,1)}\)};
    \draw [line width=0.3mm][rotate=45][fill=lightgray] (0,0) rectangle (.5,1.5);
    \draw[line width=0.3mm][rotate=45]   (0,.5) -- (.5,.5);
    \draw[line width=0.3mm][rotate=45]   (0,1) -- (.5,1);
    \draw[line width=0.5mm][rotate=45]   (0,0) -- (0,1.7);
    \draw[line width=0.5mm][rotate=45]   (0,0) -- (1.7,0);
    \end{tikzpicture}
    \begin{tikzpicture}
    \node at (-1.5,.5) {\(+ s_{(2,1)}\)};
    \draw [line width=0.3mm][rotate=45][fill=lightgray] (0,0) rectangle (.5,1);
    \draw [line width=0.3mm][rotate=45][fill=lightgray] (0,0) rectangle (1,.5);	
    \draw[line width=0.3mm][rotate=45]   (.5,0) -- (.5,.5);		
    \draw[line width=0.5mm][rotate=45]   (0,0) -- (0,1.7);
    \draw[line width=0.5mm][rotate=45]   (0,0) -- (1.7,0);
    \node at (1.4,.5) {\(+ \cdots\)};
    \end{tikzpicture}							
\end{center}
where the coefficients are
\begin{equation}
s_{(1)}=t_{-1}\; , \quad s_{(2)}=\frac{1}{2}t_{-1}^2 + t_{-2} \; ,\quad s_{(1,1)}
=\frac{1}{2}t_{-1}^2- t_{-2}\; ,\quad  \dots
\end{equation}

The Miwa coordinates are given by \(t_j = \frac{1}{j}\sum_{a = 1}^j
x_a^m\) where \(\bm{x} = (x_1, x_2, \dots, x_m)\) is a set of \(m\)
variables. If we write these Schur functions in terms of \(\bm{x}\),
these functions become symmetric polynomials, and we call them
\emph{Schur polynomials}. In terms of these coordinates, the Schur
polynomials are characters of the \(\lambda\) representations
evaluated at the diagonal matrix \(\mathrm{diag}(x_1, x_2, \dots,
x_m)\). In any case, in the text we use both terminologies as
synonyms.

%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Fermion-Boson correspondence}

The fermion boson correspondence is a relation between basis of
Hilbert space~\cite{Cordes:1994fc, Marino:2005sj}. Let start this
definition with some basic facts. The complete homogeneous symmetric
polynomials are defined by
\begin{equation}
  \label{eq:com-hom-sym-pol}
\exp\left(\sum_{k\geq 1} T_k z^k\right): = \sum_{k\geq 0} h_k(\bm{T}) z^k \; ,
\end{equation}
with \(h_0=1\) and we also define \(h_{-k}=0\  \forall\ k>0\). Explicitly we have 
\begin{equation}
  h_k(\bm{T}) = \sum_{k_1+2k_2 +\dots = k}\frac{T_1^{k_1}}{k_1!} \frac{T_2^{k_2}}{k_2!}\cdots
  \; .
\end{equation}

We can use this polynomials to express the \emph{coherent state}
\begin{equation}
\begin{split}
  \ket{\mathcal{T}}& = \exp\left( \sum_{n\geq 1} t_{n} J_{-n}\right) \ket{\bm{0}} = 
  \exp\left.\left(\sum_{n\geq 1} T_n z^n\right)\right|_{z=1}  \ket{\bm{0}} \\
& = \sum_{k\geq 0} \sum_{k_1+2k_2+\cdots = k}\frac{t_{1}^{k_1}}{k_1!}
\frac{t_{2}^{k_2}}{k_2!}\cdots \ket{k_1, k_2, k_3\dots}
\end{split}
\end{equation}
where we have used that \(T_k \equiv t_k J_{-k}\), \(\bm{t}_- = (t_1,
t_2 \dots)\) , \(\vec{k} = (k_1, \dots, k_r)\) and \(\sum_j j k_j =
|\lambda|\).  One should observe that in contrast with our previous
definition, the parameters \(\bm{t}_-\) have been changed to
\(\bm{t}_+\equiv \bm{t}\). Since there is no intrinsic meaning in those objects,
this transformation is just a relabeling.

Using the expansion of \(\ket{\vec{k}}\) in terms of characters
coefficients, we can write the explicit expression for the Schur as
\begin{equation}
\begin{split}
  \ket{\mathcal{T}}& = \sum_{\lambda} s_\lambda(\bm{t})\ket{\lambda} \\
    & = \sum_\lambda\left( \sum_{k\geq 0} \sum_{k_1+2k_2+\cdots = k}\frac{t_{1}^{k_1}}{k_1!}
    \frac{t_{2}^{k_2}}{k_2!}\cdots \chi_\lambda [C(\vec{k})] \right) \ket{\lambda}
\end{split}
\end{equation}
This formula is very hard to calculate explicily, but it is useful to know that
if we change the coefficients \(\chi_\lambda\), we have a new type of fermion boson
correspondence.

If we now identify \(t_{j} \equiv \frac{1}{j}P_j(x)\), we can
write all expressions in terms of the coordinates \(\vec{x}\). And we
conclude that
\begin{equation}
\langle \mathcal{T}| \vec{k} \rangle= \langle \vec{k} |\mathcal{T}\rangle = P_{\vec{k}}(x)\; .
\end{equation}

Moreover, we have seen that Schur polynomials are defined as
\(s_{\lambda}(t)=\langle \mathcal{T}| \lambda \rangle\), and since
Schur and Newton polynomials are related by the Frobenius formula, it
is easy to see that there is a relation between the fermionic and
bosonic states. This relation is the Fermion-Boson correspondence.

### 2024-04-10 sections/top-strings.tex
start q-boson paper

\section{Topological strings}

Moreover, we also have 
\begin{equation}
 S(N, q, \{\bm{t}\}, \{\bar{\bm{t}}\}) 
 = \sum_{\lambda \subseteq [N, \infty)} s_\lambda(\bm{t})  s_\lambda(\bar{\bm{t}}) 
\end{equation}
that is a trivial case of a hypergeometric tau function~\cite{orlov:2001}.

The partition function of D-branes on \(\mathbb{C}^3\)
is similar, see\cite{Saulina:2004da}. 
\begin{equation}
  Z_D(q, a) = Z(q) \prod_{n=1}^\infty \frac{1}{1- e^a q^{n - 1/2}}\; .
\end{equation}
We  can also write this extra factor in terms of Schur polynomials as
with \(x= (e^a, 0, 0, \dots)\) and \(y = (q^{1/2}, q^{3/2}, \dots)\),
therefore
\begin{equation}
\prod_{n=1}^\infty \frac{1}{1- e^a q^{n - 1/2}} = \sum_\lambda s_{\lambda}(e^a, 0, 0, \dots)
s_{\lambda}(q^{1/2}, q^{3/2}, \dots)\; .
\end{equation}
And using the homogeneity of the Schur polynomials, we have
\begin{equation}
s_{\lambda}(e^a, 0, 0, \dots) = e^{|\lambda|a} s_{\lambda}(1, 0, 0, \dots)\; . 
\end{equation}
One can also observe that this equation is the homogeneous polynomial at the point
\(z = e^a\) and \(x=(q^{1/2}, q^{3/2}, \dots)\).

In fact, the partition function for \(M\) D-branes in this geometry is given by 
\begin{equation}
Z_{MD} = M(q) \prod_{i<j}(1 - q^{N_i - N_j}) \prod_{i=1}^M \prod_{n_i=1}^{N_i} (1 - q^{n_i})^{-1}\; .
\end{equation}
For just one brane, we recover something similar to Bogoliubov. So, it seems
that \(M\) counts the types of bosons, while \(N_0\) is the soliton sector.

### 2024-04-10 sections/transfer.tex
start q-boson paper

\section{Transfer matrices}

This section is a review of the transfer matrix
formalism~\cite{Alexandrov:2012tr, Okounkov:2003sp, Justin2008}.  Let
us now define the following operators
\begin{subequations}
\begin{equation}
\Gamma_\pm(z):=\exp \phi_\pm(z)\; ,
\end{equation}		
where 
\begin{equation}
\phi_\pm(z):= \sum_{\pm n\geq 1}\frac{z^{n}}{n} J_{n}\; .
\end{equation}
These are the creation and annihilation components of the bosonic
vertex operator \(\mathcal(z) = \exp \phi(z)\).  By analogy with
statistical mechanics systems, we may consider that the time evolves
in a series of small steps, and for this reason we say that
\(\Gamma_\pm(x_i)\) is a transfer operator.
\end{subequations}

Moreover, observe that 
\begin{subequations}
\begin{equation}
\begin{split}
\prod_{a=1}^m \Gamma_\pm(x_a) & = \exp \sum_{a = 1}^m \phi_\pm(x_a) \\
	& = \exp \sum_{a = 1}^m  \sum_{\pm n\geq 1}\frac{x_a^{n}}{n} J_{n}\\
	& = \exp  \sum_{n\geq 1} \left(\sum_{a = 1}^m  \frac{x_a^{n}}{n} \right) J_{n}
\end{split}
\end{equation}
and using the Miwa coordinates (\ref{sympol}), we finally obtain
\begin{equation}
\label{oper}
\boxed{\prod_{i\geq 1} \Gamma_\pm(x_a) = e^{J_\pm (\mathbf{t})}}\; .
\end{equation}	
\end{subequations}

Using the expansion of skew Schur polynomials~(\ref{eq:skschur}), we have 
\begin{subequations}
\begin{equation}
   e^{\bm{J}_-(\bm{t}}\ket{\mu} \equiv 
   \prod_{a=1}^m \Gamma_-(x_a) \ket{\mu}
   = \sum_\lambda s_{\lambda/\mu}(\mathbf{t})\ket{\lambda}\; .
\end{equation}

One now can use the special case 1 we discussed above where \(\vec{x}
= (z,0,0, \dots) \), that as we alrealdy seen, corresponds to the
choice \(t_n = \tfrac{x^n}{n}\), \(h_k = z^k\) and \(e_l = 0\) for \(l > 1\).
Moreover, using~\ref{eq:schur:case1}, we can see that
\begin{equation}
  \begin{split}
    e^{\bm{J}_-(\{z^k/k\})} \ket{\mu} & \equiv \Gamma_-(z) \ket{\mu}
    = \sum_\lambda s_{\lambda/\mu}(\{z^k/k\})\ket{\lambda}\\
    & = \sum_{\lambda \succ \mu}z^{|\lambda| - |\mu|}\ket{\lambda}\; .
  \end{split}
\end{equation}
Therefore, the particular case where \(z=1\) gives the beautiful expression
\begin{equation}
    \Gamma_-(1) \ket{\mu} = \sum_{\lambda \succ \mu}\ket{\lambda}\; .
\end{equation}
It is also easy to show that
\begin{equation}
    \Gamma_+(1) \ket{\mu} = \sum_{\lambda \prec \mu}\ket{\lambda}\; .
\end{equation}
\end{subequations}
For example, the case \(\mu = \bm{\emptyset}\) gives an expansion of
one row diagrams. An in fact, \(\Gamma_+(1)^n \ket{\bm{0}}\) gives an
expansion of interlaced Young diagrams with a maximum of \(n\) rows. 
For more information on this topic, see~\cite{Okounkov:2003sp, Okounkov2001}

%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%

\subsection{Sum over partitions}

The tau-function is a function defined on the orbit of the vacuum
\(\ket{\bm{0}}\) under the action of the group element \(g\in
\widehat{GL}(\infty)\). It is defined as the Vacuum Expectation Value (VEV)
\begin{equation}
  \tau(\mathbf{t},g):= \bra{\bm{0}} e^{\mathbf{J}_+(\mathbf{t}) } g
  \ket{\bm{0}}\; , \qquad g\in \widehat{GL}(\infty)\; .
\end{equation}
This object is intrumental in the study of several integrable systems.
Additionally, we have seen that the coherent state
\(e^{\mathbf{J}_-(\mathbf{t})} \ket{\bm{0}}\) can be written in terms
of Schur polynomials. Therefore, we want to study some aspects of
these expectation values.

Let us first define the following fermionic state
\begin{equation}
\begin{split}
   \ket{\widetilde{\Psi}_0} & = \lim_{d\to \infty} \left[ 
   \sum_{\bm{a}^{<}_d;\bm{b}^{<}_d}
   q^{\frac{1}{2}\sum_{i=1}^d (a_i+b_i)} \prod_{i=1}^d \psi^\star_{a_i}\psi_{-b_i}
   \right] |\mathtt{half}\rangle \\ 
   & = q^{L_0/2} \left( 1 + \sum_{a_1; b_1}\psi^\star_{a_1} \psi_{-b_1} + 
   \sum_{\bm{a}^{<}_2 ; \bm{b}^{<}_2} \psi^\star_{a_2} \psi_{-b_2} \psi^\star_{a_1}\psi_{-b_1} + \cdots 
   \right) |\mathtt{half}\rangle 
\end{split}
\end{equation}
where \(\bm{a}^{<}_d = \{ (a_1, \dots, a_d)\ | \ 0<a_1<\cdots < a_d\}\)
and \(\bm{b}^{<}_d = \{ (b_1, \dots, b_d)\ | \ 0<b_1<\cdots < b_d\}\).
We can immediately write this state as
\begin{equation}
\ket{\Psi_0}_f = \sum_{\lambda} q^{\frac{1}{2}|\lambda|} \ket{\lambda} \; ,
\quad \textrm{where}\quad 
\ket{\lambda} = \prod_{j=1}^{d}\psi^\star_{a_j} \psi_{-b_j} \ket{\bm{0}}\; .
\end{equation}
Diagrammatically, this state is the sum
\begin{equation}
\ket{\widetilde{\Psi}_0} =  \ket{\bm{0}} + 
\ytableausetup{centertableaux, smalltableaux}
q^{1/2} |\ydiagram{1} \rangle +
q  \left|\ydiagram{1,1} \right\rangle +
q |\ydiagram{2} \rangle+ 
q^{3/2} \left| \ydiagram{3} \right\rangle+ q^{3/2} \left| \ydiagram{1,2} \right\rangle+
q^{3/2} \left| \ydiagram{1,1,1} \right\rangle+\cdots
\end{equation}
From the fermion-boson correspondence we have discussed earlier, one
can write this state in terms of bosonic states \(\ket{\vec{k}}\) as
\begin{equation}
   \ket{\widetilde{\Psi}_0}  = q^{L_0/2} \sum_{\lambda} \ket{\lambda} = 
q^{L_0/2} \sum_{\vec{k}} \sum_{\lambda} \frac{\chi_\lambda[C(\vec{k})]}{z_{\vec{k}}} \ket{\vec{k}}
\end{equation}
where \(|\lambda| = \sum_j j k_j\).

Using that under the fermion-boson correspondence, the bosonic states might also be
written in terms of conjugacy classes of the symmetric group
\(\mathfrak{S}_n\) (or the partitions of \(n\)), let us write 
\begin{equation}
  \label{eq:bos:states}
  \begin{split}
    \ket{\vec{k}}  & = \ket{(k_1, k_2, \dots, k_m)}  
    = \ket{(1^{k_1} 2^{k_2} 3^{k_3}\dots m^{k_m})} \\
    & \equiv | (\underbrace{m, \dots, m}_{k_m},m-1, \dots ,3, \underbrace{2, \dots, 2}_{k_2},
  \underbrace{1, \dots, 1}_{k_1}, 0\dots)\rangle \equiv
  \kket{\lambda}\; ,
  \end{split}
\end{equation}
where we use the double notation \(\kket{\lambda}\) to denote the bosonic partition. 
These states are not normalized, and in fact, they satisfy
\begin{equation}
\bracket{\vec{k}'}{\vec{k}} = z_{\vec{k}} \delta_{\vec{k}', \vec{k}}\; .
\end{equation}
Therefore, one can define the state
\begin{equation}
\ket{\Psi_0} = q^{L_0/2} \sum_{k\geq 0} \sum_{\vec{k}} \frac{1}{\sqrt{z_{\vec{k}}}} \ket{\vec{k}}\; . 
\end{equation}
Using the bosonization formula, we can write this state as
\begin{equation}
  \ket{\Psi_0} = q^{L_0/2} \sum_{\lambda} \sum_{\vec{k}}
  \frac{1}{\sqrt{z_{\vec{k}}}} \chi_\mu[C(\vec{k})] \ket{\lambda}\; ,
\end{equation}
and the number of boxes in the partition is given by \(|\lambda| = \sum_j j k_j\).
Using the explicit expression for the characters, we have that the
first terms in this expansion are
\begin{equation}
  \begin{split}
\ket{\Psi_0} & = \ket{\bm{0}} + 
\ytableausetup{centertableaux, smalltableaux}
q^{1/2} |\ydiagram{1} \rangle +
q \sqrt{2} |\ydiagram{2} \rangle+ 
q^{3/2} \left[ \frac{1}{\sqrt{6}} (1 + \sqrt{2} + \sqrt{3})\left| \ydiagram{3} \right\rangle + \right. \\
& + \left. \frac{1}{\sqrt{6}} (2 - \sqrt{2}) \left| \ydiagram{1,2} \right\rangle+
 \frac{1}{\sqrt{6}} (1 + \sqrt{2} - \sqrt{3})\left| \ydiagram{1,1,1} \right\rangle\right] + \\
& + q^2 \left[ \frac{1}{12}\left( \sqrt{6} + 3 \sqrt{2} + 4 \sqrt{3}
  + 12\right) \left| \ydiagram{4} \right\rangle 
  + \frac{1}{4}\left( \sqrt{6} - \sqrt{2} \right)\left| \ydiagram{1,3} \right\rangle \right.\\
  & + \left. \frac{1}{2}\left( \sqrt{2} - \frac{2\sqrt{3}}{3} + \frac{\sqrt{6}}{3} \right)
  \left| \ydiagram{2,2} \right\rangle + 
  \frac{1}{4}\left( \sqrt{6} - \sqrt{2} \right) \left| \ydiagram{1,1,2} \right\rangle + 
  \frac{1}{12}\left( \sqrt{6} + 3 \sqrt{2} + 4 \sqrt{3}
  - 12\right) \left| \ydiagram{1,1,1,1} \right\rangle + \dots
  \right]
  \end{split}
\end{equation}

\subsection{Vacuum Expectation Values and Euler Formula}

From the orthonormality of the fermionic plane partition states, we have
\begin{equation}
\begin{split}
  Z & = \bracket{\widetilde{\Psi}_0}{\widetilde{\Psi}_0}\\
  & = 1 + q + 2 q^2 + 3 q^3 + 5 q^4 + \dots = \sum_{n\geq 0} p(n) q^n \\
  & = \prod_{n=1}^\infty \frac{1}{1-q^n}\; .
\end{split}
\end{equation}
that is the Euler formula for integer partitions. It is also easy to see that 
\begin{equation}
\begin{split}
  Z & = \bracket{\Psi_0}{\Psi_0}\\
  & = 1 + q + 2 q^2 + 3 q^3 + 5 q^4 + \dots = \sum_{n\geq 0} p(n) q^n \\
  & = \prod_{n=1}^\infty \frac{1}{1-q^n}\; .
\end{split}
\end{equation}

%\(\clubsuit\) It would be very interesting to see what is the subgroup of
%\(\widehat{\mathfrak{gl}}(\infty)\) that keeps this ground state invariant. 
% See Macmahon function here~\cite{Okounkov:2003sp, Okounkov2001, Dijkgraaf:2008ua}.

### 2024-04-11 draft.tex
work on the phase model

@@
 \address{\noindent 
-Instituto de Fíısica Teórica, UNESP-Universidade Estadual Paulista
+Instituto de Física Teórica, UNESP-Universidade Estadual Paulista
 R. Dr. Bento T. Ferraz 271, Bl. II, Sao Paulo 01140-070, SP, Brazil\\
@@
 
-\email{\texttt{\href{tharaujo@if.usp.br}{tharaujo@if.usp.br}}}
+\email{\texttt{\href{tr.araujo@unesp.br}{tr.araujo@unesp.br}}}
 
@@
 %*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*
-\section{q-Bosons and the Phase Model}
+\section{Phase Model and q-Bosons}
 %*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*
 
-This section intruduces the \(q\)-boson and phase
+This section introduces the \(q\)-boson and phase
 models~\cite{Bogoliubov:1992, Bogoliubov:1997soj, Bogoliubov2005,
-  Bogoliubov:1997soj, Tsilevich:2006}. We mainly follow the notation
-of~\cite{Wheeler:2010vmq}.
+  Bogoliubov:1997soj, Tsilevich:2006}, and consists of a general
+review of the literature. Consequently, there is nothing particularly
+original, except its presentation.  In spite of that, I believe that
+it has some pedagoginal value, and a good collection of references.
+Here we follow some conventions of~\cite{Wheeler:2010vmq}, but with a
+presentation that is closer to~\cite{Tsilevich:2006}.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \end{equation}
-where \(\bar{\mathcal{N}} = \sum_{i=0}^M \mathcal{N}_i\), and we also
-impose periodic boundary conditions \(\phi_{M+1} = \phi_0\) and
+where \(\bar{\mathcal{N}} = \sum_{i=0}^M \mathcal{N}_i\), and
+periodic boundary conditions \(\phi_{M+1} = \phi_0\) and
 \(\phi_{M+1}^\dagger = \phi_0^\dagger\).
@@
 \begin{equation}
-  \ket{\Psi(y_1, \dots, y_N)} = \prod_{j=1}^N B(y_j) \ket{\bm{0}}\qquad 
-  \bra{\Psi(y_1, \dots, y_N)} = \bra{\bm{0}}\prod_{j=1}^N C(y_j) \; .
+  \ket{\widehat{\Psi}(y_1, \dots, y_N)} = \prod_{j=1}^N B(y_j) \ket{\bm{0}}\qquad 
+  \bra{\widehat{\Psi}(y_1, \dots, y_N)} = \bra{\bm{0}}\prod_{j=1}^N C(y_j) \; .
 \end{equation}
@@
 only consider \emph{off-shell} states.
+
+\subsubsection{Coordinate expansion}
+From the map defined in~\cite{Wheeler:2010vmq}, we can associate to
+each state \(\ket{\lambda}\), a Schur function \(s_\lambda(\vec{x})\),
+where the dummy coordinates are \(\vec{x} = (x_1, \dots, x_N)\), with
+\(N = \sum_{j=0}^M n_j\). Therefore, we can work in the ring of
+symmetric functions \(\Lambda_M\), with basis given by Schur functions
+\(s_\lambda\) and Young diagrams such that \(\max(\lambda_1) = M\),
+and \(\max(\ell(\lambda)) = N\).
+
+That also means that the operator \(B\) should also be realized in the
+space \(\Lambda_M\), see~\cite{Tsilevich:2006}. The operators
+\(\phi^\dagger_j\) and \(\phi_j\) are, respectively, \(1\)-particle
+creation and annihilation at the site \(j\). In terms of partitions,
+we can think of them as insertion and removal of a row of length
+\(j\). That is, given the state \(\ket{\lambda} =\ket{1^{n_1} 2^{n_2}\cdots}\), 
+we have
+\begin{equation}
+\begin{split}
+  \phi^\dagger_j \ket{\lambda} & = \ket{1^{n_1} 2^{n_2}\cdots j^{j+1}\cdots} \equiv \ket{\lambda \oplus (j)}\\
+  \phi_j \ket{\lambda} & = \ket{1^{n_1} 2^{n_2}\cdots j^{j-1}\cdots}
+  \equiv \ket{\lambda \ominus (j)}\; , 
+\end{split}
+\end{equation}
+Once again, we see that the case \(j=0\) is meaningless in the
+partition construction. Given two partitions \(\lambda\) and \(\mu\),
+we say that \(\lambda\) interlaces \(\mu\) if \(\lambda_i \geq \mu_i \geq \lambda_{i+1}\), 
+and we write \(\lambda\succ \mu\). Therefore, it is easy to see that the partition 
+\(\lambda\oplus (j)\) interlaces \(\lambda\), and \(\lambda\)
+interlaces \(\lambda \ominus (j)\). We write this condition as 
+\(\lambda \oplus (j) \succ \lambda \succ \lambda \ominus (j)\).
+
+Now, we can use the explicit expression for the \(L\)-matrix to build
+the monodromy matrix, and extract the operator \(B(x)\) as
+\begin{subequations}
+\begin{equation}
+  B(x)_a = T_a(x)_{12} = \sum_{j_1, \dots, j_M = 1,2} L_{a j_M}(x) L_{j_M j_{M-1}}(x)\cdots  L_{j_1 2}(x)\; ,
+\end{equation}
+and it is easy to see that it has the form  
+\begin{equation}
+  B(x) \equiv x^{-M/2} \mathbb{B}(x) = x^{-M/2} \sum_{k=0}^M x^k \mathcal{B}_k \; .
+\end{equation}
+The problem now is to find the action of the operators \(\mathcal{B}_k\) on
+states \(\ket{\lambda}\).
+\end{subequations}
+Let us consider some simple cases.
+
+\textbullet \ M=0: We have just one site in the  
+chain, and all particles site at the site \(0\), as we have discussed many times, 
+does not contribute to the partition representation.
+
+\textbullet \ M=1: Now we have 1 row Young diagrams \(\lambda = (1^{n_1})\), where each row
+corresponds to a boson. It is easy to show that 
+\begin{equation}
+ \mathbb{B}_{M=1}(x) =  x \phi_1^\dagger + \phi_0^\dagger \; ,
+\end{equation}
+therefore, \(\mathcal{B}_0 = \phi_0^\dagger\) and  \(\mathcal{B}_1 = \phi_1^\dagger\). 
+From this expression, we have 
+\begin{equation}
+  \mathcal{B}_0\ket{(1^{n_1})} = \ket{(1^{n_1})} \qquad  
+  \mathcal{B}_1\ket{(1^{n_1})} = \ket{(1^{n_1+1})} \; .
+\end{equation}
+All in all, we conclude that 
+\begin{equation}
+ \mathbb{B}_{M=1}(x)\ket{\lambda}  =  \ket{\lambda} + x\ket{\lambda \oplus (1)} \;.
+\end{equation}
+
+\textbullet \ M=2: Now we have 2 row Young diagrams \(\lambda = (1^{n_1}2^{n_2})\).
+Now, one can show that
+\begin{equation}
+  \mathbb{B}_{M=2}(x) = x^2 \phi_2^\dagger + x \left( \phi_1^\dagger
+  + \phi_2^\dagger \phi_1 \phi_0^\dagger \right) + \phi_0^\dagger \; ,
+\end{equation}
+therefore,
+\begin{equation}
+\mathcal{B}_0 = \phi_0^\dagger \qquad 
+\mathcal{B}_1 = \phi_1^\dagger + \phi_2^\dagger \phi_1 \phi_0^\dagger \qquad 
+\mathcal{B}_2 = \phi_2^\dagger \; .
+\end{equation}
+
+Consequently. we find
+\begin{subequations}
+\begin{equation}
+\begin{split}
+  \mathcal{B}_0\ket{(1^{n_1}2^{n_2})} & = \ket{(1^{n_1} 2^{n_2})} \qquad  
+  \mathcal{B}_1\ket{(1^{n_1}2^{n_2})} = \ket{(1^{n_1+1} 2^{n_2})} + \ket{(1^{n_1-1} 2^{n_2 + 1})} \\
+  \mathcal{B}_2\ket{(1^{n_1}2^{n_2})} & = \ket{(1^{n_1}2^{n_2+1})}\; .
+\end{split}
+\end{equation}
+\end{subequations}
+Observe that the action of the operator \(\mathbb{B}_k(x)\) on \(\ket{\lambda}\)
+generates diagrams \(\ket{\mu}\) such that \(k = |\mu| - |\lambda|\). 
+\begin{equation}
+  \mathbb{B}_{M=2}(x) \ket{\lambda}  =
+  \sum_{\lambda \prec \mu \subseteq [2,\ell(\lambda) + 1]} x^{|\mu|-|\lambda|} \ket{\mu}\; .
+\end{equation}
+
+One can keep doing this game, and we reach to the general conclusion that 
+\begin{equation}
+  \mathbb{B}(x) \ket{\lambda}  =
+  \sum_{\lambda \prec \mu \subseteq [M,\ell(\lambda) + 1]} x^{|\mu|-|\lambda|} \ket{\mu}\; .
+\end{equation}
+
+Therefore, we conclude that the action of the operators
+\(\mathcal{B}_k\) on \(\ket{\lambda}\) generates diagrams
+\(\ket{\mu}\) that are obtained from \(\lambda\) by adding \(k\) boxes,
+not in different columns. From the viewpoint of symmetric functions, this rule
+is exactly the Pieri's rule~\cite{Macdonald:1998}, that states that
+we have that
+\begin{equation}
+  s_{(k)} s_\lambda = \sum_{\mu \succ \lambda} s_\mu\;.
+\end{equation}
+
+Therefore, we have the map from the spin chain Hilbert space to the
+ring of symmetric functions given by \(\imath : \mathcal{B}_k \mapsto
+s_{(k)}\) where \(s_{(k)}\) is the Schur function with 1 row of length
+\(k\).
 
@@
 where we write the deformation parameter as \(Q = q^{-2}\). 
-In the limit \(Q\to 1\), the q-bosons are ordinary bosons. 
-In the limit \(q\to \infty\), we recover the phase model we discussed above.
+In the limit \(Q\to 1\), the q-bosons are ordinary bosons, and 
+the limit \(Q\to 0\) (\(q\to \infty\)) gives the phase model we have discussed above.
 
 \subsubsection{Representation}
+We build the representation of the q-boson algebra with some modifications of the phase 
+model we have just discussed. We suppose that the action of the operators \(\{b_i, b_i^\dagger\}\) 
+satisfies 
+\begin{subequations}
+\begin{alignat}{3}
+    & b_0\ket{n_0} =(1 - \delta_{0, n_0}) \frac{1 - Q^{n_0}}{(1 - Q)^{1/2}}\ket{n_0 - 1}
+    &\hspace{0.5cm}& b_0^\dagger \ket{n_0} =  \frac{1}{(1 - Q)^{1/2}}\ket{n_0 + 1}  \\
+    & b_i\ket{n_i} = \frac{(1 - \delta_{0, n_i})}{(1 - Q)^{1/2}}\ket{n_i - 1}
+    && b_i^\dagger \ket{n_0} =  \frac{(1 - Q^{n_i+1})}{(1 - Q)^{1/2}}\ket{n_0 + 1}\quad i\neq 0
+\end{alignat}
+\end{subequations}
+where the modes \(\{b_0, b_0^\dagger \}\) satisfies a slightly different relation.
 
 \subsubsection{Bethe Ansatz}
+The \(L\)-operator for the q-boson is given by
+\begin{equation}
+  L_n(x, Q) =
+  \begin{pmatrix}
+    x^{-1/2} & (1 - Q)^{1/2} b_n^\dagger \\ (1 - Q)^{1/2} b_n & x^{1/2}
+  \end{pmatrix}\; .
+\end{equation}
+The monodromy matrix is 
+\begin{equation}
+  T_a(x,Q) = L_{am}(x, q)  \dots  L_{a0}(x, q) = 
+  \begin{pmatrix}
+    A(x) & B(x) \\ C(x) & D(x)
+  \end{pmatrix}\; .
+\end{equation}
+
+The eigenstates of the Hamiltonian are of the form
+\begin{equation}
+  \ket{\Psi(y_1, \dots, y_N; Q)} = \prod_{j=1}^N B(y_j, Q) \ket{\bm{0}}\qquad 
+  \bra{\Psi(y_1, \dots, y_N; Q)} = \bra{\bm{0}}\prod_{j=1}^N C(y_j, Q) \; .
+\end{equation}
+When the set \(\{ y_j \ | \ j =1, \dots , N\}\) satisfies the Bethe equations 
+\begin{equation}
+  y^{N + M}_i =\prod_{\substack{j = 1 \\ j \neq i}}^N\frac{Q y_i - y_j}{y_i - Q y_j}\; , \qquad i = 1, \dots, N\; , 
+\end{equation}
+Once again, we will consider \emph{off-shell} Bethe states.
 
@@
 
-\subsection{Bethe Ansatz for the q-boson}
-
-The \(L\)-operator for the q-boson is given by
-\begin{equation}
-  L_n(\lambda) =
-  \begin{pmatrix}
-    e^\lambda & \chi B_n^\dagger \\ \chi B_n & e^{-\lambda}
-  \end{pmatrix}
-\end{equation}
-where \(\lambda \in \mathbb{C}\) is the spectral parameter and \(\chi
-= \sqrt{1 - q^{-2}}\). From this expression, we have the monodromy matrix
-\begin{equation}
-  T(\lambda) = \overleftarrow{\prod_{n = 1}^M} L_n(\lambda) =
-  \begin{pmatrix}
-    A(\lambda) & B(\lambda) \\ C(\lambda) & D(\lambda)
-  \end{pmatrix}
-\end{equation}
-The eigenstates of the Hamiltonian are of the form
-\begin{equation}
-\ket{ \psi_N(\lambda_1, \dots, \lambda_N)} = \prod_{j=1}^N B(\lambda_j) \ket{0}\; , 
-\end{equation}
-where the parameters \(\{\lambda_i\}_{i=1}^N\) satisfies the Bethe
-Algebraic Equation
-\begin{equation}
-  e^{2M\lambda_j} = \prod_{k\neq j}^N \frac{f(\lambda_k, \lambda_j)}{f(\lambda_j, \lambda_k)}\quad \textrm{with} \quad
-  f(\lambda, \mu) = \frac{\sinh(\lambda-\mu+\gamma)}{ \sinh(\lambda- \mu)}
-  \quad j = 1,\dots \ldots , N\; .
-\end{equation}
-where
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\subsection{Bethe Ansatz for the phase model}
-
-Taking the limit \(\gamma\to \infty\) we obtain the ABA for the phase model.
-In particular, we find
-\begin{equation}
-  L_n(\lambda) =
-  \begin{pmatrix}
-    e^{ip/2} & \phi_n^\dagger \\ \phi_n & e^{-ip/2}
-  \end{pmatrix}
-\end{equation}
-where we write \(\lambda = i p/2\). The Algebraic Bethe Equations are
-\begin{equation}
-\exp \left(i (M + N) p_j \right) = (-1)^{N-1} \exp \left( i \sum_{k=1}^N p_k \right)\quad j = 1, \dots, N\; ,
-\end{equation}
-that can be solved explicitly as
-\begin{equation}
-  \label{}
-p_j = \frac{ 2 \pi I_j + \sum_{k=1}^N p_k}{M+N}\; , 
-\end{equation}
-where \(i_j \in \mathbb{z}\) ( \( + 1/2\)) if \(n\) is odd (even).
-
-In what follows, we consider the transformation \(x^{-1} = e^{ip}\), see~\cite{Wheeler:2010vmq},
-in such a way that the \(L\)-operator becomes
-\begin{equation}
-  L_{an}(\lambda) =
-  \begin{pmatrix}
-    x^{-1/2} & \phi_n^\dagger \\ \phi_n & x^{1/2} 
-  \end{pmatrix}_a
-\end{equation}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\subsubsection{Bethe states}
-
-Now we also consider the Bethe states (and their dual) 
-\begin{equation}
-\ket{\Psi(\bm{x})} = \prod_{j=1}^N B(x_j) \ket{\bm{0}}\quad \; ,
-\bra{\Psi(\bm{x})} = \bra{\bm{0}}\prod_{j=1}^N C(x_j) \; ,
-\end{equation}
-where we have written \(\bm{x} = (x_1, \dots, x_N)\), and the monodromy matrix
-is given by
-\begin{equation}
-  T_a(x) = \prod_{j=1}^M L_{a j}(x) = 
-  \begin{pmatrix}
-    A(x) & B(x) \\
-    C(x) & D(x)
-  \end{pmatrix}_a \; .
-\end{equation}
-Folowing~\cite{Wheeler:2010vmq}, we define \(\mathbb{B} = x^{M/2} B(x)\) and 
-\(\mathbb{C} = x^{M/2} C(x^{-1})\).
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\subsection{Scalar product}
+%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*
+\section{Toda tau functions and the phase model}
+%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*
 

### 2024-04-12 draft.tex
work on the draft

@@
 
-\title{Comments on \(Q\)-Bosons, deformed fermions and their relations
-    with integrable hierarchies}
+\title[\(Q\)-Bosons and relations with integrable hierarchies]{Comments on \(Q\)-Bosons,
+and their relations with integrable hierarchies}
 
@@
 
+\begin{abstract}
+In this work, we study aspects of the \(Q\)-boson model, and 
+its relations with integrable hierarchies. 
+\vspace{4pt}
+\noindent \textbf{Keywords:} Integrability, Q-bosons, Schur
+
 \keywords{integrability, spin-chains}
@@
 
-\begin{abstract}
-  In this work, we study aspects of the \(Q\)-boson model, and 
-its relations with integrable hierarchies. 
-% \vspace{4pt}
- \noindent \textbf{Keywords:} Integrability, Q-bosons, Schur
 \end{abstract}
@@
 
-\setcounter{tocdepth}{1}
+\setcounter{tocdepth}{3}
 \tableofcontents
@@
 
-With these expressions, one can finally build the Bethe states and their dual 
+With these expressions, one can finally build the Bethe states
 \begin{equation}
-  \ket{\widehat{\Psi}(y_1, \dots, y_N)} = \prod_{j=1}^N B(y_j) \ket{\bm{0}}\qquad 
-  \bra{\widehat{\Psi}(y_1, \dots, y_N)} = \bra{\bm{0}}\prod_{j=1}^N C(y_j) \; .
+  \ket{\Psi(y_1, \dots, y_N)} = \prod_{j=1}^N \mathbb{B}(y_j) \ket{\bm{0}}
 \end{equation}
-When the set \(\{ y_j \ | \ j =1, \dots , N\}\) satisfies the Bethe equations 
+where \(\mathbb{B}(x) = x^{M/2} B(x)\). When the set \(\{ y_j \ | \ j
+=1, \dots , N\}\) satisfies the Bethe equations
 \begin{equation}
@@
 only consider \emph{off-shell} states.
-
-\subsubsection{Coordinate expansion}
-From the map defined in~\cite{Wheeler:2010vmq}, we can associate to
-each state \(\ket{\lambda}\), a Schur function \(s_\lambda(\vec{x})\),
-where the dummy coordinates are \(\vec{x} = (x_1, \dots, x_N)\), with
-\(N = \sum_{j=0}^M n_j\). Therefore, we can work in the ring of
-symmetric functions \(\Lambda_M\), with basis given by Schur functions
-\(s_\lambda\) and Young diagrams such that \(\max(\lambda_1) = M\),
-and \(\max(\ell(\lambda)) = N\).
-
-That also means that the operator \(B\) should also be realized in the
-space \(\Lambda_M\), see~\cite{Tsilevich:2006}. The operators
-\(\phi^\dagger_j\) and \(\phi_j\) are, respectively, \(1\)-particle
-creation and annihilation at the site \(j\). In terms of partitions,
-we can think of them as insertion and removal of a row of length
-\(j\). That is, given the state \(\ket{\lambda} =\ket{1^{n_1} 2^{n_2}\cdots}\), 
-we have
-\begin{equation}
-\begin{split}
-  \phi^\dagger_j \ket{\lambda} & = \ket{1^{n_1} 2^{n_2}\cdots j^{j+1}\cdots} \equiv \ket{\lambda \oplus (j)}\\
-  \phi_j \ket{\lambda} & = \ket{1^{n_1} 2^{n_2}\cdots j^{j-1}\cdots}
-  \equiv \ket{\lambda \ominus (j)}\; , 
-\end{split}
-\end{equation}
-Once again, we see that the case \(j=0\) is meaningless in the
-partition construction. Given two partitions \(\lambda\) and \(\mu\),
-we say that \(\lambda\) interlaces \(\mu\) if \(\lambda_i \geq \mu_i \geq \lambda_{i+1}\), 
-and we write \(\lambda\succ \mu\). Therefore, it is easy to see that the partition 
-\(\lambda\oplus (j)\) interlaces \(\lambda\), and \(\lambda\)
-interlaces \(\lambda \ominus (j)\). We write this condition as 
-\(\lambda \oplus (j) \succ \lambda \succ \lambda \ominus (j)\).
-
-Now, we can use the explicit expression for the \(L\)-matrix to build
-the monodromy matrix, and extract the operator \(B(x)\) as
-\begin{subequations}
-\begin{equation}
-  B(x)_a = T_a(x)_{12} = \sum_{j_1, \dots, j_M = 1,2} L_{a j_M}(x) L_{j_M j_{M-1}}(x)\cdots  L_{j_1 2}(x)\; ,
-\end{equation}
-and it is easy to see that it has the form  
-\begin{equation}
-  B(x) \equiv x^{-M/2} \mathbb{B}(x) = x^{-M/2} \sum_{k=0}^M x^k \mathcal{B}_k \; .
-\end{equation}
-The problem now is to find the action of the operators \(\mathcal{B}_k\) on
-states \(\ket{\lambda}\).
-\end{subequations}
-Let us consider some simple cases.
-
-\textbullet \ M=0: We have just one site in the  
-chain, and all particles site at the site \(0\), as we have discussed many times, 
-does not contribute to the partition representation.
-
-\textbullet \ M=1: Now we have 1 row Young diagrams \(\lambda = (1^{n_1})\), where each row
-corresponds to a boson. It is easy to show that 
-\begin{equation}
- \mathbb{B}_{M=1}(x) =  x \phi_1^\dagger + \phi_0^\dagger \; ,
-\end{equation}
-therefore, \(\mathcal{B}_0 = \phi_0^\dagger\) and  \(\mathcal{B}_1 = \phi_1^\dagger\). 
-From this expression, we have 
-\begin{equation}
-  \mathcal{B}_0\ket{(1^{n_1})} = \ket{(1^{n_1})} \qquad  
-  \mathcal{B}_1\ket{(1^{n_1})} = \ket{(1^{n_1+1})} \; .
-\end{equation}
-All in all, we conclude that 
-\begin{equation}
- \mathbb{B}_{M=1}(x)\ket{\lambda}  =  \ket{\lambda} + x\ket{\lambda \oplus (1)} \;.
-\end{equation}
-
-\textbullet \ M=2: Now we have 2 row Young diagrams \(\lambda = (1^{n_1}2^{n_2})\).
-Now, one can show that
-\begin{equation}
-  \mathbb{B}_{M=2}(x) = x^2 \phi_2^\dagger + x \left( \phi_1^\dagger
-  + \phi_2^\dagger \phi_1 \phi_0^\dagger \right) + \phi_0^\dagger \; ,
-\end{equation}
-therefore,
-\begin{equation}
-\mathcal{B}_0 = \phi_0^\dagger \qquad 
-\mathcal{B}_1 = \phi_1^\dagger + \phi_2^\dagger \phi_1 \phi_0^\dagger \qquad 
-\mathcal{B}_2 = \phi_2^\dagger \; .
-\end{equation}
-
-Consequently. we find
-\begin{subequations}
-\begin{equation}
-\begin{split}
-  \mathcal{B}_0\ket{(1^{n_1}2^{n_2})} & = \ket{(1^{n_1} 2^{n_2})} \qquad  
-  \mathcal{B}_1\ket{(1^{n_1}2^{n_2})} = \ket{(1^{n_1+1} 2^{n_2})} + \ket{(1^{n_1-1} 2^{n_2 + 1})} \\
-  \mathcal{B}_2\ket{(1^{n_1}2^{n_2})} & = \ket{(1^{n_1}2^{n_2+1})}\; .
-\end{split}
-\end{equation}
-\end{subequations}
-Observe that the action of the operator \(\mathbb{B}_k(x)\) on \(\ket{\lambda}\)
-generates diagrams \(\ket{\mu}\) such that \(k = |\mu| - |\lambda|\). 
-\begin{equation}
-  \mathbb{B}_{M=2}(x) \ket{\lambda}  =
-  \sum_{\lambda \prec \mu \subseteq [2,\ell(\lambda) + 1]} x^{|\mu|-|\lambda|} \ket{\mu}\; .
-\end{equation}
-
-One can keep doing this game, and we reach to the general conclusion that 
-\begin{equation}
-  \mathbb{B}(x) \ket{\lambda}  =
-  \sum_{\lambda \prec \mu \subseteq [M,\ell(\lambda) + 1]} x^{|\mu|-|\lambda|} \ket{\mu}\; .
-\end{equation}
-
-Therefore, we conclude that the action of the operators
-\(\mathcal{B}_k\) on \(\ket{\lambda}\) generates diagrams
-\(\ket{\mu}\) that are obtained from \(\lambda\) by adding \(k\) boxes,
-not in different columns. From the viewpoint of symmetric functions, this rule
-is exactly the Pieri's rule~\cite{Macdonald:1998}, that states that
-we have that
-\begin{equation}
-  s_{(k)} s_\lambda = \sum_{\mu \succ \lambda} s_\mu\;.
-\end{equation}
-
-Therefore, we have the map from the spin chain Hilbert space to the
-ring of symmetric functions given by \(\imath : \mathcal{B}_k \mapsto
-s_{(k)}\) where \(s_{(k)}\) is the Schur function with 1 row of length
-\(k\).
 
@@
 %*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*
-\section{Toda tau functions and the phase model}
+\section{Tau functions in the phase model}
 %*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*
 
-Bogoliubov~\cite{Bogoliubov2005}, in the notation~\cite{Wheeler:2010vmq},
-has shown that the the scalar product of two vectors in the
-\(N\)-particles sector in a chain of length \(M + 1\) is
+This section discusses presence of integrable hierarchies in the phase
+model.  We argue that it is also a Toda hierachy tau function, and we
+discuss some consequences of this fact, in particular, its connection
+with a matrix model. Moreover, we also point the presence of
+correlation functions in this model that also satisfy the KP hierarchy
+equations. We finally argue how these results also imply that these
+scalar products are also KP hierarchy tau functions, what is expected from 
+the results of Wheeler~\cite{Wheeler:2010vmq}.
+
+Bogoliubov~\cite{Bogoliubov2005}, has shown that the scalar product of
+two vectors in the \(N\)-particles sector in a chain of length \(M+1\)
+is
 \begin{equation}
@@
 
-This scalar product is related to the counting of plane
-partitions, with
-\begin{equation}
- S(N, M| \bm{x}, \bm{y}) 
- = \sum_{\lambda \subseteq (N, N, M)} s_{\lambda}(\bm{x}) s_{\lambda} (\bm{y})\; .
-\end{equation}
-In the limit \(M, N \to \infty\), one can use the Cauchy identity
-\begin{equation}
-  \sum_{\lambda } s_{\lambda}(\bm{x}) s_{\lambda}(\bm{y})
-  = \prod_{i,j>0} \frac{1}{1- x_iy_j}\; ,
-\end{equation}
-and specialize this expression to \( x_i = y_i = q^{i-1/2} \) to find
-the MacMahon function
-\begin{equation}
-  \sum_{\lambda } s_{\lambda}(\bm{q}_\star) s_{\lambda}(\bm{q}_\star) = \prod_{i,j>0} \frac{1}{1- q^{i +j -1}}\; ,
-\end{equation}
-where \(\bm{q}_{\star} = (q^{1/2}, q^{3/2}, \dots)\).
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-\subsection{KP and Toda tau functions}
+\subsection{Toda tau functions}
+In this section, we argue that the scalar product defined above is a tau function 
+of the Toda hierarchy. Let us first write the function \(H(z,w)\) as the geometric sum 
+\begin{equation}
+  H(z,w) = \frac{1 - (zw)^{M+N}}{1 - zw} = \sum_{k=1}^{M+N} (zw)^{k-1} \; , 
+\end{equation}
+in such a way that 
+we can write
+\begin{equation}
+  \det_{i,j} \left(H(x_i, y_j)\right) = \det_{i,j} \left( \sum_{k=1}^{M+N} x_i^{k-1} y_j^{k-1}\right)
+\end{equation}
+and we can thing of this expression as the product of a \(N\times (N+M)\) matrix \(\mathcal{X}\)
+and another \((M + N)\times N\) matrix \(\mathcal{Y}\), given by
+\begin{equation}
+  \mathcal{X} = 
+  \begin{pmatrix}
+  x_1^0 & x_1^1 & \dots & x_1^{M+N-1} \\  
+  x_2^0 & x_2^1 & \dots & x_2^{M+N-1} \\  
+  \vdots \\
+  x_N^0 & x_N^1 & \dots & x_N^{M+N-1} 
+  \end{pmatrix}\quad \textrm{and} \quad 
+  \mathcal{Y} = 
+  \begin{pmatrix}
+  y_1^0 & y_2^0 & \dots & y_N^0 \\  
+  y_1^1 & y_2^1 & \dots & y_N^1 \\  
+  \vdots & \vdots & & \vdots \\
+  y_1^{N+M-1} & y_2^{M+N-1} & \dots & y_N^{M+N-1} 
+  \end{pmatrix}
+\end{equation}
+therefore 
+\begin{equation}
+  \det_{i,j} \left(H(x_i, y_j)\right) =  \det_{i,j} \left( \mathcal{X}\mathcal{Y}\right)
+= \sum_{0 \leq \ell_{N} \leq \dots \leq \ell_1\leq N+M }  \det_{ik}(x_i^{\ell_k})  \det_{ik}(y_j^{\ell_k})\; .
+\end{equation}
+If we now define \(\ell_j = \lambda_k - k + N\), and
+using~(\ref{eq:scalar}), we have
+\begin{equation}
+S(N, M|\bm{x}, \bm{y}) = \sum_{\lambda\subseteq [N,M]}
+\frac{\det_{ik}(x_i^{\lambda_k - k + N})}{\Delta(\bm{x})} \frac{ \det_{ik}(y_j^{\lambda_k - k + N})}{\Delta(\bm{y})}
+= \sum_{\lambda\subseteq [N,M]} s_\lambda(\bm{x}) s_\lambda(\bm{y}) \; .
+\end{equation}
+where \(\Delta(\bm{x})\) and \(\Delta(\bm{x})\) are Vandermonde
+determinants. This formula agrees with the Schur expansion defined
+in~\cite{Bogoliubov2005}.
 
-Wheeler has shown in~\cite{Wheeler:2010vmq} that the the scalar
-product~(\ref{eq:scalar}) is a KP tau function with respect to the
-Miwa coordinates \(t_m = \frac{1}{m}\sum_j x_j^m\) (\(t'_m =
-\frac{1}{m}\sum_j y_j^m\)), where the Schur functions
-\(s_{\lambda}(\bm{y}) = c_\lambda\) ( \(s_{\lambda}(\bm{x}) =
-c'_\lambda\)) become Plücker coordinates, and vice-versa.
-The proof for this statement is relies on the map \(\mathcal{M}_\psi\)
-between the space \(\mathcal{V}\) os Bethe states and Fock
-space \(\mathcal{F}_\psi\) of free fermions.
-
-In fact, there is an easier way to see this fact. First one may see that 
+Let us write the Miwa coordinates 
 \begin{equation}
- S(N,M,\bm{y}|\{\bm{t}\}) 
- = \sum_{\lambda} s_\lambda(\bm{y})  s_\lambda(\bm{t}) 
- \equiv \sum_{\lambda} c_\lambda(\bm{y})  s_\lambda(\bm{t}) 
+  t_p = \frac{1}{p}\sum_{j=1}^N x_j^p\qquad 
+  t'_p = \frac{1}{p}\sum_{j=1}^N y_j^p\; ,
 \end{equation}
-is a tau function of the KP hierarchy, since
-\(c_\lambda(\bm{y}) = \det (h_{\lambda_i-i +j}(\bm{y}))\) are Plücker
-coordinates~\cite{Miwa2000, Foda:2009zz}. Therefore, the truncation
+therefore, we have 
 \begin{equation}
- S(N,M,\bm{y}|\{\bm{t}\}) 
- = \sum_{\lambda \subseteq [N, \infty)} s_\lambda(\bm{y})  s_\lambda(\bm{t}) 
+  S(N, M|\bm{t}, \bm{t}') = \sum_{\lambda\subseteq [N,M]} s_\lambda(\bm{t}) s_\lambda(\bm{t}') \; .
 \end{equation}
-is also a tau function. According to~\cite{Alexandrov:2012tr,
-  Kharchev:1991gd, Zabrodin:2010ii}, the truncation of the tau
-function corresponds to the inclusion of a projection operator
-\(\mathrm{P}^+\) in the expectation value of the tau function.
-% See section 3.3.2 of~\cite{Alexandrov:2012tr},
-
-\subsubsection{Scalar product}
-We also argue that this is also a tau function of the 2D toda hierarchy. 
-We can write the Cauchy's identity as
-\begin{subequations}
+This expression is known to be a tau function for \(M, N\to \infty\).
+In fact, using the free fermions representation~\cite{Alexandrov:2012tr},
+it can be written as
+\begin{equation}
+\sum_{\lambda } s_{\lambda}(\bm{t}) s_{\lambda}(\bm{t}')
+ = \bra{\bm{0}} e^{\bm{J}_+(\bm{t})} e^{-\bm{J}_-(\bm{t}')} \ket{\bm{0}}\; .
+\end{equation}
+It is a tau function of the Toda hierarchy with trivial element \(1 \in GL(\infty)\).
+In fact, it is nothing but the Cauchy's identity
 \begin{equation}
@@
 where \(\bm{t}\) and \(\bm{t}'\) are two independent sets of Miwa coordinates.
-It is a trivial tau function of the 2D Toda hierarchy, since one can write it as
+
+Putting all these facts together, the truncation for finite \(M\) and
+\(N\) are also tau functions of the Toda hierarchy. More specifically,
+according to~\cite{Alexandrov:2012tr, Kharchev:1991gd,
+  Zabrodin:2010ii}, the truncation of the tau function corresponds to
+the inclusion of a projection operator \(\mathrm{P}^+\) in the
+expectation value of the tau function written in the fermionic
+representation.
+
+\subsubsection{KP tau function}
+Finally, one may also observe that if we fix one set of coordinates,
+say \(\bm{y}\), then we write
 \begin{equation}
-\sum_{\lambda } s_{\lambda}(\bm{t}) s_{\lambda}(\bm{t}')
- = \bra{\bm{0}} e^{\bm{J}_+(\bm{t})} e^{-\bm{J}_-(\bm{t}')} \ket{\bm{0}}
+ S(N,M|\{\bm{t}\}) 
+ = \sum_{\lambda} s_\lambda(\bm{y})  s_\lambda(\bm{t}) 
+ \equiv \sum_{\lambda} c_\lambda(\bm{y})  s_\lambda(\bm{t}) \; ,
 \end{equation}
-and from the free-fermions representation of the tau functions,
-see~\cite{Alexandrov:2012tr}, we have that it is a tau function of the
-Toda hierarchy with trivial element \(1 \in GL(\infty)\).
-\end{subequations}
+with coefficients \(c_\lambda(\bm{y}) = \det (h_{\lambda_i-i +j}(\bm{y}))\).
+But now, it is a trivial to notice that these are Plücker
+coordinates in the Jacobi-Trudi form, see~\cite{Miwa2000, Alexandrov:2012tr}.
 
-Therefore, the truncation 
+Therefore, the expression
 \begin{equation}
- S(N,M|\{\bm{t}\},\{\bm{t}'\}) = 
-  \sum_{\lambda \subseteq [N,N,M]} s_{\lambda}(\bm{t}) s_{\lambda}(\bm{t}')
+ S(N,M|\{\bm{t}\}) = \sum_{\lambda \subseteq [N, \infty)} s_\lambda(\bm{y})  s_\lambda(\bm{t}) 
 \end{equation}
-is a also a restricted tau function. These proofs follow from simple
-properties of tau functions, and it should be contrated with the rich
-formalism in~\cite{Wheeler:2010vmq}. Moreover, from~\cite{Zabrodin:2010ii}
-we have that the this tau function can be written as the matrix integral
+is also a KP tau function: A fact that we already know
+from~\cite{Wheeler:2010vmq}, where the author has proved this
+statement using the free fermions formalism.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\subsection{Matrix models}
+
+Moreover, from~\cite{Zabrodin:2010ii} we have that the this tau
+function can be written as the matrix integral
 \begin{equation}
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+In fact, we can also write
+\begin{equation}
+  H(z,w) = \sum_{k=0}^{M+N} h_k(zw)^{k-1} \; , 
+\end{equation}
+with \(h_k=1\) if \(k\in [0, M+N-1]\) and \(0\) otherwise. In this
+case, one can define a diagonal \((N+M)\times (N+M)\) matrix
+\(\mathcal{H} = \textrm{diag}(h_0, \dots, h_{M+N-1})\). Therefore, if
+we repeat the arguments above, we find
+\begin{equation}
+S(N, M|\bm{t}, \bm{t}') = \sum_{\lambda} h_\lambda s_\lambda(\bm{t}) s_\lambda(\bm{t}') \; .
+\end{equation}
+where \(h_\lambda = 1\) if \(\lambda \subset [N,M]\) and it vanishes otherwise.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-\subsubsection{Correlation functions}
+\subsection{Correlation functions}
 Bogoliubov has also shown that the correlation functions
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %\addcontentsline{toc}{section}{References}
+
 \printbibliography

### 2024-04-13 draft.tex
finish calculations fo the scalar product

@@
 \end{equation}
-from where we can build the monodromy matrix 
+where \(x \in \mathbb{S}^1\). From where we can build the monodromy
+matrix
 \begin{equation}
@@
 \begin{equation}
-  \ket{\Psi(y_1, \dots, y_N)} = \prod_{j=1}^N \mathbb{B}(y_j) \ket{\bm{0}}
+  \ket{\Psi(y_1, \dots, y_N)} = \prod_{j=1}^N \mathbb{B}(y_j) \ket{\bm{0}} \qquad 
+  \bra{\Psi(y_1, \dots, y_N)} = \bra{\bm{0}}\prod_{j=1}^N \mathbb{C}(y_j) 
 \end{equation}
-where \(\mathbb{B}(x) = x^{M/2} B(x)\). When the set \(\{ y_j \ | \ j
+where \(\mathbb{B}(x) = x^{M/2} B(x)\) and \(\mathbb{C}(x) = x^{M/2} C(1/x)\).
+When the set \(\{ y_j \ | \ j
 =1, \dots , N\}\) satisfies the Bethe equations
 \begin{equation}
+\label{eq:bethe_eq}
   y^{N + M}_i = (-1)^{N-1} \prod_{\substack{j = 1 \\ j \neq i}}^N y_j\; , \qquad i = 1, \dots, N\; , 
@@
 \begin{equation}
+\label{eq:scalar_exp}
 S(N, M|\bm{x}, \bm{y}) = \sum_{\lambda\subseteq [N,M]}
@@
 
-Let us write the Miwa coordinates 
+Let us define two sets of Miwa coordinates \(\bm{t} = (t_1, t_2, \dots)\)
+and \(\bm{t}' = (t'_{-1}, t'_{-2}, \dots)\) by
 \begin{equation}
   t_p = \frac{1}{p}\sum_{j=1}^N x_j^p\qquad 
-  t'_p = \frac{1}{p}\sum_{j=1}^N y_j^p\; ,
+  t'_{-p} = \frac{1}{p}\sum_{j=1}^N y_j^p\; ,
 \end{equation}
@@
   \sum_{\lambda } s_{\lambda}(\bm{t}) s_{\lambda}(\bm{t}')
-    = \exp \left( \sum_{m\geq1} m t_m t'_m \right) \; .
+    = \exp \left( \sum_{m\geq1} m t_m t'_{-m} \right) \; .
 \end{equation}
@@
 representation.
+
+As a final remark, we can also write
+\begin{equation}
+  H(z,w) = \sum_{k=0}^{M+N} h_k(zw)^{k-1} \; , 
+\end{equation}
+with \(h_k=1\) if \(k\in [0, M+N-1]\) and \(0\) otherwise. In this
+case, one can define a diagonal \((N+M)\times (N+M)\) matrix
+\(\mathcal{H} = \textrm{diag}(h_0, \dots, h_{M+N-1})\). Therefore, if
+we repeat the arguments above, we find
+\begin{equation}
+S(N, M|\bm{t}, \bm{t}') = \sum_{\lambda} h_\lambda s_\lambda(\bm{t}) s_\lambda(\bm{t}') \; .
+\end{equation}
+where \(h_\lambda = 1\) if \(\lambda \subset [N,M]\) and it vanishes
+otherwise.  In this case, we have that this tau function is of the
+type considered in~\cite{orlov:2001}. Evidently that it is a trivial
+example, but we expect that in the case of the \(q\)-Boson model,
+where \(q\) is finite, the diagonal terms \(h_\lambda\) are more
+interesting. We will discuss this case next section.
+
+\subsubsection{Dyson gases}
+
+It is also interesting to consider the particular case when \(M\to
+\infty\), but with a finite number of particles \(N\). The
+partititions \(\lambda \in [N, \infty]\) satisfy the condition
+\(\ell(\lambda) \leq N\), therefore, the scalar
+product~(\ref{eq:scalar_exp}) becomes
+\begin{equation}
+  S_N(\bm{t}, \bm{t}') = \sum_{\substack{\lambda \\ \ell(\lambda) \leq N}}
+  s_\lambda(\bm{t})  s_\lambda(\bm{t}') \; .
+\end{equation}
+From~\cite{Zabrodin:2010ii}, we know that this expression can be
+written as the matrix integral
+\begin{equation}
+  S_N(\bm{t}, \bm{t}') =
+  \frac{1}{N!} \prod_{\ell=1}^N \oint_{\Gamma_\ell} \frac{dz_\ell}{2 \pi i z_\ell}
+  e^{\xi(\bm{t}, z_\ell) - \xi(\bm{t}', z_{\ell}^{-1})} \Delta(z)\Delta(z^{-1})\; ,
+\end{equation}
+where
+\begin{equation}
+\xi(\bm{t}, z) = \sum_{k\geq 0} t_k z^k \qquad
+\xi(\bm{t}', 1/z) = \sum_{k\geq 0} t'_{-k} z^{-k} \; . 
+\end{equation}
+See also~\cite{Kharchev:1991gd} for a detailed proof of this relation,
+and~\cite{Orlov:2005} for other details.
+
+What is very interesting about this result is what it implies when we 
+impose the Bethe equations~(\ref{eq:bethe_eq}). Suppose that 
+\(x_j = e^{-ip_j} \in \mathbb{S}^1\), and we set \(\bm{t}' = - \bm{t}^\star\).
+In this particular case, we have
+\begin{equation}
+\xi(\bm{t}, z) - \xi(\bm{t}', 1/z)  = 2 \ \textrm{Re}\left(\sum_k t_k z^k\right)\; .
+\end{equation}
+Then, the phase model is equivalent to an ensemble of \(N\)
+2D Coulomb particles on a circle.
 
@@
 \begin{equation}
- S(N,M|\{\bm{t}\}) = \sum_{\lambda \subseteq [N, \infty)} s_\lambda(\bm{y})  s_\lambda(\bm{t}) 
+ S(N,M|\bm{t}) = \sum_{\lambda \subseteq [N,M]} s_\lambda(\bm{y})  s_\lambda(\bm{t}) \; ,
 \end{equation}
@@
 statement using the free fermions formalism.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\subsection{Matrix models}
-
-Moreover, from~\cite{Zabrodin:2010ii} we have that the this tau
-function can be written as the matrix integral
-\begin{equation}
-  S(N,M|\{\bm{t}\}, \{\bm{t}'\}) =
-  \frac{1}{N!} \prod_{\ell=1}^N \oint_{\Gamma_\ell} \frac{dz_\ell}{2 \pi i z_\ell}
-  e^{\xi_+(\bm{t}_+, z_\ell) - \xi_-(\bm{t}_-, z_{\ell}^{-1})} \Delta(z)\Delta(z^{-1})
-\end{equation}
-
-When the Bethe equations are satisfied, we have that \(x_j = y_j = e^{ip_j}\),
-therefore \(t_{-m} = - t_m^\star\), and the the phase model is equivalent to
-an ensemble of \(N\) 2D Coulomb particles on a circle.
-
-In fact, we can also write
-\begin{equation}
-  H(z,w) = \sum_{k=0}^{M+N} h_k(zw)^{k-1} \; , 
-\end{equation}
-with \(h_k=1\) if \(k\in [0, M+N-1]\) and \(0\) otherwise. In this
-case, one can define a diagonal \((N+M)\times (N+M)\) matrix
-\(\mathcal{H} = \textrm{diag}(h_0, \dots, h_{M+N-1})\). Therefore, if
-we repeat the arguments above, we find
-\begin{equation}
-S(N, M|\bm{t}, \bm{t}') = \sum_{\lambda} h_\lambda s_\lambda(\bm{t}) s_\lambda(\bm{t}') \; .
-\end{equation}
-where \(h_\lambda = 1\) if \(\lambda \subset [N,M]\) and it vanishes otherwise.
 
@@
 Since the components \(Q_{jN}\) are independent of the coordinates \(y\), 
-we cannot write the above expression in terms of a Toda hierarchy tau
+we cannot write the above expression as a Toda hierarchy tau
 function.

### 2024-04-16 draft.tex
add correlation functions

@@
 
+\usepackage{amsaddr} % affiliation in the first page
+
 \begin{document}
@@
 
-\title[\(Q\)-Bosons and relations with integrable hierarchies]{Comments on \(Q\)-Bosons,
+\title[\(Q\)-Bosons and relations with integrable hierarchies]{\(Q\)-Bosons
 and their relations with integrable hierarchies}
@@
 \address{\noindent 
-Instituto de Física Teórica, UNESP-Universidade Estadual Paulista
+Instituto de Física Teórica, UNESP-Universidade Estadual Paulista,
 R. Dr. Bento T. Ferraz 271, Bl. II, Sao Paulo 01140-070, SP, Brazil\\
-and
-Instituto de Física\\ Universidade de S\~ao Paulo\\ 
-Rua do Matão Travessa 1371, 05508-090\\ São Paulo, SP. Brazil
+\&
+Instituto de Física, Universidade de S\~ao Paulo,
+Rua do Matão Travessa 1371, 05508-090 São Paulo, SP. Brazil
 }
@@
 its relations with integrable hierarchies. 
-\vspace{4pt}
-\noindent \textbf{Keywords:} Integrability, Q-bosons, Schur
-
-\keywords{integrability, spin-chains}
-\subjclass[2020]{82B20, 82B23}
-\date{\today}
-
 \end{abstract}
 
+%\date{\today}
+\keywords{Integrability, Bethe, Schur}
+\subjclass[2020]{82B20, 82B23}
 \maketitle
 
-\setcounter{tocdepth}{3}
+\setcounter{tocdepth}{2}
 \tableofcontents
@@
 \begin{equation}
+\label{eq:h-matrix}
   H_{ij} = H(x_i, y_j) 
@@
 
-\subsubsection{Dyson gases}
+\subsubsection{Matrix Models}
 
-It is also interesting to consider the particular case when \(M\to
+It is also interesting to notice that the particular case when \(M\to
 \infty\), but with a finite number of particles \(N\). The
-partititions \(\lambda \in [N, \infty]\) satisfy the condition
-\(\ell(\lambda) \leq N\), therefore, the scalar
+partitions \(\lambda \in [N, \infty]\) satisfy the condition
+\(\ell(\lambda) \leq N\). Therefore, the scalar
 product~(\ref{eq:scalar_exp}) becomes
@@
 From~\cite{Zabrodin:2010ii}, we know that this expression can be
-written as the matrix integral
+written as the following integral
 \begin{equation}
@@
 \end{equation}
-Then, the phase model is equivalent to an ensemble of \(N\)
-2D Coulomb particles on a circle.
+Then, the phase model is equivalent to an ensemble of \(N\) 2D Coulomb
+particles on a circle. In this case, we have that the quantities
+\(\lambda_{\ell}\) are eigenvalues of a matrix \(U\).
+
+Moreover, Zabrodin~\cite{Zabrodin:2010ii}, see citations therein, we also know
+that under the rescaling \(t_k \to T_k/ \hbar\), \(t'_k \to T_{-k}/ \hbar\)
+and \(N = T_0/ \hbar\), we get the dispersion limit tau function 
+\begin{equation}
+  F_0(\bm{T}) = \log S_{T_0}(\bm{T}, \bm{T}') + \mathcal{O}(\hbar)\; ,
+\end{equation}
+a free energy, from the viewpoit of the matrix integral partition
+function.  It is not clear how one can use this fact to determine
+properties of the integrable model, but it might be possible to study
+the analytic structure of the free energy \(F_0\) to gain some
+understanding on the Bether roots \(\bm{x}\).  This problem is
+currently under further investigation, and we hope to report some new
+results elsewhere.
 
@@
 \begin{equation}
- S(N,M|\{\bm{t}\}) 
+ S(N,M|\bm{t}) 
  = \sum_{\lambda} s_\lambda(\bm{y})  s_\lambda(\bm{t}) 
@@
 \subsection{Correlation functions}
-Bogoliubov has also shown that the correlation functions
+In~\cite{Bogoliubov2005}, Bogoliubov has also shown that the
+correlation functions
 \begin{subequations}
 \begin{equation}
-  P(m|\bm{x}, \bm{y}) = \bra{0} \prod_{j=1}^N C(x) \prod_{j=1}^{N-1} B(y) \phi_m^\dagger \ket{0}
+\begin{split}
+\label{eq:correlation}
+  P_m(N, M|\bm{x}, \bm{y}\setminus \{y_N\})
+  & = \bra{0} \prod_{j=1}^N \mathbb{C}(x_j)
+  \prod_{k=1}^{N-1} \mathbb{B}(y_k) \phi_m^\dagger \ket{0}\\
+  & =  \prod_{j=1}^N x_{j}^{M/2} \prod_{k=1}^{N-1} y_{j}^{M/2}
+  \bra{0} \prod_{j=1}^N C(1/x_j) \prod_{k=1}^{N-1} B(y_k) \phi_m^\dagger \ket{0}
+\end{split}
 \end{equation}
@@
 \begin{equation}
-  P(m|\bm{x}, \bm{y}) = \frac{(-1)^{N-1}}{(\sqrt{y_N}^{N-1}}
-  \left( \prod_{t<N} \frac{x_N - y_t}{y_t} \right) \frac{\det Q}{\det H} S(N,M|\bm{x}, \bm{y})
+  P_m(N, M|\bm{x}, \bm{y}\setminus \{y_N\}) = 
+  \frac{(-1)^{N-1}}{y_N^{(N-1)/2}} \prod_{j=1}^N x_{j}^{M/2}
+  \prod_{k=1}^{N-1} y_{j}^{M/2}
+  \left( \prod_{t<N} \frac{y_N - y_t}{y_t} \right) \frac{\det Q}{\det H} S(N,M|\bm{x}, \bm{y})
 \end{equation}
@@
 \end{equation}
-where \(H_{jk} = H(x_j, y_k)\) are the components of the matrix \(H\). 
+where \(H_{jk} = H(x_j, y_k)\) are the components of the matrix \(H\) in~(\ref{eq:h-matrix}). 
 Since the components \(Q_{jN}\) are independent of the coordinates \(y\), 
@@
 
-We already know that 
+We already know from~(\ref{eq:scalar}) that 
 \begin{equation}
@@
 \end{equation}
-therefore we define
+therefore we have
 \begin{equation}
-  \mathbb{P}(m|\bm{x}, \bm{y}) =\frac{\det Q}{\Delta(x)}\; ,
+  P_m(N, M|\bm{x}, \bm{y}\setminus{y_N})  \propto
+  \frac{\det Q}{\Delta(x)} \equiv \mathcal{P}_m(N,M|\bm{x}, \bm{y})\; ,
 \end{equation}
 and we are going to treat the coordinates \(\{ \bm{y} \}\) as
-a set of \(N\) fixed constants, in a such a way that we write \(N\)
+a set of \(N\) fixed parameters. Furthermore, we write \(N\)
 functions \(\bm{F}(z) = (F_1, \dots, F_N)\) defined as
@@
   F_j (z) = H(z, y_j) \quad \textrm{if} \ j \neq N\; , \qquad \textrm{and}\qquad 
-   F_N (z)  = z^{(M + N - 1 - 2m)/2} 
+   F_N (z)  = z^{(M + N - 1 - 2m)/2} \; .
 \end{equation}
-Moreover, we have that
+As before, we expand \(F_j(z)\) as the geometric sum
 \begin{equation}
-  H(z, y_j) = \frac{1 - (y_j z)^{M + N}}{1 - y_j z} = \sum_{n=0}^{N + M - 1} (y_j z)^n\; , 
+  F_j(z) = \frac{1 - (y_j z)^{M + N}}{1 - y_j z} = \sum_{n=0}^{N + M - 1} (y_j z)^n
+  \equiv \sum_{n=0}^{M + N -1} f_{j, n} z^n\; , 
 \end{equation}
-therefore, we write
+where we will also assume the \(M + N - 1 \in 2 \mathbb{Z}\).
+
+With these definitions, we conclude that
 \begin{equation}
-  F_j(z) =  \sum_{n=0}^{M + N -1} f_{j, n} z^n
+\mathcal{P}_m(N,M|\bm{x}, \bm{y}) =\frac{\det_{jk} F_j(x_k)}{\Delta(x)}\; .
 \end{equation}
-where we will also assume the \(M + N \in 2 \mathbb{Z}\). With these definitions,
-we conclude that
-\begin{equation}
-  \mathbb{P}(m|\bm{x}, \bm{y}) =\frac{\det_{jk} F_j(x_k)}{\Delta(x)}\; ,
-\end{equation}
-and we write~\cite{Takasaki:2010qm}, 
+From the expansion 
 \begin{equation}
@@
 \end{equation}
-With the aid of the Cauchy-Binet formula, we write  
+and with the aid of the Cauchy-Binet formula, we write  
 \begin{equation}
-  \mathbb{P}(m|\bm{x}, \bm{y}) = \sum_{0\leq \ell_N\leq \dots \leq \ell_1 \leq N+M}
-  \frac{\det_{jk}(f_{j, \ell_k}) \det_{jk}(x_k^{\ell_j})}{\Delta(x)}\; ,
+    \mathcal{P}_m(N,M|\bm{x}, \bm{y})
+  = \sum_{0\leq \ell_N\leq \dots \leq \ell_1 \leq N+M}
+  \frac{\det_{jk}(f_{j, \ell_k}) \det_{jk}(x_k^{\ell_j})}{\Delta(x)}\; . 
 \end{equation}
-and from the definition of the Schur polynomials and the
-Jacobi-Trudi expression of Plücker coordinates, we can write this last
-expression as 
+Now we basically use the definition of the Schur polynomials and the
+Jacobi-Trudi expression of Plücker coordinates, see~\cite{Alexandrov:2012tr}. 
+Finally, this last expression becomes
 \begin{equation}
-  \mathbb{P}(m|\bm{x}, \bm{y}) = \sum_{\lambda} c_\lambda(\bm{y}) s_\lambda(\bm{x}) \; ,
+\mathcal{P}_m(N,M|\bm{x}, \bm{y}) =
+\sum_{\lambda} c_\lambda(\bm{y}) s_\lambda(\bm{x}) \; ,
 \end{equation}
-where \(c_\lambda(f_{j, \ell_k})\),  \(\ell_k = \lambda_k - k +N\), define point
-in the finite dimensional Grassmannian manifold \(\mathrm{Gr}(N, N+M)\).
+where \(c_\lambda \equiv \det_{jk}(y_j^{\ell_k})\),  \(\ell_k = \lambda_k - k +N\). Therefore,
+we have that this expression is a KP tau function.
+
+From the coordinate expansion of the Bethe
+vectors~\cite{Bogoliubov2005, Tsilevich:2006}, it is easy to show that
+\begin{equation}
+\prod_{j=1}^N \mathbb{B}(x_j)\ket{\lambda}  = \sum_{\mu \supset \lambda } s_{\mu/\lambda}(\bm{x})\ket{\mu}\qquad 
+ \bra{\lambda} \prod_{j=1}^N \mathbb{C}(x_j) = \sum_{\mu \supset \lambda } s_{\mu/\lambda}(\bm{x})\bra{\mu}
+\end{equation}
+where \(s_{\mu/\lambda}\) are Schur polynomials~\cite{Macdonald:1998}.
+Therefore, we can write the correlation function~(\ref{eq:correlation}) as 
+\begin{equation}
+\begin{split}
+  P_m(N, M|\bm{x}, \bm{y}\setminus \{y_N\})
+  & = \sum_\lambda \sum_{\mu \supset (m)} s_{\lambda}(\bm{x}) s_{\mu/(m)}(\bm{y}\setminus\{y_N\})\\
+  & = \sum_\mu s_{\mu/(m)}(\bm{y}\setminus\{y_N\}) s_{\mu}(\bm{x})\; .
+\end{split}
+\end{equation}
+where we have used that \(\phi^\dagger \ket{\bm{0}} = \ket{(m)}\), and in the second line
+we sum over all partitions, since \(s_{\mu/(m)} = 0\), \(\forall \) \(\mu \not \supset (m)\).
+This expression also shows that the expresison \(P_m\) can also be written as a tau function.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-\subsubsection{Not everything is a tau-function}
-It is also worth noticing that not everything is a tau function. Consider the state
+\subsubsection{Schur functions expansion for other correlation functions}
+More generally, let us consider the correlation functions 
+\begin{equation}
+\begin{split}
+  P_{\lambda_1 \lambda_2}(N', N, M|\bm{x}, \bm{y}) & =
+  \bra{\lambda_1} \prod_{j=1}^N \mathbb{C}(x_j)
+  \prod_{k=1}^{N'} \mathbb{B}(y_k)\ket{\lambda_2} \\
+  & = \sum_{\mu} s_{\mu/\lambda_1}(\bm{x}) s_{\mu/\lambda_2}(\bm{y}) \; .
+\end{split}
+\end{equation}
+where \((N' + \ell(\lambda_2) + n_0 = N' + \ell(\lambda_2) + n'_0)\) and 
+we have also used that the expression for any Young diagram
+\(\mu\) that does not contain \(\lambda_1\) and \(\lambda_2\)
+vanishes. Using the elementary property of skew Schur
+polynomials~\cite{Macdonald:1998}
+\begin{equation}
+  \sum_{\mu} s_{\mu/\lambda_1}(\bm{x}) s_{\mu/\lambda_2}(\bm{y}) = \prod_{i,j}\frac{1}{1 - x_i y_j}
+ \sum_\nu s_{\lambda_1/\nu}(\bm{x}) s_{\lambda_2/\nu}(\bm{y})\; ,
+\end{equation}
+and the Cauchy's identity, we have 
+\begin{equation}
+\begin{split}
+  P_{\lambda_1 \lambda_2}(N', N, M|\bm{x}, \bm{y}) & =
+ \sum_\mu s_{\mu}(\bm{x}) s_{\mu}(\bm{y})
+ \sum_\nu s_{\lambda_1/\nu}(\bm{x}) s_{\lambda_2/\nu}(\bm{y})\; .
+\end{split}
+\end{equation}
+Therefore, this correlation function is nothing but the off-shell
+norm~(\ref{eq:scalar}) times a finite sum over skew Schur functions.
+This expression shows that the existence of these tau functions is not
+a trivial fact of this model.
+
+In fact, it is worth considering other quantities that are not
+necessarily tau functions. Consider the state
 \begin{equation}

### 2024-04-16 draft.tex
finish phase model

@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\subsubsection{Schur functions expansion for other correlation functions}
+\subsubsection{Skew Schur polynomials expansion}
 More generally, let us consider the correlation functions 
@@
 \end{equation}
-where \((N' + \ell(\lambda_2) + n_0 = N' + \ell(\lambda_2) + n'_0)\) and 
+where \(N' + \ell(\lambda_2) + n_0 = N' + \ell(\lambda_2) + n'_0\) and 
 we have also used that the expression for any Young diagram
@@
 
-In fact, it is worth considering other quantities that are not
-necessarily tau functions. Consider the state
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\subsection{More tau functions}
+Based on the discussion we had so far, and on the general map between 
+the Phase model and the free fermions, one can understand the general form 
+of tau functions in this phase model context.
+
+Let us consider the vertex operator construction~\cite{Okounkov2001},
+see also~\cite{Alexandrov:2012tr, Wheeler:2010vmq}. Consider the
+vacuum \(\ket{\bm{0}}\), a ``Dirac sea'', defined by the conditions \(\psi_m\ket{\bm{0}}
+= \psi_n^\star\ket{\bm{0}} =0 \) for \(m<0\) and \(n \geq 0\), where 
+\(\psi_n\) are components of a holomorphic free fermionic field. 
+In this formalism, the partition states are given by 
+\begin{equation}
+  \ket{\mu} = \textrm{sign}(\sigma) \prod_{j=1}^d \psi_{a_j} \psi^\star_{-b_j}\ket{\bm{0}}\; ,
+\end{equation}
+where \(\textrm{sign}(\sigma) = \pm 1\) are defined in a such a way
+that the Shur coefficients of the vertex operators, defined below,
+have positive coefficients.
+
+The pairs \((a_j|b_j)_{j=1}^d\) define the Frobenius notation of the
+partition \(\mu = (\mu_1, \mu_2, \dots, \mu_\ell)\), that is given by
+\(a_j = \mu_j - j\) and \(b_j = \mu^T_j - j\), where \(d\) is the
+number of boxes in the diagonal of the Young diagram, and \(\mu^T\) is
+its conjugate, or transpose.  From these definitions, we have the
+equivalence
+\begin{equation}
+  \prod_{k\geq 1} (\phi_k^\dagger)^{n_k} \mapsto \textrm{sign}(\sigma) \prod_{j=1}^d
+  \psi_{a_j} \psi^\star_{-b_j} \qquad |\mu| = \sum_k k n_k = \sum_j(a_j + b_j) + d\; .
+\end{equation}
+
+Finally, the \(\mathfrak{gl}(\infty)\) has generators given by the
+bilinears \(X = \sum_{j, j \in \mathbb{Z}} x_{ij} \bm{\colon} \psi_i \psi_j^\star\bm{\colon}
++ c\), where \(c\in \mathbb{C}\), \(x_{ij} =
+0\) for large \(|j -i|\), say \(\geq M\), and the colons denote the normal ordering
+\begin{equation}
+  \bm{\colon} \psi_i \psi_j^\ast \bm{\colon} =  \psi_i \psi_j^\ast
+  - \bra{0} \psi_i \psi_j^\ast \ket{0} \; .
+\end{equation}
+The group \(GL(\infty)\) is defined through the \(\exp:
+\mathfrak{gl}(\infty) \to GL(\infty)\) as usual.
+
+Observe that these elements have finitely many non zero elements, and
+that these include diagonal terms \(\bm{\colon} \psi_i
+\psi_i^\star\bm{\colon}\), upper and lower triangular terms. These are
+equivalent to the Number operators \(\mathcal{N}\), annihilation
+\(\phi\) and creation \(\phi^\dagger\) , respectivelly.  The central
+charge is equivalent to the vaccum projection \(\pi =
+\ket{\bm{0}}\bra{\bm{0}}\). Therefore, the elements
+\begin{equation}
+  G \in GL(\infty) \mapsto
+  \mathcal{G} = \exp \left(\sum_{i=1}^M \sum_{a=1}^3  c_{i, a} T_i^{a}  + c \pi_i\right)
+\end{equation}
+\(T^{(a)}_i \in \{ \mathcal{N}_i, \phi_i, \phi_i^\dagger \}_{i=1}^M\).
+
+The operators \(\mathbb{B}(x)\) and \(\mathbb{C}(x)\), 
+for a large enough chain \(M\to \infty\),
+are related to the vertex operators
+\(\Gamma_-(x)\) and \(\Gamma_+(x)\), respectivelly, as
+\begin{equation}
+  \begin{split}
+    \mathbb{B}(x) \mapsto \Gamma_-(x) & = \exp \left( \sum_{n\geq 1} \frac{1}{n}x^n J_{-n}\right) \\
+    \mathbb{C}(x) \mapsto \Gamma_+(x) & = \exp \left( \sum_{n\geq 1} \frac{1}{n}x^n J_{n}\right) \; ,
+  \end{split}
+\end{equation}
+where \(J_n \) is written, in terms of free fermions, as \(J_n =
+\sum_{j\in \mathbb{Z}} \bm{\colon} \psi_j \psi_{j+n}^\ast
+\bm{\colon}\),
+These set of operators generate a Heisenberg subalgebra
+\(\widehat{\mathfrak{gl}}(1) \subset \mathfrak{gl}(\infty)\) given by
+\begin{equation}
+  [J_m, J_n] = m \delta_{n+m,0}\; .
+\end{equation}
+
+Putting all these facts together, we have that the tau functions of the Toda 
+hierarchies, given by
+\begin{equation}
+  \tau_s(\bm{x}, \bm{y}) = \bra{s} \prod_{i} \Gamma_+(x_i) G \prod_j \Gamma_+(x_j) \ket{s}\; ,
+\end{equation}
+are maped into objects of the form 
+\begin{equation}
+\begin{split}
+  \tau_0(\bm{x}, \bm{y}) & = \bra{\Psi(\bm{x})} \mathcal{G} \ket{\Psi(\bm{y})} \\
+  \\ & = \bra{\bm{0}} \prod_{i} \mathbb{C}(x_i)
+  \mathcal{G} \prod_j \mathbb{B}(y_j) \ket{\bm{0}}\; ,
+\end{split}
+\end{equation}
+where we necessarily have \(s=0\) in the phase model
+
+\subsubsection{Schur polynomials expansion of some correlation functions}
+In fact, it is worth noticing that other quantities, that are not tau
+functions, might have interesting Schur polynomials expansion.
+Consider the state calculated in~\cite{Bogoliubov2005}
 \begin{equation}
   \ket{\mathcal{Y}} =
- \sum_{\vec{n}} \prod_{j=0}^M \ket{n_j}\quad \sum_j n_j = N\; .
+ \sum_{\vec{n}} \prod_{j=0}^M \ket{n_j} = \sum_\mu \ket{\mu} \qquad \sum_j n_j = N\; .
 \end{equation}
-We have~\cite{Bogoliubov2005} the correlation function 
+Then
 \begin{equation}
-  \prod_{j=1}^N x^{M/2} \bra{\mathcal{Y}} \prod_{j=1}^N B(\bm{x}) \ket{\bm{0}} = 
-  \sum_{\lambda\subseteq (N,N,M)} s_\lambda(\bm{x})\; . 
+ \bra{\mathcal{Y}} \prod_{j=1}^N B(\bm{x}) \ket{\nu} = 
+  \sum_{\lambda\subseteq (N,N,M)} s_{\lambda/\nu}(\bm{x})\; . 
 \end{equation}
-This sum of Schur polynomials are not tau functions of the KP hierarchy.
-
-We can easily see this fact, if we consider the the Fermionic
-construction. Since
-\begin{equation}
-\tau(\bm{t}) = \bra{\bm{0}} e^{\bm{J}_+(\bm{t})} G \ket{\bm{0}}
-\end{equation}
-where \(G\in GL(\infty)\), and 
-\begin{equation}
-  \bra{\bm{0}} e^{\bm{J}_+(\bm{t})} = \sum_\lambda \bra{\mu} s_\lambda(\bm{t})\; , 
-\end{equation}
-therefore
-\begin{equation}
-\tau(\bm{t}) = \sum_{\lambda} \bra{\lambda} G \ket{\bm{0}} s_\lambda(\bm{t})
-\end{equation}
-and we need \(1 = \bra{\lambda} G \ket{\bm{0}} \). Then, 
-\begin{equation}
-  \ket{\Psi_0} = G\ket{\bm{0}} = \sum_\lambda \ket{\lambda}\; ,
-\end{equation}
-is a coherent state, and we conclude that
-\begin{equation}
-G = 1 +  \sum_{d\geq 1} \prod_{j=1}^d \psi^\star_{a_j} \psi_{-b_j}\ \notin \ GL(\infty) \;. 
-\end{equation}
-It is interesting to notice that this state appears as the ground
-state of the crystal melting problem in~\cite{Dijkgraaf:2008ua}.
-
-Therefore, the restriction for diagrams with \(\ell(\lambda_1) \leq N\) is
-not a tau function either.
+And it is just a sum of skew Schur polynomials.
 

### 2024-04-16 draft.tex
remove old files and improve q-bosons

@@
 
-\subsubsection{Partitions representation}
-Let us define the i\(^{th}\)-vacuum \(\ket{0}_i\) by the the conditions 
-\(\phi_i\ket{0}_i\). The representation of this Hilbert space, denoted by
-\(\mathcal{F}_i\), is given by \(\ket{n_i}_i = \phi_i^\dagger \ket{0}_i\).
+\subsubsection{Representation}
+Let us define the i\(^{th}\)-vacuum \(\ket{0}_i\) by \(\phi_i\ket{0}_i
+= 0\).  The representation of this Hilbert space, denoted by
+\(\mathcal{F}_i\), is given by \(\ket{n_i}_i = \phi_i^\dagger
+\ket{0}_i\).
 
@@
 \phi_i^\dagger \otimes \cdots \otimes \bm{1}\).
-We will assume that these states are normalized, that is
+These states are normalized, that is
 \(\bracket{\vec{n}}{\vec{m}}=\delta_{\vec{n}, \vec{m}}\).
@@
 [\mathcal{N}_i, b_j]=-\delta_{i,j}b_i\; , \quad
-[B_i, B_j^\dagger]= \delta_{i,j} q^{-2\mathcal{N}_i}  \equiv \delta_{i,j} Q^{\mathcal{N}_i}\; , 
+[b_i, b_j^\dagger]= \delta_{i,j} q^{-2\mathcal{N}_i}  \equiv \delta_{i,j} Q^{\mathcal{N}_i}\; , 
 \end{equation}
@@
 We build the representation of the q-boson algebra with some modifications of the phase 
-model we have just discussed. We suppose that the action of the operators \(\{b_i, b_i^\dagger\}\) 
-satisfies 
+model we have just discussed. Let us define the i\(^{th}\)-vacuum \(\ket{0}_i\) by
+\(b_i\ket{0}_i = 0\). The representation of this Hilbert space, denoted by
+\(\mathcal{F}_i^Q\), is given by \(\ket{n_i}_i \propto b_i^\dagger \ket{0}_i\).
+
+Given the vacuum \(\ket{\bm{0}} = \ket{0}_0\otimes \ket{0}_1
+\otimes \cdots \otimes  \ket{0}_M\),
+the Fock space is defined as 
+\begin{equation}
+  \mathcal{F}_{Q} = \bigotimes_{i=0}^M \mathcal{F}_i^Q = 
+  \left\{\ket{\vec{n}} = \ket{n_0}\otimes \ket{n_1} \otimes \cdots
+  \otimes \ket{n_M} \ | \ n_i \in \mathbb{N} \right\}\; .
+\end{equation}
+where the actions of the operators \(\{b_i, b_i^\dagger\}\) are given by 
+the following relations
 \begin{subequations}
 \begin{alignat}{3}
-    & b_0\ket{n_0} =(1 - \delta_{0, n_0}) \frac{1 - Q^{n_0}}{(1 - Q)^{1/2}}\ket{n_0 - 1}
+    & b_0\ket{n_0} =(1 - \delta_{0, n_0}) \frac{(1 - Q^{n_0})}{(1 - Q)^{1/2}}\ket{n_0 - 1}
     &\hspace{0.5cm}& b_0^\dagger \ket{n_0} =  \frac{1}{(1 - Q)^{1/2}}\ket{n_0 + 1}  \\
     & b_i\ket{n_i} = \frac{(1 - \delta_{0, n_i})}{(1 - Q)^{1/2}}\ket{n_i - 1}
-    && b_i^\dagger \ket{n_0} =  \frac{(1 - Q^{n_i+1})}{(1 - Q)^{1/2}}\ket{n_0 + 1}\quad i\neq 0
+    && b_i^\dagger \ket{n_0} =  \frac{(1 - Q^{n_i+1})}{(1 - Q)^{1/2}}\ket{n_0 + 1}\quad i\neq 0\; .
 \end{alignat}
 \end{subequations}
-where the modes \(\{b_0, b_0^\dagger \}\) satisfies a slightly different relation.
+Observe that the modes \(\{b_0, b_0^\dagger \}\) satisfies a slightly
+different relation.
+
+Therefore, the states \(\ket{\vec{n}} \in \mathcal{F}_Q\) are defined as 
+\begin{equation}
+  \ket{\vec{n}} = \ket{n_0}\otimes \ket{n_1} \otimes \cdots \otimes \ket{n_M} 
+ : =  (b_0^\dagger)^{n_0} (b_1^\dagger)^{n_1} \cdots  (b_M^\dagger)^{n_M} \ket{\bm{0}} \; ,
+\end{equation}
+where we write \(\phi_i^\dagger \equiv \bm{1} \otimes  \cdots \otimes
+\phi_i^\dagger \otimes \cdots \otimes \bm{1}\).
+Moreover, the norm of these states satisfy 
+\begin{equation}
+  \bracket{\vec{n}}{\vec{m}} = \frac{[m_0]!}{\prod_{i=1}^M[m_i]!} \delta_{\vec{n}\vec{m}}\qquad 
+    [n]! =
+    \left\{
+    \begin{array}{ll}
+    \prod_{i=1}^n(1 - Q^i) & \textrm{if}\ \ n\neq 0 \\
+    1 & \textrm{otherwise}
+  \end{array}\right.\; .
+\end{equation}
+As before, we can associate to a given states \(\ket{\vec{n}} =
+\ket{n_0, n_1, \dots, n_M}\), a partition -- or Young diagram --
+\(\lambda = (1^{n_1} 2^{n_2} \cdots M^{n_M})\), but now we assume that 
+there is a proportionaly  factor relating these two objects, 
+\begin{equation}
+  \ket{\lambda}  = b_\lambda(Q) \ket{\vec{n}}  \ \; , \qquad b_\lambda(Q) = \prod_i [p_i(\lambda)]! 
+\end{equation}
+where \(p_i(\lambda)\) denotes the number of parts of size \(i\) in
+the partition. These partition states satisfy \(
+\bracket{\lambda}{\mu} = b_\lambda(Q) \delta_{\lambda,\mu}\). Once
+again, we see that the correspondence is not unique since the number
+of oscilators at the site \(i=\) are completely ignored in the
+partition notation.  An finally, if the number of particles is fixed,
+then \(n_0 = N - \ell(\lambda)\).
 
@@
 \begin{equation}
-  L_n(x, Q) =
+  L_{an}(x, Q) =
   \begin{pmatrix}
-    x^{-1/2} & (1 - Q)^{1/2} b_n^\dagger \\ (1 - Q)^{1/2} b_n & x^{1/2}
-  \end{pmatrix}\; .
+    x^{-1/2} & (1 - Q)^{\frac{1}{2}} b_n^\dagger \\ (1 - Q)^{\frac{1}{2}} b_n & x^{1/2}
+  \end{pmatrix}_a\; .
 \end{equation}
@@
 \begin{equation}
-  T_a(x,Q) = L_{am}(x, q)  \dots  L_{a0}(x, q) = 
+  T_a(x,Q) = L_{am}(x, Q)  \dots  L_{a0}(x, Q) = 
   \begin{pmatrix}
-    A(x) & B(x) \\ C(x) & D(x)
-  \end{pmatrix}\; .
+    A(x, Q) & B(x, Q) \\ C(x, Q) & D(x, Q)
+  \end{pmatrix}_a\; .
 \end{equation}
@@
 \begin{equation}
-  \ket{\Psi(y_1, \dots, y_N; Q)} = \prod_{j=1}^N B(y_j, Q) \ket{\bm{0}}\qquad 
-  \bra{\Psi(y_1, \dots, y_N; Q)} = \bra{\bm{0}}\prod_{j=1}^N C(y_j, Q) \; .
+  \ket{\Psi(y_1, \dots, y_N; Q)} = \prod_{j=1}^N \mathbb{B}(y_j, Q) \ket{\bm{0}}\qquad 
+  \bra{\Psi(y_1, \dots, y_N; Q)} = \bra{\bm{0}}\prod_{j=1}^N \mathbb{C}(y_j, Q) \; .
 \end{equation}
-When the set \(\{ y_j \ | \ j =1, \dots , N\}\) satisfies the Bethe equations 
+where \(\mathbb{B}(x, Q) = x^{M/2} B(x, Q)\) and \(\mathbb{C}(x, Q) =
+x^{M/2} C(1/x, Q)\).  When the set of parameters \(\{ y_j \ | \ j =1, \dots , N\}\)
+satisfies the Bethe equations
 \begin{equation}
@@
 \end{equation}
-Once again, we will consider \emph{off-shell} Bethe states.
+we say that the states \(\ket{\Psi(y_1, \dots, y_N; Q)}\) are \emph{on-shell} Bethe states, otherwise, 
+we have \emph{off-shell} Bethe states.
 
@@
 
-This section discusses presence of integrable hierarchies in the phase
+This section discusses the presence of integrable hierarchies in the phase
 model.  We argue that it is also a Toda hierachy tau function, and we
@@
 we sum over all partitions, since \(s_{\mu/(m)} = 0\), \(\forall \) \(\mu \not \supset (m)\).
-This expression also shows that the expresison \(P_m\) can also be written as a tau function.
+This expression also shows that the expression \(P_m\) can also be written as a tau function.
 
@@
 \begin{equation}
-  G \in GL(\infty) \mapsto
-  \mathcal{G} = \exp \left(\sum_{i=1}^M \sum_{a=1}^3  c_{i, a} T_i^{a}  + c \pi_i\right)
+   GL(\infty) \ni G \mapsto
+   \mathcal{G} = \exp \left[\sum_{i=1}^M \left( \sum_{a=1}^3
+   c_{i, a} T_i^{a}  + c \pi_i \right)\right]\; ,
 \end{equation}
-\(T^{(a)}_i \in \{ \mathcal{N}_i, \phi_i, \phi_i^\dagger \}_{i=1}^M\).
+where \(T^{a}_i \in \{ \mathcal{N}_i, \phi_i, \phi_i^\dagger \}_{i=1}^M\).
 
@@
 \end{equation}
-are maped into objects of the form 
+are mapped into objects of the form 
 \begin{equation}
@@
 
-\section{Scalar products for the Q-bosons}
+\section{Tau functions in the Q-bosons model}
 
-Here I would like to start considering the q-boson problem. It has
-been shown~\cite{Wheeler:2010vmq, Sulkowski:2008mx, Tsilevich:2006}
-that the scalar product of two off-shell Bethe states in the q-boson
-model is given by
+We now move out attention to the case of the q-bosons. The 
+analysis is parallel to what we did before, but the  specific 
+details are very different. We start with the study of the norm of  
+two off-shell Bethe states.
+
+It has been shown~\cite{Tsilevich:2006} conf. \cite{Sulkowski:2008mx,
+  Wheeler:2010vmq}, that the Bethe states \(\ket{\Psi(\bm{x})}\) have 
+coordiante expansion given by
+\begin{equation}
+  \ket{\psi(\bm{x})} = \sum_{\mu \subset [N,M]} P_\mu(\bm{x}, Q) \ket{\mu}  \qquad 
+  \bra{\psi(\bm{x})} = \sum_{\mu \subset [N,M]} P_\mu(\bm{x}, Q) \bra{\mu}\; ,
+\end{equation}
+where \(P_\lambda\) denote the Hall-Littlewood polynomials.  In this
+form, scalar product of two off-shell Bethe states in the q-boson
+model is easily calculated to be
 \begin{equation}
@@
 \end{equation}
-where \(P_\lambda\) denote the Hall-Littlewood polynomials.
-
-It turns out that this expansion is  also related to the Cauchy identity for 
-Hall-Littlewood polynomials~\cite{Macdonald:1998}
+where we use that \(\bracket{\lambda}{\mu} = b_\lambda(Q)
+\delta_{\lambda, \mu}\).  It turns out that this expansion is also
+related to the Cauchy identity for Hall-Littlewood
+polynomials~\cite{Macdonald:1998}
 \begin{equation}
+\label{eq:cauchy_hl}
 \sum_{\lambda} b_\lambda(Q) P_{\lambda}(\bm{x}, Q) P_{\lambda}(\bm{y}, Q)
@@
 \end{equation}
-and using the results in~\cite[see chap.3, sec. 4, eq. (4.7)]{Macdonald:1998}, 
-we can expand this identity as 
+
+Ourt goal is to consider some elementary properties of Hall-Littlewood polynomials 
+to understand aspects of this expansion.
+
+\subsection{Scalar product: Big Schur functions}
+
+Let us now consider the results in~\cite[see chap.3, sec. 4, eq. (4.7)]{Macdonald:1998}.
+We can expand the Cauchy identity~(\ref{eq:cauchy_hl}) as
 \begin{equation}
@@
 
-\subsection{Supersymmetric Schur polynomials}
+\subsection{Scalar product: Supersymmetric Schur polynomials}
 
@@
 
-%\input{./sections/fermions.tex}
-%
-%\input{./sections/schurpol.tex}
-%
-%\input{./sections/transfer.tex}
-%
-%\input{./sections/t-fermions.tex}
-%
-%\input{./sections/top-strings.tex}
-
-%\addcontentsline{toc}{section}{References}
-
 \printbibliography

### 2024-04-17 draft.tex
Add determinant formulas for the q-bosons scalar product

@@
 partition \(\mu = (\mu_1, \mu_2, \dots, \mu_\ell)\), that is given by
-\(a_j = \mu_j - j\) and \(b_j = \mu^T_j - j\), where \(d\) is the
-number of boxes in the diagonal of the Young diagram, and \(\mu^T\) is
+\(a_j = \mu_j - j\) and \(b_j = \mu'_j - j\), where \(d\) is the
+number of boxes in the diagonal of the Young diagram, and \(\mu'\) is
 its conjugate, or transpose.  From these definitions, we have the
@@
 \begin{equation}
-  \tau_s(\bm{x}, \bm{y}) = \bra{s} \prod_{i} \Gamma_+(x_i) G \prod_j \Gamma_+(x_j) \ket{s}\; ,
+  \tau_s(\bm{x}, \bm{y}) = \bra{s} \prod_{i} \Gamma_+(x_i) G \prod_j \Gamma_-(x_j) \ket{s}\; ,
 \end{equation}
@@
 
+%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*
 \section{Tau functions in the Q-bosons model}
+%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*
 
@@
 \begin{equation}
+\label{eq:scalar-hl}
 S_Q(N,M | \bm{x}, \bm{y}) = \bra{\bm{0}} \prod_{j=1}^N C(x_j, Q)
@@
 
-Let us now consider the results in~\cite[see chap.3, sec. 4, eq. (4.7)]{Macdonald:1998}.
-We can expand the Cauchy identity~(\ref{eq:cauchy_hl}) as
-\begin{equation}
-\label{eq:cachy-hl}
+Let us first consider the result originally derived
+in~\cite{Foda:2008hn} that shows how the scalar
+product~(\ref{eq:scalar-hl}) is a tau function of the KP hierarchy.
+From~\cite[see chap.3, sec. 4, eq. (4.7)]{Macdonald:1998}, we can
+expand the Cauchy identity~(\ref{eq:cauchy_hl}) as
+\begin{equation}
  \prod_{j, k=1}^\infty \frac{1-Q x_j y_k}{1 - x_j y_k} = 
@@
 \end{equation}
-where \(z\) is a formal variable. It has been argued
-in~\cite{Foda:2008hn} that if we interpret the \(Q\)-Schur functions
-as Plücker coordiantes, the identity~(\ref{eq:cachy-hl}) is a KP-tau
-function.
-
-Therefore, we see that the scalar product 
+where \(z\) is a formal variable. It has been argued in that if we
+interpret the \(Q\)-Schur functions as Plücker coordiantes, the
+expression
 \begin{equation}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
-Let us also decompose 
+We now aargue that~(\ref{eq:scalar-hl}) is also a tau function of the
+Toda hierarchy in disguise. Let us decompose 
 \begin{equation}
@@
 are homogeneous polynomials with respect the Miwa coordinates
-\(\bm{T} = (T_1, T_2, \dots)\). All in all, we conclude that 
+\(\bm{T} = (T_1, T_2, \dots)\). Then
 \begin{equation}
@@
 \end{equation}
-
-Supersymmetric Schur functions, see~\cite{Zinnjustin2009}, \(s_\lambda(\bm{\alpha}/\bm{\beta})\)
-are defined as ordinary schur functions evaluated at Miwa coodinates
-of the form
+All in all, we conclude that the big Schur functions
+\(S_\lambda(\bm{t}', Q)\) are ordinary Schur functions with respect
+the coordinates \(\bm{T}\).
+
+In fact, we can do even better. Supersymmetric (or Hook) Schur
+functions~\cite{Berele:1983}, see also see~\cite{Macdonald:1998, Moens:2003},
+\(s_\lambda(\bm{\alpha}/\bm{\beta})\) are defined as ordinary schur
+functions evaluated at Miwa coodinates of the form
 \begin{equation}
@@
 \end{equation}
-is also a restricted tau function of the Toda hierarchy.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\subsection{Toda hierarchy}
+is also a restricted tau function of the Toda hierarchy with respect
+to \(\bm{t}\) and \(\bm{T}\).
+
+There is something very interesting about this expression. Let us
+expand the supersymmetric Schur polynomials in terms of ordinary Schur
+functions as~\cite[Sec. I.5, exerc. 23]{Macdonald:1998}
+\begin{equation}
+  s_{\lambda}(\bm{x}/\bm{y}) = \sum_\mu s_{(\lambda/\mu)'}(\bm{y}) s_\mu(\bm{x})\; ,
+\end{equation}
+where the prime denotes the conjugate diagram. Then
+\begin{equation}
+  \label{eq:tau-hl}
+  \sum_{\lambda } s_\lambda(\bm{y}/ (- Q\bm{y})) s_\lambda(\bm{x}) =
+  \sum_{\lambda,\mu } s_{(\lambda/\mu)'}(- Q\bm{y}) s_\lambda(\bm{x}) s_\mu(\bm{y})\; .
+\end{equation}
+From this expression one can reach to some conclusions, and make some 
+conjectures regarding this expansion.
+
+\begin{remark}
+Observe that the skew Schur polynomials have
+determinant expressions, therefore we have that the
+\(\bm{y}\)-dependent coefficients \(c_{\mu \lambda} = s_{(\lambda/\mu)'}(-Q\bm{y})\)
+ have Jacobi-Trudi expressions. We may think of these objects as coordinate dependent 
+Pl\"ucker coordinates. 
+\end{remark}
+
+We also know that the left-hand side of~(\ref{eq:tau-hl}) is a tau
+function with respect the coodinates \(\bm{T}\equiv \{\bm{y}\cup(- Q
+\bm{y})\}\) and \(\bm{t}\). Moreover, from the Vertex operator
+formalism, we write the Schur polynomials as
+\begin{equation}
+  s_{\lambda}(\bm{x}) = \bra{\lambda} e^{J_-}\ket{\bm{0}}\qquad
+  s_{\mu}(\bm{y}) = \bra{\bm{0}} e^{J_+}\ket{\mu}\; ,
+\end{equation}
+we have
+\begin{equation}
+  \sum_{\lambda,\mu } s_\mu(\bm{y})c_{\mu\lambda}(Q,\bm{y}) s_\lambda(\bm{x}) = 
+  \sum_{\lambda,\mu } \bra{\bm{0}} e^{J_+}\ket{\mu} c_{\mu\lambda}(Q,\bm{y})
+  \bra{\lambda} e^{J_-}\ket{\bm{0}} \; ,
+\end{equation}
+therefore 
+\begin{equation}
+  c_{\mu\lambda}(Q, \bm{y}) = \bra{\mu} G_Q(\bm{y}) \ket{\lambda}\; ,
+\end{equation}
+where \(G_Q(\bm{y}) \in GL(\infty)\). We can find the expression for
+this group element as follows.
+
+Remember that \(T_m = t'_m - Q^m t'_m\), therefore we write
+\begin{subequations}
+\begin{equation}
+  \bra{\bm{0}} e^{J_+(\bm{t})} e^{J_-(\bm{T})} \ket{\bm{0}} = 
+\bra{\bm{0}} e^{J_+(\bm{t})} e^{J_-(\bm{t}')  - J_-(\bm{t}^{(Q)})} \ket{\bm{0}} 
+\end{equation}
+From the Heisenberg algebra, we conclude that \([J_-(\bm{t}),
+  J_-(\bm{t}')] = 0\), therefore
+\begin{equation}
+  \bra{\bm{0}} e^{J_+(\bm{t})} e^{J_-(\bm{T})} \ket{\bm{0}} = 
+\bra{\bm{0}} e^{J_+(\bm{t})} e^{-J_-(\bm{t}^{(Q)})} e^{J_-(\bm{t}')}  \ket{\bm{0}} 
+\end{equation}
+\end{subequations}
+from where we finally conclude that 
+\begin{equation}
+  G_{Q}(\bm{y}) = \exp \left(-J_-(\bm{t}^{(Q)})\right) \qquad
+  t^{(Q)}_n = \frac{Q^n}{n}\sum_j y_j^n\; . 
+\end{equation}
+
+In general, we do not expect that expectation values with coordinate dependent 
+group elements \(G_{Q}(\bm{y})\) are tau functions. But in the above example,
+the coordinates combine among themselves and generate a tau function with respect 
+the coordiantes \(\bm{T}\) and \(\bm{t}\).
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\subsection{Determinant formula}
+
+We would like to refine the expressions above. Let us now consider a determinant  
+expression for the scalar product~(\ref{eq:scalar-hl}). Let us continue with 
+the case where \(M\) and \(N\) are very large, but finite. In this case, we write  
+this scalar product as 
+\begin{equation}
+  S_Q(N,M | \bm{x}, \bm{y})  
+= \prod_{j_1, j_2} (1-Q x_{j_1} y_{j_2}) \prod_{k_2, k_2}(1 - x_{k_1} y_{k_2})^{-1}\; .
+\end{equation}
+
+Using the Cauchy's identity for Schur polynomials, that is 
+\begin{equation}
+  \sum_\lambda s_\lambda(\bm{x}) s_\lambda(\bm{y}) = \prod_{i,j} \frac{1}{1 - x_i y_j}\; ,
+\end{equation}
+and using the results we have derives in the Phase model, we have that 
+\begin{equation}
+  \prod_{i,j}\frac{1}{1 - x_i y_j}  \equiv S(N,M|\bm{x}, \bm{y}) = 
+  \frac{\det_{j,k}H(\bm{x},\bm{y})}{\Delta(\bm{x}) \Delta(\bm{y})}\; .
+\end{equation}
+where \(H\) is the matrix~(\ref{eq:h-matrix}). Therefore
+\begin{equation}
+  S_Q(N,M | \bm{x}, \bm{y})  
+= \frac{S(N,M|\bm{x}, \bm{y})}{S(N,M|\bm{x}, Q\bm{y})}\; , 
+\end{equation}
+and we see that the scalar product of the q-boson model is the quotient of 
+scalar products of the phase model. Moreover
+\begin{equation}
+  S_Q(N,M | \bm{x}, \bm{y})  
+  = \frac{\Delta(Q\bm{y})}{\Delta(\bm{y})}
+  \frac{\det H(\bm{x}, \bm{y})}{\det H(\bm{x},Q \bm{y})} \; .
+\end{equation}
+Using that \(\Delta(Q\bm{y})/ \Delta(\bm{y}) = \prod_{j=1}^{N-1} Q^j\), we have 
+\begin{equation}
+  S_Q(N,M | \bm{x}, \bm{y}) =  Q^{N(N-1)/2}\;
+  \frac{\det H(\bm{x}, \bm{y})}{\det H(\bm{x}, Q \bm{y})} \; .
+\end{equation}
+
+Let us denote \(H(\bm{x}, Q\bm{y}) = H_{Q}\), therefore
+\begin{equation}
+  S_Q(N,M | \bm{x}, \bm{y}) = Q^{N(N-1)/2} \det H_{Q}^{-1} \det H
+= Q^{N(N-1)/2} \det \mathcal{H}(\bm{x}, \bm{y}) \; ,
+\end{equation}
+where \(\mathcal{H} = H_{Q}^{-1} H\) is a matrix with components 
+\(\mathcal{H}_{i,j} \equiv \mathcal{H}(x_i, y_j)\) where
+\begin{equation}
+  \mathcal{H}(z, w) = \sum_{k} \frac{(1 - z y_k)}{(1 - Q z y_k)} 
+\frac{(1 - Q x_k w)}{(1 - x_k w)}\; .
+\end{equation}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\subsection{Kostka-Foulkes expansion}
 
 One of the greatest advantages of the expressions above is that we do not 
-need to consider the t-Schur fermions.
+need to consider the t-fermions.
 

### 2024-04-19 draft.tex
finish bulk of the draft

@@
 \tableofcontents
+
+%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*
+\section{Introduction}
+%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*
 
@@
 \begin{equation}
-  \ket{\lambda}  = b_\lambda(Q) \ket{\vec{n}}  \ \; , \qquad b_\lambda(Q) = \prod_i [p_i(\lambda)]! 
+  \ket{\lambda}_Q  = b_\lambda(Q) \ket{\vec{n}}  \ \; , \qquad b_\lambda(Q) = \prod_i [p_i(\lambda)]! 
 \end{equation}
@@
 the partition. These partition states satisfy \(
-\bracket{\lambda}{\mu} = b_\lambda(Q) \delta_{\lambda,\mu}\). Once
+\bracket{\lambda}{\mu}_Q = b_\lambda(Q) \delta_{\lambda,\mu}\). Once
 again, we see that the correspondence is not unique since the number
-of oscilators at the site \(i=\) are completely ignored in the
-partition notation.  An finally, if the number of particles is fixed,
+of oscilators at the site \(i=0\) are completely ignored in the
+partition notation. And finally, if the number of particles is fixed,
 then \(n_0 = N - \ell(\lambda)\).
@@
 \label{eq:scalar}
-  S(N, M|\bm{x}, \bm{y}) & =
+  \mathcal{I}(N, M|\bm{x}, \bm{y}) & =
   \bra{\bm{0}} \prod_{j=1}^N \mathbb{C}(x_j) \prod_{j=1}^N \mathbb{B}(y_j) \ket{\bm{0}} \\ 
@@
 \begin{equation}
+\label{eq:h-exp}
   H(z,w) = \frac{1 - (zw)^{M+N}}{1 - zw} = \sum_{k=1}^{M+N} (zw)^{k-1} \; , 
@@
 \label{eq:scalar_exp}
-S(N, M|\bm{x}, \bm{y}) = \sum_{\lambda\subseteq [N,M]}
+\mathcal{I}(N, M|\bm{x}, \bm{y}) = \sum_{\lambda\subseteq [N,M]}
 \frac{\det_{ik}(x_i^{\lambda_k - k + N})}{\Delta(\bm{x})} \frac{ \det_{ik}(y_j^{\lambda_k - k + N})}{\Delta(\bm{y})}
@@
 \begin{equation}
-  S(N, M|\bm{t}, \bm{t}') = \sum_{\lambda\subseteq [N,M]} s_\lambda(\bm{t}) s_\lambda(\bm{t}') \; .
+  \mathcal{I}(N, M|\bm{t}, \bm{t}') = \sum_{\lambda\subseteq [N,M]} s_\lambda(\bm{t}) s_\lambda(\bm{t}') \; .
 \end{equation}
@@
 \begin{equation}
-S(N, M|\bm{t}, \bm{t}') = \sum_{\lambda} h_\lambda s_\lambda(\bm{t}) s_\lambda(\bm{t}') \; .
+\mathcal{I}(N, M|\bm{t}, \bm{t}') = \sum_{\lambda} h_\lambda s_\lambda(\bm{t}) s_\lambda(\bm{t}') \; .
 \end{equation}
@@
 \begin{equation}
-  S_N(\bm{t}, \bm{t}') = \sum_{\substack{\lambda \\ \ell(\lambda) \leq N}}
+  \mathcal{I}_N(\bm{t}, \bm{t}') = \sum_{\substack{\lambda \\ \ell(\lambda) \leq N}}
   s_\lambda(\bm{t})  s_\lambda(\bm{t}') \; .
@@
 \begin{equation}
-  S_N(\bm{t}, \bm{t}') =
+  \mathcal{I}_N(\bm{t}, \bm{t}') =
   \frac{1}{N!} \prod_{\ell=1}^N \oint_{\Gamma_\ell} \frac{dz_\ell}{2 \pi i z_\ell}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \subsubsection{KP tau function}
@@
 \begin{equation}
- S(N,M|\bm{t}) 
+ \mathcal{I}(N,M|\bm{t}) 
  = \sum_{\lambda} s_\lambda(\bm{y})  s_\lambda(\bm{t}) 
@@
 \begin{equation}
- S(N,M|\bm{t}) = \sum_{\lambda \subseteq [N,M]} s_\lambda(\bm{y})  s_\lambda(\bm{t}) \; ,
+ \mathcal{I}(N,M|\bm{t}) = \sum_{\lambda \subseteq [N,M]} s_\lambda(\bm{y})  s_\lambda(\bm{t}) \; ,
 \end{equation}
@@
 \label{eq:correlation}
-  P_m(N, M|\bm{x}, \bm{y}\setminus \{y_N\})
+  A_m(N, M|\bm{x}, \bm{y}\setminus \{y_N\})
   & = \bra{0} \prod_{j=1}^N \mathbb{C}(x_j)
@@
 \begin{equation}
-  P_m(N, M|\bm{x}, \bm{y}\setminus \{y_N\}) = 
+  A_m(N, M|\bm{x}, \bm{y}\setminus \{y_N\}) = 
   \frac{(-1)^{N-1}}{y_N^{(N-1)/2}} \prod_{j=1}^N x_{j}^{M/2}
@@
 \begin{equation}
-   S(N,M|\bm{x}, \bm{y}) =  \Delta(x) \Delta(y) \det H \; ,
+    \frac{\mathcal{I}(N,M|\bm{x}, \bm{y})}{\det H}=  \frac{1}{\Delta(x)\Delta(y)} \; ,
 \end{equation}
@@
 \begin{equation}
-  P_m(N, M|\bm{x}, \bm{y}\setminus{y_N})  \propto
-  \frac{\det Q}{\Delta(x)} \equiv \mathcal{P}_m(N,M|\bm{x}, \bm{y})\; ,
+  A_m(N, M|\bm{x}, \bm{y}\setminus{y_N})  \propto
+  \frac{\det Q}{\Delta(x)} \equiv \mathcal{A}_m(N,M|\bm{x}, \bm{y})\; ,
 \end{equation}
@@
 \begin{equation}
-\mathcal{P}_m(N,M|\bm{x}, \bm{y}) =\frac{\det_{jk} F_j(x_k)}{\Delta(x)}\; .
+\mathcal{A}_m(N,M|\bm{x}, \bm{y}) =\frac{\det_{jk} F_j(x_k)}{\Delta(x)}\; .
 \end{equation}
@@
 \begin{equation}
-    \mathcal{P}_m(N,M|\bm{x}, \bm{y})
+    \mathcal{A}_m(N,M|\bm{x}, \bm{y})
   = \sum_{0\leq \ell_N\leq \dots \leq \ell_1 \leq N+M}
@@
 \begin{equation}
-\mathcal{P}_m(N,M|\bm{x}, \bm{y}) =
+\mathcal{A}_m(N,M|\bm{x}, \bm{y}) =
 \sum_{\lambda} c_\lambda(\bm{y}) s_\lambda(\bm{x}) \; ,
@@
 \begin{split}
-  P_m(N, M|\bm{x}, \bm{y}\setminus \{y_N\})
+  A_m(N, M|\bm{x}, \bm{y}\setminus \{y_N\})
   & = \sum_\lambda \sum_{\mu \supset (m)} s_{\lambda}(\bm{x}) s_{\mu/(m)}(\bm{y}\setminus\{y_N\})\\
@@
 \begin{split}
-  P_{\lambda_1 \lambda_2}(N', N, M|\bm{x}, \bm{y}) & =
+  A_{\lambda_1 \lambda_2}(N', N, M|\bm{x}, \bm{y}) & =
   \bra{\lambda_1} \prod_{j=1}^N \mathbb{C}(x_j)
@@
 \begin{split}
-  P_{\lambda_1 \lambda_2}(N', N, M|\bm{x}, \bm{y}) & =
+  A_{\lambda_1 \lambda_2}(N', N, M|\bm{x}, \bm{y}) & =
  \sum_\mu s_{\mu}(\bm{x}) s_{\mu}(\bm{y})
@@
 \label{eq:scalar-hl}
-S_Q(N,M | \bm{x}, \bm{y}) = \bra{\bm{0}} \prod_{j=1}^N C(x_j, Q)
+\mathcal{I}_Q(N,M | \bm{x}, \bm{y}) = \bra{\bm{0}} \prod_{j=1}^N C(x_j, Q)
 \prod_{k=1}^N B(y_k, Q) \ket{\bm{0}}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\subsection{Determinant formula}
+
+We would like to refine the expressions above.  Let us first consider
+a determinant expression for the scalar product~(\ref{eq:scalar-hl}).
+Let us continue with the case where \(M\) and \(N\) are very large,
+but finite. In this case, we write this scalar product as
+\begin{equation}
+  \mathcal{I}_Q(N,M | \bm{x}, \bm{y})  
+= \prod_{j_1, j_2} (1-Q x_{j_1} y_{j_2}) \prod_{k_2, k_2}(1 - x_{k_1} y_{k_2})^{-1}\; .
+\end{equation}
+
+Using the Cauchy's identity for Schur polynomials, that is 
+\begin{equation}
+  \sum_\lambda s_\lambda(\bm{x}) s_\lambda(\bm{y}) = \prod_{i,j} \frac{1}{1 - x_i y_j}\; ,
+\end{equation}
+and using the results we have derives in the Phase model, we have that 
+\begin{equation}
+  \prod_{i,j}\frac{1}{1 - x_i y_j}  \equiv \mathcal{I}(N,M|\bm{x}, \bm{y}) = 
+  \frac{\det_{j,k}H(\bm{x},\bm{y})}{\Delta(\bm{x}) \Delta(\bm{y})}\; .
+\end{equation}
+where \(H\) is the matrix~(\ref{eq:h-matrix}). Therefore
+\begin{equation}
+  \mathcal{I}_Q(N,M | \bm{x}, \bm{y})  
+= \frac{\mathcal{I}(N,M|\bm{x}, \bm{y})}{\mathcal{I}(N,M|\bm{x}, Q\bm{y})}\; , 
+\end{equation}
+and we see that the scalar product of the q-boson model is the quotient of 
+scalar products of the phase model. Observe that in the case \(Q = 0\), we have 
+\(\mathcal{I}(N, M| \bm{x}, \bm{0}) = 1\), therefore
+\(\mathcal{I}_0(N,M | \bm{x}, \bm{y}) = \mathcal{I}(N,M | \bm{x}, \bm{y})\), 
+as expected.
+
+Moreover
+\begin{equation}
+  \mathcal{I}_Q(N,M | \bm{x}, \bm{y})  
+  = \frac{\Delta(Q\bm{y})}{\Delta(\bm{y})}
+  \frac{\det H(\bm{x}, \bm{y})}{\det H(\bm{x},Q \bm{y})} \; .
+\end{equation}
+Using that \(\Delta(Q\bm{y})/ \Delta(\bm{y}) = \prod_{j=1}^{N-1} Q^j\), we have 
+\begin{equation}
+  \mathcal{I}_Q(N,M | \bm{x}, \bm{y}) =  Q^{N(N-1)/2}\;
+  \frac{\det H(\bm{x}, \bm{y})}{\det H(\bm{x}, Q \bm{y})} \; .
+\end{equation}
+Let us denote \(H(\bm{x}, Q\bm{y}) = H_{Q}\), therefore
+\begin{equation}
+  \mathcal{I}_Q(N,M | \bm{x}, \bm{y}) = Q^{N(N-1)/2} \det H_{Q}^{-1} \det H
+= Q^{N(N-1)/2} \det \mathcal{H}(\bm{x}, \bm{y}) \; ,
+\end{equation}
+where \(\mathcal{H} = H_{Q}^{-1} H\) is a matrix with components 
+\(\mathcal{H}_{i,j} \equiv \mathcal{H}(x_i, y_j)\) where
+\begin{equation}
+  \mathcal{H}(z, w) = \sum_{k} \frac{(1 - z y_k)}{(1 - Q z y_k)} 
+\frac{(1 - Q x_k w)}{(1 - x_k w)}\; .
+\end{equation}
+Observe that from this expression, we cannot decompose this function
+as in~(\ref{eq:h-exp}) since the coefficients in this expansion are 
+are dependent on \(\bm{x}\) and \(\bm{y}\).
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \subsection{Scalar product: Big Schur functions}
 
-Let us first consider the result originally derived
+Let us now recover the result originally derived
 in~\cite{Foda:2008hn} that shows how the scalar
@@
  \prod_{j, k=1}^\infty \frac{1-Q x_j y_k}{1 - x_j y_k} = 
-\sum_{\lambda} S_{\lambda}(\bm{x}, Q) s_{\lambda}(\bm{y}) =
-\sum_{\lambda} S_{\lambda}(\bm{y}, Q) s_{\lambda}(\bm{x}) \; ,
-\end{equation}
-where the polynomials, the big-Schur~\cite{Wheeler:2018}, are given by 
-a Jacobi-Trudi formula
-\begin{equation}
-  S_{\lambda} (\bm{y}, Q) = \det(q_{\lambda_i -i + j}(\bm{y}, Q))\; , 
+\sum_{\lambda} \mathcal{S}_{\lambda}(\bm{x}, Q) s_{\lambda}(\bm{y}) =
+\sum_{\lambda} \mathcal{S}_{\lambda}(\bm{y}, Q) s_{\lambda}(\bm{x}) \; ,
+\end{equation}
+where the polynomials \(\mathcal{S}_\lambda\), the big-Schur
+functions, are given by a Jacobi-Trudi formula
+\begin{equation}
+  \mathcal{S}_{\lambda} (\bm{y}, Q) = \det(q_{\lambda_i -i + j}(\bm{y}, Q))\; , 
 \end{equation}
@@
 \end{equation}
-where \(z\) is a formal variable. It has been argued in that if we
-interpret the \(Q\)-Schur functions as Plücker coordiantes, the
-expression
-\begin{equation}
-  S_Q(N,M | \bm{x}, \bm{y})
-  = \sum_{\lambda \subseteq (N,N,M)} S_{\lambda}(\bm{x}, Q) s_{\lambda}(\bm{y})
-  = \sum_{\lambda \subseteq (N,N,M)} S_{\lambda}(\bm{y}, Q) s_{\lambda}(\bm{x})\; ,
+where \(z\) is a formal variable. It has been
+argued~\cite{Foda:2008hn} in that if we interpret the big-Schur
+functions as Plücker coordinates, the expression
+\begin{equation}
+  \mathcal{I}_Q(N,M | \bm{x}, \bm{y})
+  = \sum_{\lambda \subseteq (N,N,M)} \mathcal{S}_{\lambda}(\bm{x}, Q) s_{\lambda}(\bm{y})
+  = \sum_{\lambda \subseteq (N,N,M)} \mathcal{S}_{\lambda}(\bm{y}, Q) s_{\lambda}(\bm{x})\; ,
 \end{equation}
@@
 
+\subsection{Kostka-Foulkes expansion}
+Let us expand these polynomials in a Schur polynomials basis, that is
+\begin{equation}
+P_\lambda(\bm{x}, Q) = \sum_{\mu} K^{-1}_{\lambda \mu}(Q) s_\lambda(\bm{x})\; , 
+\end{equation}
+where \(K^{-1}_{\mu\nu}(Q) \in \mathbb{Z}[Q]\) are inverse
+Kostka-Foulkes polynomials~\cite{Macdonald:1998, Wheeler:2018}.
+
+Expanding now the big-Schur functions as
+\begin{equation}
+\label{eq:kf-exp}
+  \mathcal{S}_\lambda(\bm{x}, Q) = \sum_{\mu} c_{\lambda\mu}(Q) s_\mu(\bm{x})\; ,
+\end{equation}
+we can fix the coefficients \(c_{\lambda\mu}\) in terms of Kostka-Foulkes
+polynomials using the orthogonality relations of these polynomials~\cite{Macdonald:1998}. 
+In particular, there is an inner product in the ring of symmetric functions such that
+\begin{equation}
+  \frac{1}{b_\mu}\langle P_\mu(\bm{x}, Q), P_\mu(\bm{x}, Q)\rangle = \langle
+  \mathcal{S}_\mu(\bm{x}, Q), s_\mu(\bm{x})\rangle = \delta_{\mu\nu}\; .
+\end{equation}
+Therefore,
+\begin{equation}
+  \begin{split}
+\delta_{\mu\nu} & = \sum_\lambda c_{\mu\lambda}  \langle s_\lambda(\bm{x}), s_\mu(\bm{x})\rangle\\
+& = \sum_\lambda \sum_{\pi_1, \pi_2} c_{\mu\lambda}  K_{\lambda \pi_1} K_{\mu \pi_2}
+\langle P_{\pi_1}(\bm{x}, Q), P_{\pi_2}(\bm{x}, Q)\rangle \\ 
+& = \sum_\lambda \sum_{\pi} c_{\mu\lambda} b_{\mu}^{-1}  K_{\lambda \pi} K_{\mu \pi}\; ,
+  \end{split}
+\end{equation}
+and we conclude that 
+\begin{equation}
+\label{eq:indices}
+c_{\mu \nu} = \sum_\pi K_{\mu\pi}^{-1} b_\mu (K_{\nu\pi}^{-1})^T\; .
+\end{equation}
+\marginpar[left]{\tiny Fix these \\ indices later.}
+
+Hence, we conclude that 
+\begin{equation}
+\sum_{\mu} S_{\mu}(\bm{x},Q) s_{\nu}(\bm{y}) = 
+\sum_{\mu , \nu} c_{\mu\nu} s_{\mu}(\bm{x}) s_{\nu}(\bm{y}) 
+\end{equation}
+and everything boils down to the analysis of the
+indices~(\ref{eq:indices}). In has been shown
+in~\cite{Necoechea:2019wbg} that the big Schur functions themselves
+are KP-tau functions\footnote{ We give a simple argument for this fact
+in the remark 2 below.} therefore, the coefficients \((c_\lambda)_\mu
+\equiv c_{\lambda\mu}\) must satisfy Plücker relations for the KP
+hierarchy.
+
+\begin{remark}
+ This expression show somethign interesting. We can formulate this
+problem in terms of partitions defined in the phase
+model. Equivalently, we can use the usual vertex operator
+construction, and not the q-deformed version of Jing~\emph{\cite{Jing1991,
+  Jing1995}}, that is much harder to deal with.
+\end{remark}
+
+If these coefficients satisfy the Toda hierarchy Plücker relations, then 
+one can find a \(Q\)-dependent group element \(G_Q\in GL(\infty)\) such that 
+\begin{equation}
+  c_{\mu\nu} = \bra{\mu} G_Q \ket{\nu}\; ,
+\end{equation}
+where \(\ket{\mu}\) and \(\ket{\nu}\) are phase model
+states. Unfortunatelly, we could not find any good reason to explore
+this expansion in a consistent way. In what follows, we try to explore
+alternative expansions that give a double Schur expansion.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \subsection{Scalar product: Supersymmetric Schur polynomials}
 
-We now aargue that~(\ref{eq:scalar-hl}) is also a tau function of the
-Toda hierarchy in disguise. Let us decompose 
+With this result, we now argue that~(\ref{eq:scalar-hl}) is also a tau
+function of the Toda hierarchy in disguise. Let us decompose
 \begin{equation}
@@
 and \( -Q\bm{y} = (-Qy_1, -Qy_2, \dots)\).
-
-%One can also define the set of coordinates \(\bm{\xi} =
-%\{\xi_i(\bm{x}, Q)\}\), in such a way that
-%\begin{equation}
-% \sum_{m} q_m(\bm{x}, Q) z^m := \sum_{m} h_m(\bm{\xi}) z^m =
-% \prod_k \frac{1}{(1 - \xi_k z)} \;.
-%\end{equation}
-%In this sense, the \(Q\)-Schur function is an ordinary Schur function
-%with respect the variables \(\bm{\xi}\).
 
@@
 \begin{equation}
-  S_\lambda(\bm{t}', Q) = \det \left(h_{\lambda_i - i +j}(\bm{T})\right) = s_\lambda(\bm{T})\; .
+  \mathcal{S}_\lambda(\bm{t}', Q) = \det \left(h_{\lambda_i - i +j}(\bm{T})\right) = s_\lambda(\bm{T})\; .
 \end{equation}
@@
 
-There is something very interesting about this expression. Let us
-expand the supersymmetric Schur polynomials in terms of ordinary Schur
-functions as~\cite[Sec. I.5, exerc. 23]{Macdonald:1998}
+Now we can consider an expansion that differs from
+from~(\ref{eq:kf-exp}). Let us consider the supersymmetric Schur
+polynomials in terms of ordinary Schur functions as~\cite[Sec. I.5,
+  exerc. 23]{Macdonald:1998}
 \begin{equation}
@@
 \end{equation}
-From this expression one can reach to some conclusions, and make some 
-conjectures regarding this expansion.
+
+From this expression we conclude (and speculate) the following.
 
 \begin{remark}
-Observe that the skew Schur polynomials have
-determinant expressions, therefore we have that the
-\(\bm{y}\)-dependent coefficients \(c_{\mu \lambda} = s_{(\lambda/\mu)'}(-Q\bm{y})\)
- have Jacobi-Trudi expressions. We may think of these objects as coordinate dependent 
-Pl\"ucker coordinates. 
+Since the skew Schur polynomials have determinant expressions,
+therefore we have that the \(\bm{y}\)-dependent coefficients \(c_{\mu
+  \lambda} = s_{(\lambda/\mu)'}(-Q\bm{y})\) have Jacobi-Trudi
+expressions. It is tempting to think of these objects as
+\(\bm{y}\)-dependent Pl\"ucker coordinates. In this sense, we would have a curve in 
+the indinite Grassmannian, instead a point.
 \end{remark}
 
-We also know that the left-hand side of~(\ref{eq:tau-hl}) is a tau
-function with respect the coodinates \(\bm{T}\equiv \{\bm{y}\cup(- Q
-\bm{y})\}\) and \(\bm{t}\). Moreover, from the Vertex operator
+\begin{remark}
+Remeber that one of the easiest nontrivial solutions of the KP
+hierarchies are the Schur functions
+themselves~\emph{\cite{Zabrodin2018}}. We have just seen that the
+big-Schur polynomials can be written as supersymmetric Shur
+polynomials, and these are nothing but ordinary Schur polynomials in a
+particular choice of Miwa coordiantes. Then, we also see that 
+the big Schur polynomials are also KP tau functions. This direct conclusion 
+is an alternative proof of this fact, see~\emph{\cite{Necoechea:2019wbg}}.
+\end{remark}
+
+Putting these facts together, we have that the left-hand side
+of~(\ref{eq:tau-hl}) is a tau function with respect the coordinates
+\(\bm{T}\) and \(\bm{t}\). Furthermore, from the Vertex operator
 formalism, we write the Schur polynomials as
@@
   \bra{\bm{0}} e^{J_+(\bm{t})} e^{J_-(\bm{T})} \ket{\bm{0}} = 
-\bra{\bm{0}} e^{J_+(\bm{t})} e^{J_-(\bm{t}')  - J_-(\bm{t}^{(Q)})} \ket{\bm{0}} 
+\bra{\bm{0}} e^{J_+(\bm{t})} e^{J_-(\bm{t}')  - J_-(\bm{t}^{(Q)})} \ket{\bm{0}} \; .
 \end{equation}
@@
   \bra{\bm{0}} e^{J_+(\bm{t})} e^{J_-(\bm{T})} \ket{\bm{0}} = 
-\bra{\bm{0}} e^{J_+(\bm{t})} e^{-J_-(\bm{t}^{(Q)})} e^{J_-(\bm{t}')}  \ket{\bm{0}} 
+\bra{\bm{0}} e^{J_+(\bm{t})} e^{-J_-(\bm{t}^{(Q)})} e^{J_-(\bm{t}')}  \ket{\bm{0}} \; ,
 \end{equation}
@@
 
-In general, we do not expect that expectation values with coordinate dependent 
-group elements \(G_{Q}(\bm{y})\) are tau functions. But in the above example,
-the coordinates combine among themselves and generate a tau function with respect 
-the coordiantes \(\bm{T}\) and \(\bm{t}\).
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\subsection{Determinant formula}
-
-We would like to refine the expressions above. Let us now consider a determinant  
-expression for the scalar product~(\ref{eq:scalar-hl}). Let us continue with 
-the case where \(M\) and \(N\) are very large, but finite. In this case, we write  
-this scalar product as 
-\begin{equation}
-  S_Q(N,M | \bm{x}, \bm{y})  
-= \prod_{j_1, j_2} (1-Q x_{j_1} y_{j_2}) \prod_{k_2, k_2}(1 - x_{k_1} y_{k_2})^{-1}\; .
-\end{equation}
-
-Using the Cauchy's identity for Schur polynomials, that is 
-\begin{equation}
-  \sum_\lambda s_\lambda(\bm{x}) s_\lambda(\bm{y}) = \prod_{i,j} \frac{1}{1 - x_i y_j}\; ,
-\end{equation}
-and using the results we have derives in the Phase model, we have that 
-\begin{equation}
-  \prod_{i,j}\frac{1}{1 - x_i y_j}  \equiv S(N,M|\bm{x}, \bm{y}) = 
-  \frac{\det_{j,k}H(\bm{x},\bm{y})}{\Delta(\bm{x}) \Delta(\bm{y})}\; .
-\end{equation}
-where \(H\) is the matrix~(\ref{eq:h-matrix}). Therefore
-\begin{equation}
-  S_Q(N,M | \bm{x}, \bm{y})  
-= \frac{S(N,M|\bm{x}, \bm{y})}{S(N,M|\bm{x}, Q\bm{y})}\; , 
-\end{equation}
-and we see that the scalar product of the q-boson model is the quotient of 
-scalar products of the phase model. Moreover
-\begin{equation}
-  S_Q(N,M | \bm{x}, \bm{y})  
-  = \frac{\Delta(Q\bm{y})}{\Delta(\bm{y})}
-  \frac{\det H(\bm{x}, \bm{y})}{\det H(\bm{x},Q \bm{y})} \; .
-\end{equation}
-Using that \(\Delta(Q\bm{y})/ \Delta(\bm{y}) = \prod_{j=1}^{N-1} Q^j\), we have 
-\begin{equation}
-  S_Q(N,M | \bm{x}, \bm{y}) =  Q^{N(N-1)/2}\;
-  \frac{\det H(\bm{x}, \bm{y})}{\det H(\bm{x}, Q \bm{y})} \; .
-\end{equation}
-
-Let us denote \(H(\bm{x}, Q\bm{y}) = H_{Q}\), therefore
-\begin{equation}
-  S_Q(N,M | \bm{x}, \bm{y}) = Q^{N(N-1)/2} \det H_{Q}^{-1} \det H
-= Q^{N(N-1)/2} \det \mathcal{H}(\bm{x}, \bm{y}) \; ,
-\end{equation}
-where \(\mathcal{H} = H_{Q}^{-1} H\) is a matrix with components 
-\(\mathcal{H}_{i,j} \equiv \mathcal{H}(x_i, y_j)\) where
-\begin{equation}
-  \mathcal{H}(z, w) = \sum_{k} \frac{(1 - z y_k)}{(1 - Q z y_k)} 
-\frac{(1 - Q x_k w)}{(1 - x_k w)}\; .
-\end{equation}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\subsection{Kostka-Foulkes expansion}
-
-One of the greatest advantages of the expressions above is that we do not 
-need to consider the t-fermions.
-
-It would be very interesting to have some basic understanding of what
-is happenning in the fermionic representation of this tau function.
-Let us expand these polynomials in a Schur polynomials basis, that is
-\begin{equation}
-P_\lambda(\bm{x}, Q) = \sum_{\mu} K^{-1}_{\lambda \mu}(Q) s_\lambda(\bm{x})\; , 
-\end{equation}
-where \(K^{-1}_{\mu\nu}(Q) \in \mathbb{Z}[Q]\) are inverse
-Kostka–Foulkes coefficients, and \(K^{-1}_{\lambda\mu} = 0 \) if
-\(|\lambda|\neq |\mu|\)~\cite{Macdonald:1998, Wheeler:2018}.
-
-Expanding now the \(Q\)-Schur functions as
-\begin{equation}
-  S_\lambda(\bm{x}, Q) = \sum_{\mu} c_{\lambda\mu}(Q) s_\mu(\bm{x})\; ,
-\end{equation}
-we can fix the coefficients \(c_{\lambda\mu}\) in terms of Kostka-Foulkes
-polynomials as follows. Since~\cite{Macdonald:1998} 
-\begin{equation}
-  \frac{1}{b_\mu}\langle P_\mu(\bm{x}, Q), P_\mu(\bm{x}, Q)\rangle = \langle
-  S_\mu(\bm{x}, Q), s_\mu(\bm{x})\rangle = \delta_{\mu\nu}\; ,  
-\end{equation}
-we have
-\begin{equation}
-  \begin{split}
-\delta_{\mu\nu} & = \sum_\lambda c_{\mu\lambda}  \langle s_\lambda(\bm{x}), s_\mu(\bm{x})\rangle\\
-& = \sum_\lambda \sum_{\pi_1, \pi_2} c_{\mu\lambda}  K_{\lambda \pi_1} K_{\mu \pi_2}
-\langle P_{\pi_1}(\bm{x}, Q), P_{\pi_2}(\bm{x}, Q)\rangle \\ 
-& = \sum_\lambda \sum_{\pi} c_{\mu\lambda} b_{\mu}^{-1}  K_{\lambda \pi} K_{\mu \pi}\; ,
-  \end{split}
-\end{equation}
-and we conclude that 
-\begin{equation}
-\label{eq:indices}
-c_{\mu \nu} = \sum_\pi K_{\mu\pi}^{-1} b_\mu (K_{\nu\pi}^{-1})^T\; .
-\end{equation}
-\marginpar[left]{\tiny Fix these \\ indices later.}
-
-Hence, we conclude that 
-\begin{equation}
-\sum_{\mu} S_{\mu}(\bm{x},Q) s_{\nu}(\bm{y}) = 
-\sum_{\mu , \nu} c_{\mu\nu} s_{\mu}(\bm{x}) s_{\nu}(\bm{y}) 
-\end{equation}
-and everything boils down to the analysis of the
-indices~(\ref{eq:indices}).
-
-Do they satisfy Plücker relations? We already know that the
-big Schur functions themselves are KP-tau
-functions, see also~\cite{Necoechea:2019wbg}, therefore, the coefficients
-\((c_\lambda)_\mu \equiv c_{\lambda\mu}\) must satisfy Plücker relations
-for the KP hierarchy.
-
-From the above considerations, we can define 2 set of functions \(F_i(x)\) and
-\(G_i(y)\) such that
-\begin{equation}
- F_i(x) = \sum_{n} F_{i,n} x^n \qquad 
- G_i(y) = \sum_{n} G_{i,n} y^n\; ,
-\end{equation}
-in such a way that 
-\begin{equation}
-  c_{\mu\nu} = \det(F_{i, m_j} | G_{i, n_k})\; ,
-\end{equation}
-and we conclude that the scalar product is, indeed, a Toda tau function. 
-\marginpar[left]{\tiny Improve this\\ discussion}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\subsection{!! Fermionic construction}
+In general, the expectation values with coordinate dependent group
+elements \(G_{Q}(\bm{y})\) are not Toda hierarchy tau functions. But
+in the above example, the coordinates combine among themselves and
+generate a tau function with respect the coordiantes \(\bm{T}\) and
+\(\bm{t}\).
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*
+\section{Conclusions and Perspectives}
+%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*
 

### 2024-04-22 draft.tex
improve section 2.1 and change title.

@@
 
-\title[\(Q\)-Bosons and relations with integrable hierarchies]{\(Q\)-Bosons
-and their relations with integrable hierarchies}
+\title{\(Q\)-Boson model and relations with integrable hierarchies}
 
@@
 
-This section introduces the \(q\)-boson and phase
+This section introduces the \emph{phase} and \emph{\(q\)-boson}
 models~\cite{Bogoliubov:1992, Bogoliubov:1997soj, Bogoliubov2005,
-  Bogoliubov:1997soj, Tsilevich:2006}, and consists of a general
-review of the literature. Consequently, there is nothing particularly
-original, except its presentation.  In spite of that, I believe that
-it has some pedagoginal value, and a good collection of references.
-Here we follow some conventions of~\cite{Wheeler:2010vmq}, but with a
-presentation that is closer to~\cite{Tsilevich:2006}.
+  Tsilevich:2006}. It serves as a comprehensive review of existing
+literature, offering insights into established concepts. While the
+content may not be entirely original, its presentation adds value,
+particularly from a pedagogical standpoint. Additionally, it provides
+a well-curated collection of references for further exploration.
+
+In structuring this discussion, we adopt certain conventions
+from~\cite{Wheeler:2010vmq}, while aligning our presentation style
+more closely with that of~\cite{Tsilevich:2006}.
 
@@
 \end{equation}
-where \(\bar{\mathcal{N}} = \sum_{i=0}^M \mathcal{N}_i\), and
-periodic boundary conditions \(\phi_{M+1} = \phi_0\) and
-\(\phi_{M+1}^\dagger = \phi_0^\dagger\).
-
-These operators appear in the context of quantum optics, and for
-this reason it is referred as the \emph{phase model}. It 
-corresponds to the strongly correlated \(q\)-bosons
-model~\cite{Bogoliubov:1997soj} the we will define soon.
+where \(\bar{\mathcal{N}} = \sum_{i=0}^M \mathcal{N}_i\) is the total
+number operator, and we also impose periodic boundary conditions
+\(\phi_{M+1} = \phi_0\) and \(\phi_{M+1}^\dagger = \phi_0^\dagger\).
+
+These operators appear in the context of quantum optics, and for this
+reason, it is referred to as the \emph{phase model}. It corresponds to
+the strongly correlated limit of the \(q\)-bosons
+model~\cite{Bogoliubov:1997soj}, which we will define shortly.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
 \subsubsection{Representation}
-Let us define the i\(^{th}\)-vacuum \(\ket{0}_i\) by \(\phi_i\ket{0}_i
-= 0\).  The representation of this Hilbert space, denoted by
-\(\mathcal{F}_i\), is given by \(\ket{n_i}_i = \phi_i^\dagger
-\ket{0}_i\).
+The representation of the phase model algebra, is constructed using
+the vacuum state, defined by \(\ket{0}_i\) by \(\phi_i\ket{0}_i =
+0\). In this context, the state with \(n_i\) bosons (oscillators) is
+given by \(\ket{n_i}_i = \phi_i^\dagger \ket{0}_i\).
 
 Given the vacuum \(\ket{\bm{0}} = \ket{0}_0\otimes \ket{0}_1
-\otimes \cdots \otimes  \ket{0}_M\),
-the Fock space is defined as 
+\otimes \cdots \otimes  \ket{0}_M\), the Fock space is defined as 
 \begin{equation}
@@
   \left\{\ket{\vec{n}} = \ket{n_0}\otimes \ket{n_1} \otimes \cdots
-  \otimes \ket{n_M} \ | \ n_i \in \mathbb{N} \right\}\; .
+  \otimes \ket{n_M} \ | \ n_i \in \mathbb{N} \right\}\; ,
 \end{equation}
@@
   \ket{\vec{n}} = \ket{n_0}\otimes \ket{n_1} \otimes \cdots \otimes \ket{n_M} 
- : =  (\phi_0^\dagger)^{n_0} (\phi_1^\dagger)^{n_1} \cdots  (\phi_M^\dagger)^{n_M} \ket{\bm{0}} \; ,
-\end{equation}
-where we write \(\phi_i^\dagger \equiv \bm{1} \otimes  \cdots \otimes
-\phi_i^\dagger \otimes \cdots \otimes \bm{1}\).
-These states are normalized, that is
-\(\bracket{\vec{n}}{\vec{m}}=\delta_{\vec{n}, \vec{m}}\).
-
-Finally, the actions of the operators \(\mathcal{N}_i\) and \(\pi_i\) are
+ \equiv  (\phi_0^\dagger)^{n_0} (\phi_1^\dagger)^{n_1} \cdots  (\phi_M^\dagger)^{n_M} \ket{\bm{0}} \; ,
+\end{equation}
+with \(\phi_i^\dagger \equiv \bm{1} \otimes \cdots \otimes
+\phi_i^\dagger \otimes \cdots \otimes \bm{1}\).  Moreover, it is easy
+to see that these states are normalized, that is
+\(\bracket{\vec{n}}{\vec{m}}=\delta_{\vec{n}, \vec{m}}\).  Finally,
+the actions of the operators \(\mathcal{N}_i\) and \(\pi_i\) are
 \begin{equation}
@@
 associate a partition \( \lambda = (0^{n_0} 1^{n_1} 2^{n_2} \cdots
-M^{n_M})\). Observe that this correspondence is not unique, since the
-partition \(\lambda\) ignores the number \(n_0\) of particles in the
-site \(0\). From the total number of particles, we can, on the other hand,
-determine the number of particles \(n_0 = N - \ell(\lambda)\), where
-\(\ell(\lambda)\) is the number of rows in the Young diagram defined
-by \(\lambda\). This is very important, since in our consideations 
-below, the \(N\) particle sector is fixed, thanks to the integrability of 
-the model, therefore, the number \(n_0\) is known once we define the 
-partition \(\lambda\).
-
-In fact, from the correspondence above, Wheeler~\cite{Wheeler:2010vmq}
-defines a map \(\mathcal{M}_\psi: \mathcal{F}\to
-\mathcal{F}^{(0)}_\psi\), where \(\mathcal{F}^{(0)}_\psi\) is the
-space Fock space of charged free fermions built upon the neutral
-(fermionic) vacuum. We will discuss a map from the bosonic space to
-the space of symmetric functions.
+M^{n_M})\).  It's worth noting that this correspondence is not unique,
+as the partition \(\lambda\) does not account for the number of
+particles \(n_0\).  If the total number of particles \(N\) is known,
+we can determine \(n_0 = N - \ell(\lambda)\), where \(\ell(\lambda)\)
+represents the number of rows in the Young diagram defined by
+partition \(\lambda\). This determination is crucial, as in our
+subsequent considerations, the \(N\) particle sector remains fixed due
+to the integrability of the model. Consequently, the value of \(n_0\)
+becomes known once we specify the partition \(\lambda\).
+
+Finally, based on the correspondence we mentioned above,
+Wheeler~\cite{Wheeler:2010vmq} defines a map \(\mathcal{M}_\psi:
+\mathcal{F} \to \mathcal{F}^{(0)}_\psi\), where
+\(\mathcal{F}^{(0)}_\psi\) denotes the Fock space of charged free
+fermions constructed from the neutral (fermionic) vacuum.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 \end{equation}
-where \(x \in \mathbb{S}^1\). From where we can build the monodromy
-matrix
+where \(x \in \mathbb{S}^1\). We also have the monodromy matrix
 \begin{equation}
@@
 where \(\mathbb{B}(x) = x^{M/2} B(x)\) and \(\mathbb{C}(x) = x^{M/2} C(1/x)\).
-When the set \(\{ y_j \ | \ j
-=1, \dots , N\}\) satisfies the Bethe equations
+
+When the set \(\{ y_j \ | \ j =1, \dots , N\}\) satisfies the Bethe equations
 \begin{equation}
@@
 \end{equation}
-we says that the Bethe states are \emph{on-sheel}. In what follows, we will 
-only consider \emph{off-shell} states.
+we say that the Bethe states are \emph{on-shell}; otherwise, we have
+\emph{off-shell} states. In what follows, we will only consider
+\emph{off-shell} states.
 

### 2024-04-23 draft.tex
improve tau function section

@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 \end{equation}
-where \(\mathbb{B}(x) = x^{M/2} B(x)\) and \(\mathbb{C}(x) = x^{M/2} C(1/x)\).
-
-When the set \(\{ y_j \ | \ j =1, \dots , N\}\) satisfies the Bethe equations
+where \(\mathbb{B}(y) = y^{M/2} B(y)\) and \(\mathbb{C}(y) = y^{M/2} C(1/y)\).
+
+When the coordinates \(\{ y_j \ | \ j =1, \dots , N\}\) satisfy the Bethe equations
 \begin{equation}
@@
 \subsection{q-Bosons}
-The Hamiltonian for the q-boson model is
-given by
-\begin{equation}
-  \mathcal{H} = -\frac{1}{2} \sum_{n=0}^M
-  \left(b_n^\dagger b_{n+1} + b_n b_{n+1}^\dagger \right) + \bar{\mathcal{N}}\; ,
-\end{equation}
-where \(\bar{\mathcal{N}} = \sum_{i=0}^M \mathcal{N}_i\), and as in
-the case of the phase mode, we impose periodic boundary conditions
-\(b_{M+1} = b_0\) and \(b_{M+1}^\dagger = b_0^\dagger\). The set of
-operators \(\{b_i, b^\dagger,\mathcal{N}_i\}_{i=0}^M\) form the
-\(M+1\) copies of independent q-boson algebras:
+The set of operators \(\{b_i, b^\dagger,\mathcal{N}_i\}_{i=0}^M\) constitute
+\(M+1\) independent q-boson algebras defined by
 \begin{equation}
@@
 \end{equation}
-where we write the deformation parameter as \(Q = q^{-2}\). 
-In the limit \(Q\to 1\), the q-bosons are ordinary bosons, and 
-the limit \(Q\to 0\) (\(q\to \infty\)) gives the phase model we have discussed above.
+where we denote the deformation parameter as \(Q = q^{-2}\).
+
+The q-boson model is characterized by its Hamiltonian
+\begin{equation}
+  \mathcal{H} = -\frac{1}{2} \sum_{n=0}^M
+  \left(b_n^\dagger b_{n+1} + b_n b_{n+1}^\dagger \right) + \bar{\mathcal{N}}\; ,
+\end{equation}
+where \(\bar{\mathcal{N}} = \sum_{i=0}^M \mathcal{N}_i\), and we also
+impose periodic boundary conditions \(b_{M+1} = b_0\) and
+\(b_{M+1}^\dagger = b_0^\dagger\).
+
+In the limit \(Q \to 1\), the q-bosons behave as ordinary bosons,
+while the limit \(Q \to 0\) (\(q \to \infty\)) corresponds to the
+phase model discussed earlier.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
 \subsubsection{Representation}
-We build the representation of the q-boson algebra with some modifications of the phase 
-model we have just discussed. Let us define the i\(^{th}\)-vacuum \(\ket{0}_i\) by
-\(b_i\ket{0}_i = 0\). The representation of this Hilbert space, denoted by
-\(\mathcal{F}_i^Q\), is given by \(\ket{n_i}_i \propto b_i^\dagger \ket{0}_i\).
+We construct the representation of the q-boson algebra with
+modifications of the phase model discussed earlier. Let us define the
+\(i\)-th vacuum \(\ket{0}_i\) such that \(b_i\ket{0}_i = 0\). The
+representation of this Hilbert space, denoted by \(\mathcal{F}_i^Q\),
+is given by \(\ket{n_i}_i \propto b_i^\dagger \ket{0}_i\).
 
@@
   \left\{\ket{\vec{n}} = \ket{n_0}\otimes \ket{n_1} \otimes \cdots
-  \otimes \ket{n_M} \ | \ n_i \in \mathbb{N} \right\}\; .
-\end{equation}
-where the actions of the operators \(\{b_i, b_i^\dagger\}\) are given by 
-the following relations
+  \otimes \ket{n_M} \ | \ n_i \in \mathbb{N} \right\}\; ,
+\end{equation}
+where the actions of the operators \(\{b_i, b_i^\dagger\}\) are given
+by the following relations
 \begin{subequations}
@@
 \end{subequations}
-Observe that the modes \(\{b_0, b_0^\dagger \}\) satisfies a slightly
-different relation.
-
-Therefore, the states \(\ket{\vec{n}} \in \mathcal{F}_Q\) are defined as 
+
+Consequently, the states \(\ket{\vec{n}} \in \mathcal{F}_Q\) are
 \begin{equation}
@@
 \end{equation}
-where we write \(\phi_i^\dagger \equiv \bm{1} \otimes  \cdots \otimes
-\phi_i^\dagger \otimes \cdots \otimes \bm{1}\).
-Moreover, the norm of these states satisfy 
+where we write \(b_i^\dagger \equiv \bm{1} \otimes  \cdots \otimes
+b_i^\dagger \otimes \cdots \otimes \bm{1}\).
+Moreover, this space has an inner product that satisfies
 \begin{equation}
@@
 \end{equation}
-As before, we can associate to a given states \(\ket{\vec{n}} =
-\ket{n_0, n_1, \dots, n_M}\), a partition -- or Young diagram --
-\(\lambda = (1^{n_1} 2^{n_2} \cdots M^{n_M})\), but now we assume that 
-there is a proportionaly  factor relating these two objects, 
+
+Similar to the phase model, we can associate to a given states
+\(\ket{\vec{n}} = \ket{n_0, n_1, \dots, n_M}\), a Young diagram
+\(\lambda = (1^{n_1} 2^{n_2} \cdots M^{n_M})\). It is useful to define
+a proportionaly factor relating these two objects
 \begin{equation}
@@
 where \(p_i(\lambda)\) denotes the number of parts of size \(i\) in
-the partition. These partition states satisfy \(
-\bracket{\lambda}{\mu}_Q = b_\lambda(Q) \delta_{\lambda,\mu}\). Once
-again, we see that the correspondence is not unique since the number
-of oscilators at the site \(i=0\) are completely ignored in the
-partition notation. And finally, if the number of particles is fixed,
-then \(n_0 = N - \ell(\lambda)\).
+the partition. These partition states satisfy
+\(\bracket{\lambda}{\mu}_Q = b_\lambda(Q) \delta_{\lambda,\mu}\). Once
+again, we see that the correspondence is not unique, as the number of
+oscillators at site \(i=0\) is completely ignored in the partition
+notation. Finally, if the number of particles is fixed, then \(n_0 = N
+- \ell(\lambda)\).
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
     x^{-1/2} & (1 - Q)^{\frac{1}{2}} b_n^\dagger \\ (1 - Q)^{\frac{1}{2}} b_n & x^{1/2}
-  \end{pmatrix}_a\; .
-\end{equation}
-The monodromy matrix is 
+  \end{pmatrix}_a\; ,
+\end{equation}
+and the monodromy matrix is 
 \begin{equation}
@@
 
-The eigenstates of the Hamiltonian are of the form
+As before, the eigenstates of the Hamiltonian have the form
 \begin{equation}
@@
 \end{equation}
-where \(\mathbb{B}(x, Q) = x^{M/2} B(x, Q)\) and \(\mathbb{C}(x, Q) =
-x^{M/2} C(1/x, Q)\).  When the set of parameters \(\{ y_j \ | \ j =1, \dots , N\}\)
-satisfies the Bethe equations
+where \(\mathbb{B}(y, Q) = y^{M/2} B(y, Q)\) and \(\mathbb{C}(y, Q) =
+y^{M/2} C(1/y, Q)\). When the parameters \(\{ y_j \ | \ j =1, \dots , N\}\)
+satisfy the Bethe equations given by
 \begin{equation}
@@
 \end{equation}
-we say that the states \(\ket{\Psi(y_1, \dots, y_N; Q)}\) are \emph{on-shell} Bethe states, otherwise, 
-we have \emph{off-shell} Bethe states.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+we have \emph{on-shell} Bethe states; otherwise, they are \emph{off-shell} states.
 
@@
 
-This section discusses the presence of integrable hierarchies in the phase
-model.  We argue that it is also a Toda hierachy tau function, and we
-discuss some consequences of this fact, in particular, its connection
-with a matrix model. Moreover, we also point the presence of
-correlation functions in this model that also satisfy the KP hierarchy
-equations. We finally argue how these results also imply that these
-scalar products are also KP hierarchy tau functions, what is expected from 
-the results of Wheeler~\cite{Wheeler:2010vmq}.
-
-Bogoliubov~\cite{Bogoliubov2005}, has shown that the scalar product of
-two vectors in the \(N\)-particles sector in a chain of length \(M+1\)
-is
+This section explores the presence of integrable hierarchies in the
+phase model. We demonstrate its correspondence to a Toda hierarchy tau
+function and discuss some implications, particularly its connection to
+a matrix model. Additionally, we highlight the existence of
+correlation functions in this model that satisfy the KP hierarchy
+equations. Finally, we argue that these results imply that these
+scalar products are also KP hierarchy tau functions, aligning with the
+findings of Wheeler~\cite{Wheeler:2010vmq}.
+
+Bogoliubov~\cite{Bogoliubov2005} has demonstrated that the scalar
+product of two vectors in the \(N\)-particle sector of a chain with
+length \(M+1\) is
 \begin{equation}
@@
   \mathcal{I}(N, M|\bm{x}, \bm{y}) & =
-  \bra{\bm{0}} \prod_{j=1}^N \mathbb{C}(x_j) \prod_{j=1}^N \mathbb{B}(y_j) \ket{\bm{0}} \\ 
+  \bra{\bm{0}} \prod_{i=1}^N \mathbb{C}(x_i) \prod_{j=1}^N \mathbb{B}(y_j) \ket{\bm{0}} \\ 
  & = \frac{ \det H(\bm{x},\bm{y})}{ \prod_{i<j}(x_i - x_j)(y_i - y_j)} \; ,
@@
 \end{equation}
-where \(H\) is an \(N\times N\) matrix with components
+where \(H\equiv [H_{ij}]_{i,j=1}^N\) is an \(N\times N\) matrix with components
 \begin{equation}
@@
 \subsection{Toda tau functions}
-In this section, we argue that the scalar product defined above is a tau function 
-of the Toda hierarchy. Let us first write the function \(H(z,w)\) as the geometric sum 
+We now argue that the scalar product defined above is a tau function
+of the Toda hierarchy. Let us first write the function \(H(z,w)\) as
+the geometric sum
 \begin{equation}
 \label{eq:h-exp}
-  H(z,w) = \frac{1 - (zw)^{M+N}}{1 - zw} = \sum_{k=1}^{M+N} (zw)^{k-1} \; , 
-\end{equation}
-in such a way that 
-we can write
+  H(z,w) = \frac{1 - (zw)^{M+N}}{1 - zw} = \sum_{k=1}^{M+N} (zw)^{k-1} \; .
+\end{equation}
+Hence, we express the determinant of the \(H\) matrix as
 \begin{equation}
@@
 \end{equation}
-and we can thing of this expression as the product of a \(N\times (N+M)\) matrix \(\mathcal{X}\)
-and another \((M + N)\times N\) matrix \(\mathcal{Y}\), given by
+Furthermore, we can interpret this expression as the result of
+multiplying an \(N\times (N+M)\) matrix \(\mathcal{X}\) by another
+\((M + N)\times N\) matrix \(\mathcal{Y}\), which are given by
 \begin{equation}
@@
   y_1^{N+M-1} & y_2^{M+N-1} & \dots & y_N^{M+N-1}
-  \end{pmatrix}
+  \end{pmatrix}\; ,
 \end{equation}
@@
 \begin{equation}
-  \det_{i,j} \left(H(x_i, y_j)\right) =  \det_{i,j} \left( \mathcal{X}\mathcal{Y}\right)
+  \det_{i,j} \left(H(x_i, y_j)\right)
+  \equiv  \det_{i,j} \left( \mathcal{X}\mathcal{Y}\right)
   = \sum_{0 \leq \ell_{N} \leq \dots \leq \ell_1\leq N+M } \det_{ik}(x_i^{\ell_k}) \det_{ik}(y_j^{\ell_k})\; .
@@
 
-Let us define two sets of Miwa coordinates \(\bm{t} = (t_1, t_2, \dots)\)
-and \(\bm{t}' = (t'_{-1}, t'_{-2}, \dots)\) by
+Define two sets of Miwa coordinates \(\bm{t} = (t_1, t_2, \dots)\) and
+\(\bm{t}' = (t'_{-1}, t'_{-2}, \dots)\) by
 \begin{equation}
   t_p = \frac{1}{p}\sum_{j=1}^N x_j^p\qquad 
-  t'_{-p} = \frac{1}{p}\sum_{j=1}^N y_j^p\; ,
-\end{equation}
-therefore, we have 
+  t'_{-p} = \frac{1}{p}\sum_{j=1}^N y_j^p\; .
+\end{equation}
+One can write the inner product in terms of these coordinates, that is
 \begin{equation}
@@
 This expression is known to be a tau function for \(M, N\to \infty\).
-In fact, using the free fermions representation~\cite{Alexandrov:2012tr},
-it can be written as
-\begin{equation}
-\sum_{\lambda } s_{\lambda}(\bm{t}) s_{\lambda}(\bm{t}')
+As such, utilizing the free fermions representation~\cite{Alexandrov:2012tr}, it can be written as
+\begin{equation}
+  \lim_{M,N\to \infty}\mathcal{I}(N, M|\bm{t}, \bm{t}')
  = \bra{\bm{0}} e^{\bm{J}_+(\bm{t})} e^{-\bm{J}_-(\bm{t}')} \ket{\bm{0}}\; .
 \end{equation}
-It is a tau function of the Toda hierarchy with trivial element \(1 \in GL(\infty)\).
-In fact, it is nothing but the Cauchy's identity
+It is a tau function of the Toda hierarchy with trivial element \(\mathbb{1} \in GL(\infty)\).
+It is nothing but the Cauchy's identity
 \begin{equation}
@@
 \end{equation}
-where \(\bm{t}\) and \(\bm{t}'\) are two independent sets of Miwa coordinates.
-
-Putting all these facts together, the truncation for finite \(M\) and
-\(N\) are also tau functions of the Toda hierarchy. More specifically,
-according to~\cite{Alexandrov:2012tr, Kharchev:1991gd,
+
+Bringing all these facts together, the truncation for finite \(M\) and
+\(N\) also yields tau functions of the Toda hierarchy. More
+specifically, according to~\cite{Alexandrov:2012tr, Kharchev:1991gd,
   Zabrodin:2010ii}, the truncation of the tau function corresponds to
@@
 case, one can define a diagonal \((N+M)\times (N+M)\) matrix
-\(\mathcal{H} = \textrm{diag}(h_0, \dots, h_{M+N-1})\). Therefore, if
-we repeat the arguments above, we find
+\(\mathcal{H} = \textrm{diag}(h_0, \dots, h_{M+N-1})\). Consequently,
+if we repeat the arguments above, we find
 \begin{equation}
@@
 \end{equation}
-where \(h_\lambda = 1\) if \(\lambda \subset [N,M]\) and it vanishes
-otherwise.  In this case, we have that this tau function is of the
-type considered in~\cite{orlov:2001}. Evidently that it is a trivial
-example, but we expect that in the case of the \(q\)-Boson model,
-where \(q\) is finite, the diagonal terms \(h_\lambda\) are more
-interesting. We will discuss this case next section.
+Here, \(h_\lambda\) is equal to \(1\) if \(\lambda \in [N,M]\), and it
+is zero otherwise.
+
+In this case, we find that this tau function is a trivial example of
+the tau functions considered in~\cite{orlov:2001}. We anticipate that
+in the analysis of more general correlation functions, the diagonal
+terms \(h_\lambda\) will be more interesting. We will revisit this
+discussion soon.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
 \subsubsection{Matrix Models}
-
-It is also interesting to notice that the particular case when \(M\to
-\infty\), but with a finite number of particles \(N\). The
-partitions \(\lambda \in [N, \infty]\) satisfy the condition
+It is also interesting to note the particular case when \(M\to
+\infty\), but with a finite number of particles \(N\). In this case,
+the partitions \(\lambda \in [N, \infty]\) satisfy the condition
 \(\ell(\lambda) \leq N\). Therefore, the scalar
@@
 \begin{equation}
-  \mathcal{I}_N(\bm{t}, \bm{t}') = \sum_{\substack{\lambda \\ \ell(\lambda) \leq N}}
-  s_\lambda(\bm{t})  s_\lambda(\bm{t}') \; .
+  \mathcal{I}_N(\bm{t}, \bm{t}')\equiv 
+  \lim_{M\to \infty}\mathcal{I}(N, M|\bm{t}, \bm{t}')
+  = \sum_{\substack{\lambda \\ \ell(\lambda) \leq N}} s_\lambda(\bm{t})  s_\lambda(\bm{t}') \; .
 \end{equation}
@@
 
-What is very interesting about this result is what it implies when we 
-impose the Bethe equations~(\ref{eq:bethe_eq}). Suppose that 
-\(x_j = e^{-ip_j} \in \mathbb{S}^1\), and we set \(\bm{t}' = - \bm{t}^\star\).
-In this particular case, we have
+From the results of~\cite{Zabrodin:2010ii}, see citations therein, we
+have an interesting consequence of this representation. Impose the
+Bethe equations~(\ref{eq:bethe_eq}) to one set of variables, say \(x_j
+= e^{-ip_j} \in \mathbb{S}^1\). Additionally, let us set \(\bm{t}' = -
+\bm{t}^\star\). In this particular case, we have
 \begin{equation}
@@
 Then, the phase model is equivalent to an ensemble of \(N\) 2D Coulomb
-particles on a circle. In this case, we have that the quantities
+particles on a circle. In this case, we find that the quantities
 \(\lambda_{\ell}\) are eigenvalues of a matrix \(U\).
 
-Moreover, Zabrodin~\cite{Zabrodin:2010ii}, see citations therein, we also know
-that under the rescaling \(t_k \to T_k/ \hbar\), \(t'_k \to T_{-k}/ \hbar\)
-and \(N = T_0/ \hbar\), we get the dispersion limit tau function 
+Furthermore, according to Zabrodin~\cite{Zabrodin:2010ii}, see
+citations therein, we also know that under the rescaling \(t_k \to
+T_k/ \hbar\), \(t'_k \to T_{-k}/ \hbar\) and \(N = T_0/ \hbar\), we
+obtain the dispersion limit tau function
 \begin{equation}
@@
 \end{equation}
-a free energy, from the viewpoit of the matrix integral partition
-function.  It is not clear how one can use this fact to determine
-properties of the integrable model, but it might be possible to study
-the analytic structure of the free energy \(F_0\) to gain some
-understanding on the Bether roots \(\bm{x}\).  This problem is
-currently under further investigation, and we hope to report some new
-results elsewhere.
-
+that is a free energy, from the viewpoit of the matrix integral
+partition function.
+
+It remains unclear how one can use this fact to determine properties
+of the integrable model, but it might be possible to study the
+analytic structure of the free energy \(F_0\) to gain some
+understanding of the Bethe roots \(\bm{x}\). This problem is currently
+under further investigation, and we hope to report new results
+elsewhere.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \subsubsection{KP tau function}
-Finally, one may also observe that if we fix one set of coordinates,
-say \(\bm{y}\), then we write
+Lastly, one may also observe that if we fix one set of coordinates,
+say \(\bm{y}\), then we can write
 \begin{equation}
@@
 \end{equation}
-with coefficients \(c_\lambda(\bm{y}) = \det (h_{\lambda_i-i +j}(\bm{y}))\).
-But now, it is a trivial to notice that these are Plücker
-coordinates in the Jacobi-Trudi form, see~\cite{Miwa2000, Alexandrov:2012tr}.
-
-Therefore, the expression
+with coefficients \(c_\lambda(\bm{y}) = \det (h_{\lambda_i-i
+  +j}(\bm{y}))\). But now, it is trivial to notice that these are
+Plücker coordinates in the Jacobi-Trudi form, as seen
+in~\cite{Miwa2000, Alexandrov:2012tr}.
+
+Hence, the following expression
 \begin{equation}
@@
 \end{equation}
-is also a KP tau function: A fact that we already know
-from~\cite{Wheeler:2010vmq}, where the author has proved this
-statement using the free fermions formalism.
+is also a KP tau function, a fact that we already know
+from~\cite{Wheeler:2010vmq}, where the author proved this statement
+using the free fermions formalism.
 
@@
 \subsection{Correlation functions}
-In~\cite{Bogoliubov2005}, Bogoliubov has also shown that the
+Bogoliubov has also shown in~\cite{Bogoliubov2005}, that the
 correlation functions
@@
   \prod_{k=1}^{N-1} y_{j}^{M/2}
-  \left( \prod_{t<N} \frac{y_N - y_t}{y_t} \right) \frac{\det Q}{\det H} S(N,M|\bm{x}, \bm{y})
+  \left( \prod_{t<N} \frac{y_N - y_t}{y_t} \right)
+  \frac{\det Q}{\det H} \mathcal{I}(N,M|\bm{x}, \bm{y})
 \end{equation}
 \end{subequations}
-where the \(N\times N\) matrix \(Q\) has components 
+where \(Q\) is an \(N\times N\) matrix with components 
 \begin{equation}
@@
 \end{equation}
-where \(H_{jk} = H(x_j, y_k)\) are the components of the matrix \(H\) in~(\ref{eq:h-matrix}). 
-Since the components \(Q_{jN}\) are independent of the coordinates \(y\), 
-we cannot write the above expression as a Toda hierarchy tau
-function.
+and \(H_{jk} = H(x_j, y_k)\) are the components of the matrix \(H\)
+in~(\ref{eq:h-matrix}). Since the components \(Q_{jN}\) are
+independent of the coordinates \(y\), we cannot express the above
+expression as a Toda hierarchy tau function.
 

### 2024-04-24 draft.tex
finish phase model

@@
 and \(H_{jk} = H(x_j, y_k)\) are the components of the matrix \(H\)
-in~(\ref{eq:h-matrix}). Since the components \(Q_{jN}\) are
-independent of the coordinates \(y\), we cannot express the above
+in~(\ref{eq:h-matrix}).  The components \(Q_{jN}\) are independent of
+the coordinates \(y\); therefore, we cannot express the above
 expression as a Toda hierarchy tau function.
@@
 \end{equation}
-therefore we have
-\begin{equation}
-  A_m(N, M|\bm{x}, \bm{y}\setminus{y_N})  \propto
+then
+\begin{equation}
+  A_m(N, M|\bm{x}, \bm{y}\setminus \{y_N\})  \propto
   \frac{\det Q}{\Delta(x)} \equiv \mathcal{A}_m(N,M|\bm{x}, \bm{y})\; ,
 \end{equation}
-and we are going to treat the coordinates \(\{ \bm{y} \}\) as
-a set of \(N\) fixed parameters. Furthermore, we write \(N\)
-functions \(\bm{F}(z) = (F_1, \dots, F_N)\) defined as
+and we treat the coordinates \(\{ \bm{y} \}\) as a set of \(N\) fixed
+parameters.
+
+Furthermore, we define vector field \(\bm{F}(z) = (F_1, \dots, F_N)\),
+where the components are given by
 \begin{equation}
@@
   F_j(z) = \frac{1 - (y_j z)^{M + N}}{1 - y_j z} = \sum_{n=0}^{N + M - 1} (y_j z)^n
-  \equiv \sum_{n=0}^{M + N -1} f_{j, n} z^n\; , 
-\end{equation}
-where we will also assume the \(M + N - 1 \in 2 \mathbb{Z}\).
+  \equiv \sum_{n=0}^{M + N -1} f_{j, n} z^n\; .
+\end{equation}
 
@@
 \end{equation}
-and with the aid of the Cauchy-Binet formula, we write  
+By utilizing the Cauchy-Binet formula, we can write
 \begin{equation}
@@
 \end{equation}
-Now we basically use the definition of the Schur polynomials and the
-Jacobi-Trudi expression of Plücker coordinates, see~\cite{Alexandrov:2012tr}. 
-Finally, this last expression becomes
+
+We now employ the definition of the Schur polynomials and utilize the
+Jacobi-Trudi expression of Plücker coordinates, as detailed
+in~\cite{Alexandrov:2012tr}. Ultimately, this leads to the following
+expression:
 \begin{equation}
@@
 \end{equation}
-where \(c_\lambda \equiv \det_{jk}(y_j^{\ell_k})\),  \(\ell_k = \lambda_k - k +N\). Therefore,
-we have that this expression is a KP tau function.
-
+where \(c_\lambda \equiv \det_{jk}(y_j^{\ell_k})\),  \(\ell_k = \lambda_k - k +N\).
+Therefore, we have that this expression is a KP tau function.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\subsubsection{Skew Schur polynomials expansion}
 From the coordinate expansion of the Bethe
@@
 \end{equation}
-where \(s_{\mu/\lambda}\) are Schur polynomials~\cite{Macdonald:1998}.
+where \(s_{\mu/\lambda}\) are skew Schur polynomials~\cite{Macdonald:1998}.
 Therefore, we can write the correlation function~(\ref{eq:correlation}) as 
@@
 we sum over all partitions, since \(s_{\mu/(m)} = 0\), \(\forall \) \(\mu \not \supset (m)\).
-This expression also shows that the expression \(P_m\) can also be written as a tau function.
-
-\subsubsection{Skew Schur polynomials expansion}
+This expression also shows that the expression \(A_m\) can also be written as a tau function.
+
 More generally, let us consider the correlation functions 
@@
 where \(N' + \ell(\lambda_2) + n_0 = N' + \ell(\lambda_2) + n'_0\) and
-we have also used that the expression for any Young diagram
+we have also used that the skew Schur polynomial for any Young diagram
 \(\mu\) that does not contain \(\lambda_1\) and \(\lambda_2\)
@@
 Therefore, this correlation function is nothing but the off-shell
-norm~(\ref{eq:scalar}) times a finite sum over skew Schur functions.
-This expression shows that the existence of these tau functions is not
-a trivial fact of this model.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\subsection{More tau functions}
-Based on the discussion we had so far, and on the general map between 
-the Phase model and the free fermions, one can understand the general form 
-of tau functions in this phase model context.
+norm~(\ref{eq:scalar}) times a finite sum over skew Schur
+functions. This expression illustrates that the existence of these tau
+functions is not a trivial aspect of this model.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\subsubsection{Schur polynomials expansion of some correlation functions}
+Indeed, it is worth noting that other quantities, which are not tau
+functions, may have interesting Schur polynomial expansions. Consider
+the state calculated in~\cite{Bogoliubov2005}.
+\begin{equation}
+  \ket{\mathcal{Y}} =
+ \sum_{\vec{n}} \prod_{j=0}^M \ket{n_j} = \sum_\mu \ket{\mu} \qquad \sum_j n_j = N\; .
+\end{equation}
+Then
+\begin{equation}
+ \bra{\mathcal{Y}} \prod_{j=1}^N B(\bm{x}) \ket{\nu} = 
+  \sum_{\lambda\subseteq (N,N,M)} s_{\lambda/\nu}(\bm{x})\; . 
+\end{equation}
+While it may not be a tau function, it has an interesting expansion as
+a sum of skew Schur polynomials.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\subsection{General tau functions}
+Based on the discussion we have had so far, and on the general mapping
+between the Phase model and free fermions, one can grasp the general
+form of tau functions in the context of this phase model.
 
 Let us consider the vertex operator construction~\cite{Okounkov2001},
-see also~\cite{Alexandrov:2012tr, Wheeler:2010vmq}. Consider the
-vacuum \(\ket{\bm{0}}\), a ``Dirac sea'', defined by the conditions \(\psi_m\ket{\bm{0}}
-= \psi_n^\star\ket{\bm{0}} =0 \) for \(m<0\) and \(n \geq 0\), where 
-\(\psi_n\) are components of a holomorphic free fermionic field. 
-In this formalism, the partition states are given by 
+as also discussed in~\cite{Alexandrov:2012tr, Wheeler:2010vmq}. We
+consider the vacuum state \(\ket{\bm{0}}\), often referred to as a
+``Dirac sea'', defined by the conditions \(\psi_m\ket{\bm{0}} =
+\psi_n^\star\ket{\bm{0}} =0 \) for \(m<0\) and \(n \geq 0\), where
+\(\psi_n\) are components of a holomorphic free fermionic field. In
+this formalism, the partition states are given by
 \begin{equation}
@@
 where \(\textrm{sign}(\sigma) = \pm 1\) are defined in a such a way
-that the Shur coefficients of the vertex operators, defined below,
+that the Shur coefficients of the vertex operators (defined below)
 have positive coefficients.
 
-The pairs \((a_j|b_j)_{j=1}^d\) define the Frobenius notation of the
-partition \(\mu = (\mu_1, \mu_2, \dots, \mu_\ell)\), that is given by
-\(a_j = \mu_j - j\) and \(b_j = \mu'_j - j\), where \(d\) is the
-number of boxes in the diagonal of the Young diagram, and \(\mu'\) is
-its conjugate, or transpose.  From these definitions, we have the
-equivalence
+The pairs \(\{(a_j|b_j)\}_{j=1}^d\) define the Frobenius notation of
+the partition \(\mu = (\mu_1, \mu_2, \dots, \mu_\ell)\). In this
+notation, \(a_j\) is given by \(\mu_j - j\) and \(b_j\) is given by
+\(\mu'_j - j\), where \(d\) represents the number of boxes in the
+diagonal of the Young diagram, and \(\mu'\) is its conjugate, or
+transpose, diagram. From these definitions, we have the equivalence
 \begin{equation}
@@
 
-Finally, the \(\mathfrak{gl}(\infty)\) has generators given by the
+Finally, the \(\mathfrak{gl}(\infty)\) algebra has generators given by the
 bilinears \(X = \sum_{j, j \in \mathbb{Z}} x_{ij} \bm{\colon} \psi_i \psi_j^\star\bm{\colon}
@@
 
-Observe that these elements have finitely many non zero elements, and
-that these include diagonal terms \(\bm{\colon} \psi_i
-\psi_i^\star\bm{\colon}\), upper and lower triangular terms. These are
-equivalent to the Number operators \(\mathcal{N}\), annihilation
-\(\phi\) and creation \(\phi^\dagger\) , respectivelly.  The central
-charge is equivalent to the vaccum projection \(\pi =
-\ket{\bm{0}}\bra{\bm{0}}\). Therefore, the elements
+Note that these elements have finitely many non-zero entries,
+including diagonal terms, upper triangular terms, and lower triangular
+terms. These are equivalent to the number operators \(\mathcal{N}\),
+annihilation operators \(\phi\), and creation operators
+\(\phi^\dagger\), respectively. The central charge is equivalent to
+the vacuum projection \(\pi = \ket{\bm{0}}\bra{\bm{0}}\). Therefore,
 \begin{equation}
@@
 \end{equation}
-where \(J_n \) is written, in terms of free fermions, as \(J_n =
+where \(J_n \) is written in terms of free fermions as \(J_n =
 \sum_{j\in \mathbb{Z}} \bm{\colon} \psi_j \psi_{j+n}^\ast
-\bm{\colon}\),
-These set of operators generate a Heisenberg subalgebra
-\(\widehat{\mathfrak{gl}}(1) \subset \mathfrak{gl}(\infty)\) given by
+\bm{\colon}\). These set of operators generate a Heisenberg subalgebra
+\(\widehat{\mathfrak{gl}}(1) \subset \mathfrak{gl}(\infty)\)
 \begin{equation}
@@
 \begin{equation}
+\label{eq:tau-corr}
 \begin{split}
-  \tau_0(\bm{x}, \bm{y}) & = \bra{\Psi(\bm{x})} \mathcal{G} \ket{\Psi(\bm{y})} \\
-  \\ & = \bra{\bm{0}} \prod_{i} \mathbb{C}(x_i)
+  \tau(\bm{x}, \bm{y}) & = \bra{\Psi(\bm{x})} \mathcal{G} \ket{\Psi(\bm{y})} \\
+  & = \bra{\bm{0}} \prod_{i} \mathbb{C}(x_i)
   \mathcal{G} \prod_j \mathbb{B}(y_j) \ket{\bm{0}}\; ,
@@
 
-\subsubsection{Schur polynomials expansion of some correlation functions}
-In fact, it is worth noticing that other quantities, that are not tau
-functions, might have interesting Schur polynomials expansion.
-Consider the state calculated in~\cite{Bogoliubov2005}
-\begin{equation}
-  \ket{\mathcal{Y}} =
- \sum_{\vec{n}} \prod_{j=0}^M \ket{n_j} = \sum_\mu \ket{\mu} \qquad \sum_j n_j = N\; .
-\end{equation}
-Then
-\begin{equation}
- \bra{\mathcal{Y}} \prod_{j=1}^N B(\bm{x}) \ket{\nu} = 
-  \sum_{\lambda\subseteq (N,N,M)} s_{\lambda/\nu}(\bm{x})\; . 
-\end{equation}
-And it is just a sum of skew Schur polynomials.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\subsubsection{Hypergeometric tau functions}
+From these expressions, we can conclude that if we consider a diagonal
+group element \(\mathcal{G} = \exp \left( \sum_{i\geq 0} c_i
+\mathcal{N}_i\right)\), we have that~(\ref{eq:tau-corr}) becomes
+\begin{equation}
+\begin{split}
+  \tau(\bm{x}, \bm{y}) 
+  & = \bra{\bm{0}} \prod_{i} \mathbb{C}(x_i)
+  e^{\sum_i c_i \mathcal{N}_i} \prod_j \mathbb{B}(y_j) \ket{\bm{0}}\\
+  & = \sum_{\mu, \nu \subseteq [N,M]} c_{\mu\nu} s_\mu(\bm{x}) s_\nu(\bm{x})
+\end{split}
+\end{equation}
+where
+\begin{equation}
+  c_{\mu \nu} = \bra{\mu} e^{\sum_i c_i \mathcal{N}_i}\ket{\nu} \; .
+\end{equation}
+From the representation of the phase model, we have that for the partition
+\(\mu = (1^{n_1} 2^{n_2} \dots M^{n_M})\), we have
+\begin{equation}
+  \mathcal{N}_i \ket{\mu} = n_i \ket{\mu} \qquad n_0 = N - \ell(\mu)\; .
+\end{equation}
+
+Consequently, we have
+\begin{equation}
+  c_{\mu \nu} \equiv \delta_{\mu\nu} c_{\mu}  =  \delta_{\mu\nu} \prod_i e^{c_i n_i}\; , 
+\end{equation}
+and we conclude that 
+\begin{equation}
+  \tau(\bm{x}, \bm{y}) 
+  = \sum_{\mu, \nu \subseteq [N,M]} c_{\mu} s_\mu(\bm{x}) s_\mu(\bm{x})\; ,
+\end{equation}
+and we see that this tau function has diagonal coordinates
+\(c_\lambda\).  These tau functions are of the hypergeometric type,
+considered in~\cite{Orlov:2000, orlov:2001, Orlov:2001b, Orlov:2005}.
 

### 2024-04-24 draft.tex
start improving the q-bosons model

@@
  \bra{\mathcal{Y}} \prod_{j=1}^N B(\bm{x}) \ket{\nu} = 
-  \sum_{\lambda\subseteq (N,N,M)} s_{\lambda/\nu}(\bm{x})\; . 
+  \sum_{\lambda\subseteq [N,M]} s_{\lambda/\nu}(\bm{x})\; . 
 \end{equation}
@@
 \end{equation}
-and we see that this tau function has diagonal coordinates
-\(c_\lambda\).  These tau functions are of the hypergeometric type,
+We observe that this tau function has diagonal coordinates
+\(c_\lambda\). These tau functions belong to the hypergeometric type
 considered in~\cite{Orlov:2000, orlov:2001, Orlov:2001b, Orlov:2005}.
@@
 
-We now move out attention to the case of the q-bosons. The 
-analysis is parallel to what we did before, but the  specific 
-details are very different. We start with the study of the norm of  
-two off-shell Bethe states.
-
-It has been shown~\cite{Tsilevich:2006} conf. \cite{Sulkowski:2008mx,
-  Wheeler:2010vmq}, that the Bethe states \(\ket{\Psi(\bm{x})}\) have 
-coordiante expansion given by
-\begin{equation}
-  \ket{\psi(\bm{x})} = \sum_{\mu \subset [N,M]} P_\mu(\bm{x}, Q) \ket{\mu}  \qquad 
-  \bra{\psi(\bm{x})} = \sum_{\mu \subset [N,M]} P_\mu(\bm{x}, Q) \bra{\mu}\; ,
-\end{equation}
-where \(P_\lambda\) denote the Hall-Littlewood polynomials.  In this
-form, scalar product of two off-shell Bethe states in the q-boson
-model is easily calculated to be
+We now shift our focus to the case of q-bosons. The analysis parallels
+what we did before, but the specific details are markedly
+different. We begin by examining the norm of two off-shell Bethe
+states.
+
+It has been shown~\cite{Tsilevich:2006} (see
+also~\cite{Sulkowski:2008mx, Wheeler:2010vmq}) that the Bethe states
+\(\ket{\Psi(\bm{x})}\) have coordinate expansions
+\begin{equation}
+  \ket{\Psi(\bm{x})} = \sum_{\mu \subseteq [N,M]} P_\mu(\bm{x}, Q) \ket{\mu}_Q  \qquad 
+  \bra{\Psi(\bm{x})} = \sum_{\mu \subseteq [N,M]} P_\mu(\bm{x}, Q) \prescript{}{Q}{\bra{\mu}}\; ,
+\end{equation}
+where \(P_\lambda\) denote Hall-Littlewood polynomials. In this form,
+the scalar product of two off-shell Bethe states in the q-boson model
+can be easily calculated to be
 \begin{equation}
 \label{eq:scalar-hl}
-\mathcal{I}_Q(N,M | \bm{x}, \bm{y}) = \bra{\bm{0}} \prod_{j=1}^N C(x_j, Q)
-\prod_{k=1}^N B(y_k, Q) \ket{\bm{0}}
-= \sum_{\lambda \subseteq
-  (N,N,M)} b_\lambda(Q) P_{\lambda}(\bm{x}, Q) P_{\lambda}(\bm{y}, Q)
+\mathcal{I}_Q(N,M | \bm{x}, \bm{y}) = \bra{\bm{0}} \prod_{j=1}^N \mathbb{C}(x_j, Q)
+\prod_{k=1}^N \mathbb{B}(y_k, Q) \ket{\bm{0}}
+= \sum_{\lambda \subseteq [N,M]} b_\lambda(Q) P_{\lambda}(\bm{x}, Q) P_{\lambda}(\bm{y}, Q)
 \end{equation}
 where we use that \(\bracket{\lambda}{\mu} = b_\lambda(Q)
-\delta_{\lambda, \mu}\).  It turns out that this expansion is also
-related to the Cauchy identity for Hall-Littlewood
-polynomials~\cite{Macdonald:1998}
+\delta_{\lambda, \mu}\) and the completeness relation
+\begin{equation}
+  \sum_{\lambda} \frac{1}{b_\lambda(Q)} \ket{\lambda}_Q {}_Q\bra{\lambda}  = \mathbb{1}\; .
+\end{equation}
+
+It turns out that this expansion is the Cauchy identity for
+Hall-Littlewood polynomials~\cite{Macdonald:1998}
 \begin{equation}
@@
 
-Ourt goal is to consider some elementary properties of Hall-Littlewood polynomials 
-to understand aspects of this expansion.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\subsection{Determinant formula}
-
-We would like to refine the expressions above.  Let us first consider
-a determinant expression for the scalar product~(\ref{eq:scalar-hl}).
-Let us continue with the case where \(M\) and \(N\) are very large,
-but finite. In this case, we write this scalar product as
-\begin{equation}
+Our goal is to explore some properties of Hall-Littlewood polynomials
+to gain insight into aspects of this expansion.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\subsection{Scalar product: determinant formula}
+We aim to refine the expressions above. First, we consider a
+determinant expression for the scalar product~(\ref{eq:scalar-hl}). We
+proceed with the scenario where \(M\) and \(N\) are very large but
+finite. In this case, we express this scalar product as
+\begin{equation}
+\label{eq:q-inner}
   \mathcal{I}_Q(N,M | \bm{x}, \bm{y})  
@@
 \end{equation}
-and using the results we have derives in the Phase model, we have that 
+and from the results derived in the phase model, we find that
 \begin{equation}
@@
 \end{equation}
-where \(H\) is the matrix~(\ref{eq:h-matrix}). Therefore
+where \(H\) is the matrix~(\ref{eq:h-matrix}). Consequently, the scalar
+product~(\ref{eq:q-inner}) becomes
 \begin{equation}
@@
 \end{equation}
-and we see that the scalar product of the q-boson model is the quotient of 
-scalar products of the phase model. Observe that in the case \(Q = 0\), we have 
-\(\mathcal{I}(N, M| \bm{x}, \bm{0}) = 1\), therefore
-\(\mathcal{I}_0(N,M | \bm{x}, \bm{y}) = \mathcal{I}(N,M | \bm{x}, \bm{y})\), 
-as expected.
-
-Moreover
+and we see that it is the quotient of scalar products of the phase
+model. Hence, it is the quotient of two Toda tau functions.
+Observe that in the case \(Q = 0\), we have \(\mathcal{I}(N, M|
+\bm{x}, \bm{0}) = 1\), then \(\mathcal{I}_0(N,M | \bm{x}, \bm{y})
+= \mathcal{I}(N,M | \bm{x}, \bm{y})\), as expected.
+
+Additionally, we write
 \begin{equation}
@@
 \end{equation}
-Using that \(\Delta(Q\bm{y})/ \Delta(\bm{y}) = \prod_{j=1}^{N-1} Q^j\), we have 
+Using \(\Delta(Q\bm{y})/ \Delta(\bm{y}) = \prod_{j=1}^{N-1} Q^j\), we have 
 \begin{equation}
@@
 \end{equation}
-Observe that from this expression, we cannot decompose this function
-as in~(\ref{eq:h-exp}) since the coefficients in this expansion are 
-are dependent on \(\bm{x}\) and \(\bm{y}\).
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\subsection{Scalar product: Big Schur functions}
+Note that from this expression, we cannot decompose this function as
+in~(\ref{eq:h-exp}) since the coefficients in this expansion depend on
+\(\bm{x}\) and \(\bm{y}\).
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\subsection{Scalar product: big Schur functions}
 
@@
   \mathcal{I}_Q(N,M | \bm{x}, \bm{y})
-  = \sum_{\lambda \subseteq (N,N,M)} \mathcal{S}_{\lambda}(\bm{x}, Q) s_{\lambda}(\bm{y})
-  = \sum_{\lambda \subseteq (N,N,M)} \mathcal{S}_{\lambda}(\bm{y}, Q) s_{\lambda}(\bm{x})\; ,
+  = \sum_{\lambda \subseteq [N,M]} \mathcal{S}_{\lambda}(\bm{x}, Q) s_{\lambda}(\bm{y})
+  = \sum_{\lambda \subseteq [N,M]} \mathcal{S}_{\lambda}(\bm{y}, Q) s_{\lambda}(\bm{x})\; ,
 \end{equation}

### 2024-04-26 draft.tex
finish draft

@@
 
-\subsection{Scalar product: big Schur functions}
-
-Let us now recover the result originally derived
-in~\cite{Foda:2008hn} that shows how the scalar
-product~(\ref{eq:scalar-hl}) is a tau function of the KP hierarchy.
-From~\cite[see chap.3, sec. 4, eq. (4.7)]{Macdonald:1998}, we can
-expand the Cauchy identity~(\ref{eq:cauchy_hl}) as
+\subsection{Big Schur functions expansion}
+Now, let's revisit the result originally derived in~\cite{Foda:2008hn}
+which demonstrates that the scalar product~(\ref{eq:scalar-hl}) is a
+tau function of the KP hierarchy. According to~\cite[Chapter 3,
+  Section 4, Equation (4.7)]{Macdonald:1998}, we can expand the Cauchy
+identity~(\ref{eq:cauchy_hl}) as
 \begin{equation}
@@
 \sum_{\lambda} \mathcal{S}_{\lambda}(\bm{x}, Q) s_{\lambda}(\bm{y}) =
-\sum_{\lambda} \mathcal{S}_{\lambda}(\bm{y}, Q) s_{\lambda}(\bm{x}) \; ,
-\end{equation}
-where the polynomials \(\mathcal{S}_\lambda\), the big-Schur
-functions, are given by a Jacobi-Trudi formula
+\sum_{\lambda} \mathcal{S}_{\lambda}(\bm{y}, Q) s_{\lambda}(\bm{x}) \; .
+\end{equation}
+Here, the polynomials \(\mathcal{S}_\lambda\), known as the big-Schur
+functions, are defined by a Jacobi-Trudi formula:
 \begin{equation}
@@
 \end{equation}
-and the coefficients \(q_m\) can be obtained from the following expression
+and the coefficients \(q_m\) are obtained from the expression:
 \begin{equation}
  \sum_{m} q_m(\bm{y}, Q) z^m =
- \prod_{j} \frac{1-Q y_j z}{1 - y_j z}
+ \prod_{j} \frac{1-Q y_j z}{1 - y_j z}\; .
 \end{equation}
@@
 \begin{equation}
+\label{eq:big-expansion}
   \mathcal{I}_Q(N,M | \bm{x}, \bm{y})
@@
 
+Based on these findings, we conclude that the inner product can be
+expressed as a quotient of two Toda tau functions or as a KP tau
+function with coefficients given by the big Schur
+polynomials. However, further exploration of these results is
+warranted.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \subsection{Kostka-Foulkes expansion}
-Let us expand these polynomials in a Schur polynomials basis, that is
+In a Schur polynomial basis, we write
 \begin{equation}
@@
 
-Expanding now the big-Schur functions as
+If we now insert this expansion in the inner product, we have 
+\begin{equation}
+    \sum_\lambda b_\lambda(Q) P_\lambda(\bm{x}, Q) P_\lambda(\bm{y}, Q) = 
+    \sum_{\mu,\nu} \left(\sum_\lambda K_{\mu \lambda}^{-1 \; T} b_\lambda(Q)  K_{\lambda \nu}^{-1} \right)
+    s_{\mu}(\bm{x}) s_{\nu}(\bm{y}) 
+\end{equation}
+and we have the coefficients 
+\begin{equation}
+\label{eq:c-expan}
+ \tilde{c}_{\mu\nu}(Q) = \sum_\lambda K_{\mu \lambda}^{-1 \; T} b_\lambda(Q)  K_{\lambda \nu}^{-1} \; ,
+\end{equation}
+for the double expansion of the inner product in terms of Schur
+polynomials. Based on our discussion so far, we know that these
+coefficients are not Plücker coordinates of the Toda hierarchy.
+
+But we can say something interesting about these coefficients. Let us
+now expand the big-Schur functions as
 \begin{equation}
 \label{eq:kf-exp}
-  \mathcal{S}_\lambda(\bm{x}, Q) = \sum_{\mu} c_{\lambda\mu}(Q) s_\mu(\bm{x})\; ,
-\end{equation}
-we can fix the coefficients \(c_{\lambda\mu}\) in terms of Kostka-Foulkes
-polynomials using the orthogonality relations of these polynomials~\cite{Macdonald:1998}. 
-In particular, there is an inner product in the ring of symmetric functions such that
-\begin{equation}
-  \frac{1}{b_\mu}\langle P_\mu(\bm{x}, Q), P_\mu(\bm{x}, Q)\rangle = \langle
-  \mathcal{S}_\mu(\bm{x}, Q), s_\mu(\bm{x})\rangle = \delta_{\mu\nu}\; .
+  \mathcal{S}_\mu(\bm{x}, Q) = \sum_{\lambda} \mathcal{C}^{\mu}_\lambda(Q) s_\lambda(\bm{x})\; ,
+\end{equation}
+we can fix the coefficients \(\mathcal{C}_\lambda^\mu(Q)\) in terms of
+Kostka-Foulkes polynomials using the orthogonality relations of these
+polynomials~\cite{Macdonald:1998}. In particular, there is an inner
+product in the ring of symmetric functions such that
+\begin{equation}
+  b_\mu \langle P_\mu(\bm{x}, Q), P_\nu(\bm{x}, Q)\rangle = \langle
+  \mathcal{S}_\mu(\bm{x}, Q), s_\nu(\bm{x})\rangle = \delta_{\mu\nu}\; .
 \end{equation}
@@
   \begin{split}
-\delta_{\mu\nu} & = \sum_\lambda c_{\mu\lambda}  \langle s_\lambda(\bm{x}), s_\mu(\bm{x})\rangle\\
-& = \sum_\lambda \sum_{\pi_1, \pi_2} c_{\mu\lambda}  K_{\lambda \pi_1} K_{\mu \pi_2}
-\langle P_{\pi_1}(\bm{x}, Q), P_{\pi_2}(\bm{x}, Q)\rangle \\ 
-& = \sum_\lambda \sum_{\pi} c_{\mu\lambda} b_{\mu}^{-1}  K_{\lambda \pi} K_{\mu \pi}\; ,
+\delta_{\mu\nu} & = \sum_\lambda \mathcal{C}_\lambda^\mu \langle s_\lambda(\bm{x}), s_\nu(\bm{x})\rangle
+ = \sum_\lambda \sum_{\pi, \sigma} \mathcal{C}_\lambda^\mu K_{\lambda \pi} K_{\nu \sigma}
+\langle P_{\pi}(\bm{x}, Q), P_{\sigma}(\bm{x}, Q)\rangle \\ 
+& = \sum_\lambda \mathcal{C}_\lambda^\mu \sum_{\sigma}K_{\lambda \sigma} b_{\sigma}^{-1} K_{\nu \sigma}\; ,
   \end{split}
@@
 \label{eq:indices}
-c_{\mu \nu} = \sum_\pi K_{\mu\pi}^{-1} b_\mu (K_{\nu\pi}^{-1})^T\; .
-\end{equation}
-\marginpar[left]{\tiny Fix these \\ indices later.}
+(\mathcal{C}^{-1})^\lambda_\nu = \sum_\sigma K_{\lambda\sigma} b^{-1}_\sigma (K_{\sigma\nu})^T\; ,
+\end{equation}
+and consistency with~(\ref{eq:c-expan}) implies that
+\(\mathcal{C}^\lambda_\nu \equiv \tilde{c}_{\lambda\nu}\).
 
@@
 \begin{equation}
-\sum_{\mu} S_{\mu}(\bm{x},Q) s_{\nu}(\bm{y}) = 
-\sum_{\mu , \nu} c_{\mu\nu} s_{\mu}(\bm{x}) s_{\nu}(\bm{y}) 
-\end{equation}
-and everything boils down to the analysis of the
-indices~(\ref{eq:indices}). In has been shown
-in~\cite{Necoechea:2019wbg} that the big Schur functions themselves
-are KP-tau functions\footnote{ We give a simple argument for this fact
-in the remark 2 below.} therefore, the coefficients \((c_\lambda)_\mu
-\equiv c_{\lambda\mu}\) must satisfy Plücker relations for the KP
-hierarchy.
+S_{\mu}(\bm{t},Q) = \sum_{\nu} \tilde{c}_{\mu\nu}(Q) s_{\mu}(\bm{t})\; ,
+\end{equation}
+where we have written the big Schur polynomial in terms of the Miwa
+coordinates \(\bm{t}\).
 
 \begin{remark}
- This expression show somethign interesting. We can formulate this
-problem in terms of partitions defined in the phase
-model. Equivalently, we can use the usual vertex operator
-construction, and not the q-deformed version of Jing~\emph{\cite{Jing1991,
-  Jing1995}}, that is much harder to deal with.
+This expression also reveals something interesting. We can formulate
+this problem in terms of partitions defined in the phase
+model. Alternatively, we can utilize the conventional vertex operator
+construction rather than the q-deformed version proposed by
+Jing~\cite{Jing1991, Jing1995}, which is much more challenging to
+handle.
 \end{remark}
 
-If these coefficients satisfy the Toda hierarchy Plücker relations, then 
-one can find a \(Q\)-dependent group element \(G_Q\in GL(\infty)\) such that 
-\begin{equation}
-  c_{\mu\nu} = \bra{\mu} G_Q \ket{\nu}\; ,
-\end{equation}
-where \(\ket{\mu}\) and \(\ket{\nu}\) are phase model
-states. Unfortunatelly, we could not find any good reason to explore
-this expansion in a consistent way. In what follows, we try to explore
-alternative expansions that give a double Schur expansion.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\subsection{Scalar product: Supersymmetric Schur polynomials}
-
-With this result, we now argue that~(\ref{eq:scalar-hl}) is also a tau
-function of the Toda hierarchy in disguise. Let us decompose
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\subsection{Supersymmetric Schur polynomials expansion}
+With this result, we now argue that although~(\ref{eq:scalar-hl}) is
+not a Toda tau function with respect to the coordinates \(\{\bm{t};
+\bm{t}'\}\), we can define a new set of coordinates
+\(\{\bm{t};\bm{T}\}\) that render the inner product a trivial Toda tau
+function.
+
+Let us first decompose
 \begin{equation}
@@
 \end{equation}
-And reorganizing this sum, we conclude that 
+Reorganizing this sum, we conclude that
 \begin{equation}
@@
 \end{equation}
-where in the last equality we have used the homogeneity of the elementary symmetric polynomials,
-and \( -Q\bm{y} = (-Qy_1, -Qy_2, \dots)\).
-
-It is easier to consider the following expansions 
+where in the last equality we have used the homogeneity of the
+elementary symmetric polynomials, and \( Q\bm{y} \equiv (Qy_1, Qy_2,
+\dots)\).
+
+It is easy to see that
 \begin{subequations}
@@
   \begin{split}
-    \prod_{j=1}^N (1 - (Q y_i)z) & = \prod_{j=1}^N e^{ \ln  (1 - (Q y_i)z)} = 
+    \prod_{j=1}^N (1 - (Q y_j)z) & = \prod_{j=1}^N e^{ \ln  (1 - (Q y_j)z)} = 
     \prod_{j=1}^N \exp \left( - \sum_{n>0} \frac{Q^n y^n_j}{n} z^n \right) \\ 
@@
 \end{equation}
-where \(t^{(Q)}_n = \frac{Q^n}{n} \sum_j y_j^n = Q^n t'_n\), and 
-\begin{equation}
-    \prod_{j=1}^N \frac{1}{(1 - y_iz)} = \prod_{j=1}^N e^{- \ln  (1 - y_iz)} = 
+where \(t^{(Q)}_n = \frac{Q^n}{n} \sum_j y_j^n = Q^n t'_n\). Additionally
+\begin{equation}
+    \prod_{j=1}^N \frac{1}{(1 - y_jz)} = \prod_{j=1}^N e^{- \ln  (1 - y_jz)} = 
     \prod_{j=1}^N \exp \left( \sum_{n>0} \frac{y_j}{n} z^n \right)=
@@
 
-Therefore 
+Consequently
 \begin{equation}
  \sum_{m} q_m(\bm{y}, Q) z^m =
- \prod_j \frac{1 + y_j (-Q z)}{1 - y_k z}
+ \prod_j \frac{1 - Q y_j z}{1 - y_k z}
  = \exp \left( \sum_{n\geq 1} (t'_n - t_n^{(Q)})z^n \right)
@@
 All in all, we conclude that the big Schur functions
-\(S_\lambda(\bm{t}', Q)\) are ordinary Schur functions with respect
+\(S_\lambda(\bm{t}', Q)\) are ordinary Schur functions with respect to
 the coordinates \(\bm{T}\).
 
-In fact, we can do even better. Supersymmetric (or Hook) Schur
-functions~\cite{Berele:1983}, see also see~\cite{Macdonald:1998, Moens:2003},
-\(s_\lambda(\bm{\alpha}/\bm{\beta})\) are defined as ordinary schur
-functions evaluated at Miwa coodinates of the form
+Moreover, a more refined approach is also possible. Supersymmetric (or
+Hook) Schur functions~\cite{Berele:1983}, as discussed in works such
+as~\cite{Macdonald:1998, Moens:2003}, denoted by
+\(s_\lambda(\bm{\alpha}/\bm{\beta})\), are defined as ordinary Schur
+functions evaluated at Miwa coordinates of the form
 \begin{equation}
   T_n = \frac{1}{n} \left( \sum_{i=1}^{\dim(\alpha)} \alpha_i^n
-       - \sum_{i=1}^{\dim(\beta)} (-\beta_i)^n\right)\; ,
-\end{equation}
-and comparing with the results above, we can see that the big Schur functions
-\(S_\lambda(\bm{t}', Q)\)\; , is the supersymmetric Schur functions for \(\bm{\alpha} = \bm{y}\) 
-and \(\bm{\beta} = - Q \bm{y}\), that is 
+       - \sum_{i=1}^{\dim(\beta)} (-\beta_i)^n\right)\; .
+\end{equation}
+Comparing this expression with the results above, we can see that the
+big Schur functions \(S_\lambda(\bm{t}', Q)\) correspond to the
+supersymmetric Schur functions for \(\bm{\alpha} = \bm{y}\) and
+\(\bm{\beta} = - Q \bm{y}\), which is
 \begin{equation}
@@
 
-Putting these facts together, we immediately conclude that since 
-\begin{equation}
-  \sum_\lambda s_\lambda(\bm{T}) s_\lambda (\bm{t}) \; ,
-\end{equation}
-is a Toda hierarchy tau function, we obviously have that 
-\begin{equation}
-  S_Q(N,M| \bm{x}, \bm{y}) \equiv S_Q(N,M| \bm{t}, \bm{T})
-  = \sum_{\lambda \subseteq [N,N,M]} s_\lambda(\bm{T}) s_\lambda(\bm{t})\; , 
+Putting these facts together, we immediately conclude that
+\begin{equation}
+\lim_{M,N\to \infty}\mathcal{I}_Q(N,M|\bm{x}, \bm{y}) = 
+  \sum_\lambda s_\lambda(\bm{T}) s_\lambda (\bm{t})
+\end{equation}
+is a Toda hierarchy tau function with respect to
+\(\{\bm{t};\bm{T}\}\). As a result, we obviously have that
+\begin{equation}
+  \mathcal{I}_Q(N,M| \bm{x}, \bm{y}) \equiv \mathcal{I}_Q(N,M| \bm{t}, \bm{T})
+  = \sum_{\lambda \subseteq [N,M]} s_\lambda(\bm{T}) s_\lambda(\bm{t})
 \end{equation}
@@
 
-Now we can consider an expansion that differs from
-from~(\ref{eq:kf-exp}). Let us consider the supersymmetric Schur
-polynomials in terms of ordinary Schur functions as~\cite[Sec. I.5,
-  exerc. 23]{Macdonald:1998}
+Let us also express the supersymmetric Schur polynomials in terms of
+ordinary Schur functions, as shown in~\cite[Sec. I.5, exerc. 23]{Macdonald:1998}:
 \begin{equation}
@@
 \end{equation}
-where the prime denotes the conjugate diagram. Then
+where the prime denotes the conjugate diagram. Compare this expansion
+with~(\ref{eq:kf-exp}). Then
 \begin{equation}
   \label{eq:tau-hl}
-  \sum_{\lambda } s_\lambda(\bm{y}/ (- Q\bm{y})) s_\lambda(\bm{x}) =
-  \sum_{\lambda,\mu } s_{(\lambda/\mu)'}(- Q\bm{y}) s_\lambda(\bm{x}) s_\mu(\bm{y})\; .
+  \sum_{\lambda } s_\lambda(\bm{T}) s_\lambda(\bm{t}) =
+  \sum_{\lambda,\mu } s_{(\lambda/\mu)'}(- Q\bm{y}) s_\lambda(\bm{t}) s_\mu(\bm{t}')\; .
 \end{equation}
@@
 \begin{remark}
-Since the skew Schur polynomials have determinant expressions,
-therefore we have that the \(\bm{y}\)-dependent coefficients \(c_{\mu
-  \lambda} = s_{(\lambda/\mu)'}(-Q\bm{y})\) have Jacobi-Trudi
-expressions. It is tempting to think of these objects as
-\(\bm{y}\)-dependent Pl\"ucker coordinates. In this sense, we would have a curve in 
-the indinite Grassmannian, instead a point.
+Since the skew Schur polynomials have determinant expressions, we can
+deduce that the \(\bm{y}\)-dependent coefficients \(c_{\mu \lambda} =
+s_{(\lambda/\mu)'}(-Q\bm{y})\) have Jacobi-Trudi expressions. It is
+tempting to regard these objects as \(\bm{y}\)-dependent Plücker
+coordinates. In this sense, we would have a curve in the infinite
+Grassmannian instead of a point.
 \end{remark}
@@
 \begin{remark}
-Remeber that one of the easiest nontrivial solutions of the KP
-hierarchies are the Schur functions
-themselves~\emph{\cite{Zabrodin2018}}. We have just seen that the
-big-Schur polynomials can be written as supersymmetric Shur
-polynomials, and these are nothing but ordinary Schur polynomials in a
-particular choice of Miwa coordiantes. Then, we also see that 
-the big Schur polynomials are also KP tau functions. This direct conclusion 
-is an alternative proof of this fact, see~\emph{\cite{Necoechea:2019wbg}}.
+It is worth noting that one of the simplest nontrivial solutions of
+the KP hierarchies are the Schur functions
+themselves~\emph{\cite{Zabrodin2018}}. As we have just demonstrated, the big
+Schur polynomials can be expressed as supersymmetric Schur
+polynomials, which are essentially ordinary Schur polynomials in a
+specific choice of Miwa coordinates. Therefore, we conclude that the
+big Schur polynomials are also KP tau functions. This direct
+conclusion serves as an alternative proof of this fact, originally derived in
+in~\emph{\cite{Necoechea:2019wbg}} using the KP bilinear identity.
 \end{remark}
 
-Putting these facts together, we have that the left-hand side
-of~(\ref{eq:tau-hl}) is a tau function with respect the coordinates
-\(\bm{T}\) and \(\bm{t}\). Furthermore, from the Vertex operator
-formalism, we write the Schur polynomials as
-\begin{equation}
-  s_{\lambda}(\bm{x}) = \bra{\lambda} e^{J_-}\ket{\bm{0}}\qquad
-  s_{\mu}(\bm{y}) = \bra{\bm{0}} e^{J_+}\ket{\mu}\; ,
+Combining these observations, we conclude that the left-hand side of
+equation (\ref{eq:tau-hl}) serves as a tau function with respect to
+the coordinates \(\{\bm{t}; \bm{T}\}\). Additionally, utilizing the
+vertex operator formalism, we can express the Schur polynomials as
+\begin{equation}
+  s_{\lambda}(\bm{t}') = \bra{\lambda} e^{J_-(\bm{t}')}\ket{\bm{0}}\qquad
+  s_{\mu}(\bm{t}) = \bra{\bm{0}} e^{J_+(\bm{t})}\ket{\mu}\; ,
 \end{equation}
@@
 \begin{equation}
-  \sum_{\lambda,\mu } s_\mu(\bm{y})c_{\mu\lambda}(Q,\bm{y}) s_\lambda(\bm{x}) = 
-  \sum_{\lambda,\mu } \bra{\bm{0}} e^{J_+}\ket{\mu} c_{\mu\lambda}(Q,\bm{y})
-  \bra{\lambda} e^{J_-}\ket{\bm{0}} \; ,
-\end{equation}
-therefore 
+  \sum_{\lambda,\mu } c_{\mu\lambda}(Q,\bm{y}) s_\mu(\bm{t}') s_\lambda(\bm{t}) = 
+  \sum_{\lambda,\mu } \bra{\bm{0}} e^{J_+(\bm{t})}\ket{\mu} c_{\mu\lambda}(Q,\bm{y})
+  \bra{\lambda} e^{J_-(\bm{t}')}\ket{\bm{0}} \; .
+\end{equation}
+It is important to emphasize once again that this expression does not
+constitute a tau function with respect to \(\{\bm{t},
+\bm{t}'\}\). Despite this crucial distinction, we proceed under the
+assumption that the coefficients can be expressed as
 \begin{equation}
@@
 \end{equation}
-where \(G_Q(\bm{y}) \in GL(\infty)\). We can find the expression for
-this group element as follows.
-
-Remember that \(T_m = t'_m - Q^m t'_m\), therefore we write
+where \(G_Q(\bm{y}) \in GL(\infty)\). Hence
+\begin{equation}
+  \lim_{N,M\to\infty} \mathcal{I}_Q(N, M| \bm{t}, \bm{t}'(\bm{y})) = 
+  \bra{\bm{0}} e^{J_+(\bm{t})} G_Q(\bm{y}) e^{J_-(\bm{t}')}\ket{\bm{0}}\; .
+\end{equation}
+where \(t'_n = \frac{1}{n}\sum_i y_i^n\).
+
+To derive the expression for this group element, recall that \(T_m =
+t'_m - Q^m t'_m\). With this in mind, we can express the scalar
+product as follows:
 \begin{subequations}
@@
 \end{equation}
-From the Heisenberg algebra, we conclude that \([J_-(\bm{t}),
+From the Heisenberg algebra, we deduce that \([J_-(\bm{t}),
   J_-(\bm{t}')] = 0\), therefore
@@
   \bra{\bm{0}} e^{J_+(\bm{t})} e^{J_-(\bm{T})} \ket{\bm{0}} = 
-\bra{\bm{0}} e^{J_+(\bm{t})} e^{-J_-(\bm{t}^{(Q)})} e^{J_-(\bm{t}')}  \ket{\bm{0}} \; ,
+\bra{\bm{0}} e^{J_+(\bm{t})} e^{-J_-(\bm{t}^{(Q)})} e^{J_-(\bm{t}')}  \ket{\bm{0}} \; .
 \end{equation}
 \end{subequations}
-from where we finally conclude that 
+We finally conclude that 
 \begin{equation}
@@
 
-In general, the expectation values with coordinate dependent group
-elements \(G_{Q}(\bm{y})\) are not Toda hierarchy tau functions. But
-in the above example, the coordinates combine among themselves and
-generate a tau function with respect the coordiantes \(\bm{T}\) and
-\(\bm{t}\).
+Generally, the expectation values involving coordinate-dependent group
+elements \(G_{Q}(\bm{y})\) do not form Toda hierarchy tau
+functions. However, in the example above, the coordinates combine in
+such a way that they generate a tau function with respect to the
+coordinates \(\bm{T}\) and \(\bm{t}\).
 
@@
 
-Let us now consider the fermionic construction of this tau function. We write 
-the limit \(N,M\to \infty\) of the scalar product
-\begin{subequations}
-\begin{equation}
-  S_Q(\bm{T}, \bm{t}') = \sum_{\mu} s_\mu(\bm{T}) s_\mu(\bm{t}')
-\end{equation}
-that is quite quite trivial from the viewpoint of the free fermion construction, 
-\begin{equation}
-  S_Q(\bm{T}, \bm{t}') = \bra{\bm{0}} e^{\bm{J}_+(\bm{T})} e^{\bm{J}_-(\bm{t}')} \ket{\bm{0}}\; .
-\end{equation}
-
-Moreover, we also know that
-\begin{equation}
-  S_Q(\bm{t}, \bm{t}') = \sum_{\mu\nu} c_{\mu\nu} s_\mu(\bm{t}) s_\mu(\bm{t}')
-\end{equation}
-where \(n t = \sum_j x^n_j\) and \(n t' = \sum_j y^n_j\) are Miwa coordinates, then  
-\begin{equation}
-  S_Q(\bm{t}, \bm{t}') = \bra{\bm{0}} e^{J_+(\bm{t})} G_Q  e^{J_-(\bm{t}')} \ket{\bm{0}}
-\end{equation}
-and \(G_Q \in GL(\infty)\) that depends on \(Q\). We could also think of \(G_Q\)
-as a twist
-\(G_Q = e^{Q H} G e^{-Q H}\), with the action of \(H\in \mathfrak{gl}(\infty)\)
-on ordinary fermions generating the Q-fermions. 
-\marginpar[left]{\tiny Investigate this\\ twists}
-\end{subequations}
-
-It seems that we have some different representations for this tau function, 
-and it is worth investigating what is happening here. That is what we want 
-to do now.
+%Let us now consider the fermionic construction of this tau function. We write 
+%the limit \(N,M\to \infty\) of the scalar product
+%\begin{subequations}
+%\begin{equation}
+%  S_Q(\bm{T}, \bm{t}') = \sum_{\mu} s_\mu(\bm{T}) s_\mu(\bm{t}')
+%\end{equation}
+%that is quite quite trivial from the viewpoint of the free fermion construction, 
+%\begin{equation}
+%  S_Q(\bm{T}, \bm{t}') = \bra{\bm{0}} e^{\bm{J}_+(\bm{T})} e^{\bm{J}_-(\bm{t}')} \ket{\bm{0}}\; .
+%\end{equation}
+%
+%Moreover, we also know that
+%\begin{equation}
+%  S_Q(\bm{t}, \bm{t}') = \sum_{\mu\nu} c_{\mu\nu} s_\mu(\bm{t}) s_\mu(\bm{t}')
+%\end{equation}
+%where \(n t = \sum_j x^n_j\) and \(n t' = \sum_j y^n_j\) are Miwa coordinates, then  
+%\begin{equation}
+%  S_Q(\bm{t}, \bm{t}') = \bra{\bm{0}} e^{J_+(\bm{t})} G_Q  e^{J_-(\bm{t}')} \ket{\bm{0}}
+%\end{equation}
+%and \(G_Q \in GL(\infty)\) that depends on \(Q\). We could also think of \(G_Q\)
+%as a twist
+%\(G_Q = e^{Q H} G e^{-Q H}\), with the action of \(H\in \mathfrak{gl}(\infty)\)
+%on ordinary fermions generating the Q-fermions. 
+%\marginpar[left]{\tiny Investigate this\\ twists}
+%\end{subequations}
+%
+%It seems that we have some different representations for this tau function, 
+%and it is worth investigating what is happening here. That is what we want 
+%to do now.
 

### 2024-04-27 draft.tex
finish introduction and add definitions

@@
 
-\usepackage{/home/thiago/.config/dot-files/latex/globaldef}
-\input{/home/thiago/.config/dot-files/latex/definitions.tex}
+\usepackage{globaldef}
+\input{definitions.tex}
 
@@
 R. Dr. Bento T. Ferraz 271, Bl. II, Sao Paulo 01140-070, SP, Brazil\\
-\&
+\&\\
 Instituto de Física, Universidade de S\~ao Paulo,
@@
 %\date{\today}
-\keywords{Integrability, Bethe, Schur}
+%\keywords{Integrability, Bethe, Schur}
 \subjclass[2020]{82B20, 82B23}
@@
 %*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*
+
+The exploration of connections between quantum and classical
+exactly solvable models constitutes a significant endeavor aimed at
+elucidating the overarching structure of integrable systems.  This
+research avenue has yielded fruitful insights, as evidenced
+by~\cite{Its:1992bj, Foda:2009zz, Alexandrov:2011aa, Araujo:2021ghu},
+to name a few.  The present work is situated within this research field.
+
+Here, we examine the emergence of classical integrable structures
+within correlation functions of the q-boson system. This quantum
+integrable system characterizes strongly correlated bosons confined to
+a one-dimensional chain~\cite{Bogoliubov:1992, Bogoliubov:1997soj,
+  Bogoliubov2005}. Notably, this model bears close relation to the AL
+(Ablowitz-Ladik) model, which is an integrable discretization of the
+nonlinear Schrödinger equation.
+
+In~\cite{Bogoliubov2005}, the author demonstrates that the \(q = 0\)
+limit of the q-boson, known as the \emph{phase model}, is associated
+with the enumeration of plane partitions. This connection allows for a
+Schur polynomials expansion for the Bethe states. Subsequently, this
+result was extended to \(q > 0\) in~\cite{Tsilevich:2006}, where the
+Bethe states are expressed in terms of Hall-Littlewood polynomials.
+
+Foda and Wheeler~\cite{Foda:2008hn, Wheeler:2010vmq} utilize a rich
+set of dualities between the ring of symmetric functions and
+\(q\)-deformations of the Heisenberg algebra~\cite{Jing1991, Jing1995}
+to explore the combinatorial properties of this spin chain. As a
+consequence, it has been shown that the scalar product of off-shell
+Bethe states are restricted tau functions of the KP
+(Kadomtsev-Petviashvili) hierarchy.
+
+This underlying combinatorial structure underscores the significance
+of this model in various physical phenomena, notably in the context of
+crystal melting and string theory~\cite{Okounkov:2003sp,
+  Saulina:2004da}. The application of the q-boson model in these
+contexts has been demonstrated in~\cite{Sulkowski:2008mx}, with the
+string data recoverable as the chain length approaches infinity.
+
+This paper builds upon these developments and aims to extend some of
+these results. Firstly, we delve deeper into the connection between
+the scalar product and tau functions, exploring them in a wider
+context, particularly in the case of the Toda
+hierarchy~\cite{Takasaki:2018wsv}. It is important to note that both
+the AL and the KP hierarchies are reductions of the Toda
+hierarchy. This paper takes steps towards resolving this problem,
+which is instrumental in understanding the finer details of these
+relationships.
+
+Additionally, we explore other correlation functions within this model
+and examine how they fit into the classical integrable context. It is
+noteworthy that even when the object is not a tau function itself, it
+may have an interesting expansion in terms of different classes of
+Schur polynomials, highlighting some of its combinatorial
+properties. We also investigate some of these expansions in this
+paper.
+
+In Section 2, we provide a review of the phase model and
+q-bosons. This section serves to establish notation and emphasize the
+key aspects relevant to our analysis.  Section 3 delves into the
+analysis of the Phase Model and its connections with the KP and Toda
+integrable hierarchies. We explore the general expansion of a phase
+model correlation function that yields a Toda system solution.  In
+Section 4, we shift our focus to the q-bosons. We discuss a
+determinantal formula for the scalar product of two off-shell Bethe
+states and its expansion in terms of big, supersymmetric Schur
+functions, as well as in terms of Kostka-Foulkes polynomials.
+Finally, in Section 5, we conclude with an overview of open problems.
 
@@
 %*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*
+
+% Ablowitz-Ladik and Toda hierarchy - reduction -- understands this problem
 

### 2024-04-29 draft.tex
finish paper

@@
 \begin{abstract}
-In this work, we study aspects of the \(Q\)-boson model, and 
-its relations with integrable hierarchies. 
+This work delves into the complex interplay between the q-boson model,
+a quantum integrable system, and classical integrable systems like the
+Toda and KP hierarchies. By analyzing scalar products of Bethe states
+and correlation functions, we investigate their connections to tau
+functions and explore Schur polynomial expansions. This exploration
+unravels the intricate relationship between quantum and classical
+integrable systems.
 \end{abstract}
@@
 Schur polynomials, highlighting some of its combinatorial
-properties. We also investigate some of these expansions in this
-paper.
+properties\footnote{Numerous results involving Schur and
+Hall-Littlewood polynomials have been confirmed using the
+software~\cite{Araujo:2024piv}.}. We also investigate some of these
+expansions in this paper.
 
@@
 \begin{equation}
-  F_0(\bm{T}) = \log S_{T_0}(\bm{T}, \bm{T}') + \mathcal{O}(\hbar)\; ,
+  F_0(\bm{T}) = \log \mathcal{I}_{T_0}(\bm{T}, \bm{T}') + \mathcal{O}(\hbar)\; ,
 \end{equation}
@@
 
-% Ablowitz-Ladik and Toda hierarchy - reduction -- understands this problem
-
-%Let us now consider the fermionic construction of this tau function. We write 
-%the limit \(N,M\to \infty\) of the scalar product
-%\begin{subequations}
-%\begin{equation}
-%  S_Q(\bm{T}, \bm{t}') = \sum_{\mu} s_\mu(\bm{T}) s_\mu(\bm{t}')
-%\end{equation}
-%that is quite quite trivial from the viewpoint of the free fermion construction, 
-%\begin{equation}
-%  S_Q(\bm{T}, \bm{t}') = \bra{\bm{0}} e^{\bm{J}_+(\bm{T})} e^{\bm{J}_-(\bm{t}')} \ket{\bm{0}}\; .
-%\end{equation}
-%
-%Moreover, we also know that
-%\begin{equation}
-%  S_Q(\bm{t}, \bm{t}') = \sum_{\mu\nu} c_{\mu\nu} s_\mu(\bm{t}) s_\mu(\bm{t}')
-%\end{equation}
-%where \(n t = \sum_j x^n_j\) and \(n t' = \sum_j y^n_j\) are Miwa coordinates, then  
-%\begin{equation}
-%  S_Q(\bm{t}, \bm{t}') = \bra{\bm{0}} e^{J_+(\bm{t})} G_Q  e^{J_-(\bm{t}')} \ket{\bm{0}}
-%\end{equation}
-%and \(G_Q \in GL(\infty)\) that depends on \(Q\). We could also think of \(G_Q\)
-%as a twist
-%\(G_Q = e^{Q H} G e^{-Q H}\), with the action of \(H\in \mathfrak{gl}(\infty)\)
-%on ordinary fermions generating the Q-fermions. 
-%\marginpar[left]{\tiny Investigate this\\ twists}
-%\end{subequations}
-%
-%It seems that we have some different representations for this tau function, 
-%and it is worth investigating what is happening here. That is what we want 
-%to do now.
+In this work, we have explored the connections between a quantum
+integrable system, the q-boson model, and solutions of a classical
+integrable system, the Toda hierarchy. Our investigation has extended
+some early findings in this field and has also unveiled new research
+avenues, which we aim to explore in future studies.  Let us now delve
+into some of these promising directions.
+
+In Section 3, we delved into various facets of the phase model. We
+presented the scalar product of two off-shell Bethe states as an
+elementary example of the Toda system tau function. Additionally, we
+revealed an intriguing alternative portrayal of these scalar products
+using a matrix integral framework, which corresponds to an ensemble of
+Coulomb particles. This discovery opens up promising avenues for
+further inquiry. Specifically, it would be interesting to delve deeper
+into this subject and explore whether the matrix model description
+offers valuable insights into the phase model and its Bethe
+roots. Such an investigation promises to illuminate the underlying
+dynamics of the phase model and its relationship with classical
+integrable systems.
+
+We have elucidated how various correlation functions align with this
+framework, revealing a remarkably rich structure. Notably, we explored
+a mapping between the phase model data and the vertex operator
+representation of free bosons. An intriguing avenue for future
+research is to delve into the properties of the hypergeometric tau
+functions uncovered in Section 3.3.1. Investigating these functions
+promises to yield valuable insights, and comparing them with existing
+literature will further enhance our understanding of their
+characteristics and implications within the context of integrable
+systems.
+
+In Section 4, we addressed the same problem within the context of the
+q-boson model. Initially, we derived a determinant formula for the
+q-boson, followed by a discussion on its expression as a quotient of
+two phase model tau functions. This result establishes a connection
+between the q-boson and phase model scalar product. Furthermore, we
+explored three different expansions for these scalar products. Notably, we
+observed that they can be expanded in terms of Big Schur and
+supersymmetric tau functions. Consequently, we introduced a new set of
+coordinates and demonstrated that it can be expressed as a Toda tau
+function in terms of this new coordinate set.
+
+One of our most pressing challenges lies in elucidating the intricate
+connection between our findings and the Ablowitz-Ladik
+hierarchy. Since this hierarchy arises as a reduction of the Toda
+hierarchy, it becomes imperative to understand how we can capture the
+structure of the former using the results we've derived. Given that
+the q-boson model can be seen as a quantization of the Ablowitz-Ladik
+equation, it stands to reason that we should observe some resemblance of
+this classical problem within the framework of the q-boson
+system. Understanding this relationship promises valuable insights
+into the interplay between classical and quantum integrable systems.
+
+We hope to address some of these challenges in future publications.
+
+\subsection*{Acknowledgments}
+This work is supported by Fapesp through grant \textbf{2022/06599-0}.
 

### 2024-04-30 draft.tex
improve draft

@@
 \address{\noindent 
-Instituto de Física Teórica, UNESP-Universidade Estadual Paulista,
-R. Dr. Bento T. Ferraz 271, Bl. II, Sao Paulo 01140-070, SP, Brazil\\
-\&\\
-Instituto de Física, Universidade de S\~ao Paulo,
-Rua do Matão Travessa 1371, 05508-090 São Paulo, SP. Brazil
+Instituto de Física Teórica, UNESP-Universidade Estadual Paulista, \\
+Rua Dr. Bento T. Ferraz 271, Bl. II, São Paulo 01140-070, SP, Brazil
+\\ \& \\
+Instituto de Física, Universidade de São Paulo,\\
+Rua do Matão, Travessa 1371, 05508-090 São Paulo, SP. Brazil
 }
@@
 \begin{abstract}
-This work delves into the complex interplay between the q-boson model,
-a quantum integrable system, and classical integrable systems like the
-Toda and KP hierarchies. By analyzing scalar products of Bethe states
-and correlation functions, we investigate their connections to tau
-functions and explore Schur polynomial expansions. This exploration
-unravels the intricate relationship between quantum and classical
-integrable systems.
+This work investigates the intricate relationship between the q-boson
+model, a quantum integrable system, and classical integrable systems
+such as the Toda and KP hierarchies. Initially, we analyze scalar
+products of off-shell Bethe states and explore their connections to
+tau functions of integrable hierarchies. Furthermore, we discuss
+correlation functions within this formalism, examining their
+representations in terms of tau functions, as well as their Schur
+polynomial expansions.
 \end{abstract}
 
-%\date{\today}
-%\keywords{Integrability, Bethe, Schur}
+\date{\today}
+\keywords{Integrability, Bethe states, Schur, Hall-Littlewood, Toda, KP}
 \subjclass[2020]{82B20, 82B23}
+
 \maketitle
 
-\setcounter{tocdepth}{2}
+\setcounter{tocdepth}{1}
 \tableofcontents
@@
 within correlation functions of the q-boson system. This quantum
-integrable system characterizes strongly correlated bosons confined to
+integrable system describes q-deformed bosons confined to
 a one-dimensional chain~\cite{Bogoliubov:1992, Bogoliubov:1997soj,
@@
 
-Foda and Wheeler~\cite{Foda:2008hn, Wheeler:2010vmq} utilize a rich
+The authors of~\cite{Foda:2008hn, Wheeler:2010vmq} utilize a rich
 set of dualities between the ring of symmetric functions and
@@
 This paper builds upon these developments and aims to extend some of
-these results. Firstly, we delve deeper into the connection between
+these results. Firstly, we investigate the connections between
 the scalar product and tau functions, exploring them in a wider
@@
 noteworthy that even when the object is not a tau function itself, it
-may have an interesting expansion in terms of different classes of
+may have an interesting expansion in terms of different types of
 Schur polynomials, highlighting some of its combinatorial
-properties\footnote{Numerous results involving Schur and
-Hall-Littlewood polynomials have been confirmed using the
+properties\footnote{Results involving Schur and Hall-Littlewood
+polynomials have been confirmed using the
 software~\cite{Araujo:2024piv}.}. We also investigate some of these
@@
 key aspects relevant to our analysis.  Section 3 delves into the
-analysis of the Phase Model and its connections with the KP and Toda
+analysis of the phase model and its connections with the KP and Toda
 integrable hierarchies. We explore the general expansion of a phase
@@
 literature, offering insights into established concepts.  While the
-content may not be entirely original, its presentation adds value,
-particularly from a pedagogical standpoint. Additionally, it provides
-a well-curated collection of references for further exploration.
-
-In structuring this discussion, we adopt certain conventions
-from~\cite{Wheeler:2010vmq}, while aligning our presentation style
-more closely with that of~\cite{Tsilevich:2006}.
+content does not introduce novel ideas, its presentation adds value,
+particularly from a pedagogical standpoint.  In structuring this
+discussion, we adopt certain conventions from~\cite{Wheeler:2010vmq},
+while aligning our presentation style more closely with that
+of~\cite{Tsilevich:2006}.
 
@@
 Consider the \((M+1)\) set of operators \(\{\phi_i,
-\phi^\dagger,\mathcal{N}_i\}_{i=0}^M\) such that
+\phi_i^\dagger,\mathcal{N}_i\}_{i=0}^M\) such that
 \begin{equation}
@@
 These operators appear in the context of quantum optics, and for this
-reason, it is referred to as the \emph{phase model}. It corresponds to
+reason, this model is referred to as the \emph{phase model}. It corresponds to
 the strongly correlated limit of the \(q\)-bosons
@@
 \subsubsection{Representation}
-The representation of the phase model algebra, is constructed using
-the vacuum state, defined by \(\ket{0}_i\) by \(\phi_i\ket{0}_i =
+The representation of the phase model algebra is constructed using
+the vacuum state defined by \(\ket{0}_i\) by \(\phi_i\ket{0}_i =
 0\). In this context, the state with \(n_i\) bosons (oscillators) is
-given by \(\ket{n_i}_i = \phi_i^\dagger \ket{0}_i\).
+given by \(\ket{n_i}_i = (\phi_i^\dagger)^{n_i} \ket{0}_i\).
 
@@
 \begin{equation}
-  \mathcal{F} = \bigotimes_{i=0}^M \mathcal{F}_i = 
+  \mathcal{F} = \bigotimes_{i=0}^M \mathcal{F}_i = \mathrm{span}
   \left\{\ket{\vec{n}} = \ket{n_0}\otimes \ket{n_1} \otimes \cdots
@@
 Given a state \(\ket{\vec{n}} = \ket{n_0, n_1, \dots, n_M}\), we
-associate a partition \( \lambda = (0^{n_0} 1^{n_1} 2^{n_2} \cdots
+associate a partition \( \lambda = (1^{n_1} 2^{n_2} \cdots
 M^{n_M})\).  It's worth noting that this correspondence is not unique,
@@
 represents the number of rows in the Young diagram defined by
-partition \(\lambda\). This determination is crucial, as in our
-subsequent considerations, the \(N\) particle sector remains fixed due
-to the integrability of the model. Consequently, the value of \(n_0\)
-becomes known once we specify the partition \(\lambda\).
-
-Finally, based on the correspondence we mentioned above,
+partition \(\lambda\).  This aspect is crucial because, in our
+subsequent considerations, the \(N\) particle sector remains fixed
+owing to the integrability of the model. Consequently, the value of
+\(n_0\) becomes known once we specify the partition \(\lambda\).
+
+Finally, based on the correspondence we mentioned in the introduction,
 Wheeler~\cite{Wheeler:2010vmq} defines a map \(\mathcal{M}_\psi:
@@
 \subsection{q-Bosons}
-The set of operators \(\{b_i, b^\dagger,\mathcal{N}_i\}_{i=0}^M\) constitute
+The set of operators \(\{b_i, b_i^\dagger,\mathcal{N}_i\}_{i=0}^M\) constitute
 \(M+1\) independent q-boson algebras defined by
@@
 \subsubsection{Representation}
-We construct the representation of the q-boson algebra with
-modifications of the phase model discussed earlier. Let us define the
-\(i\)-th vacuum \(\ket{0}_i\) such that \(b_i\ket{0}_i = 0\). The
-representation of this Hilbert space, denoted by \(\mathcal{F}_i^Q\),
-is given by \(\ket{n_i}_i \propto b_i^\dagger \ket{0}_i\).
+Let us define the \(i\)-th vacuum \(\ket{0}_i\) such that
+\(b_i\ket{0}_i = 0\). The representation of this Hilbert space,
+denoted by \(\mathcal{F}_i^Q\), is defined by the states
+\(\ket{n_i}_i \propto (b_i^\dagger)^{n_i} \ket{0}_i\).
 
@@
   \mathcal{F}_{Q} = \bigotimes_{i=0}^M \mathcal{F}_i^Q = 
-  \left\{\ket{\vec{n}} = \ket{n_0}\otimes \ket{n_1} \otimes \cdots
+  \mathrm{span}\left\{\ket{\vec{n}} = \ket{n_0}\otimes \ket{n_1} \otimes \cdots
   \otimes \ket{n_M} \ | \ n_i \in \mathbb{N} \right\}\; ,
@@
 This section explores the presence of integrable hierarchies in the
-phase model. We demonstrate its correspondence to a Toda hierarchy tau
+phase model. We demonstrate its relations with the Toda hierarchy tau
 function and discuss some implications, particularly its connection to
-a matrix model. Additionally, we highlight the existence of
-correlation functions in this model that satisfy the KP hierarchy
-equations. Finally, we argue that these results imply that these
-scalar products are also KP hierarchy tau functions, aligning with the
-findings of Wheeler~\cite{Wheeler:2010vmq}.
+a matrix model. We also argue that these results imply that the scalar
+products in the model also serve as KP hierarchy tau functions,
+consistent with the findings of
+Wheeler~\cite{Wheeler:2010vmq}. Additionally, we highlight the
+existence of correlation functions in this model that satisfy the KP
+hierarchy equations.
 
@@
 \end{equation}
-It is a tau function of the Toda hierarchy with trivial element \(\mathbb{1} \in GL(\infty)\).
-It is nothing but the Cauchy's identity
+It is a tau function of the Toda hierarchy with trivial element
+\(\mathbb{1} \in GL(\infty)\), and it is nothing but the Cauchy's
+identity
 \begin{equation}
@@
   Zabrodin:2010ii}, the truncation of the tau function corresponds to
-the inclusion of a projection operator \(\mathrm{P}^+\) in the
-expectation value of the tau function written in the fermionic
-representation.
+the inclusion of a projection operator in the expectation value of the
+tau function written in the fermionic representation.
 
@@
 \begin{equation}
-  H(z,w) = \sum_{k=0}^{M+N} h_k(zw)^{k-1} \; , 
+  H(z,w) = \sum_{k=0}^{M+N-1} h_k(zw)^{k} \; , 
 \end{equation}
@@
 \begin{equation}
-\mathcal{I}(N, M|\bm{t}, \bm{t}') = \sum_{\lambda} h_\lambda s_\lambda(\bm{t}) s_\lambda(\bm{t}') \; .
+\mathcal{I}(N, M|\bm{t}, \bm{t}') = \sum_{\lambda \subseteq [N,M]} h_\lambda s_\lambda(\bm{t}) s_\lambda(\bm{t}') \; .
 \end{equation}
@@
 particles on a circle. In this case, we find that the quantities
-\(\lambda_{\ell}\) are eigenvalues of a matrix \(U\).
+\(z_{\ell}\) are eigenvalues of a matrix \(U\).
 
@@
 \subsection{Correlation functions}
-Bogoliubov has also shown in~\cite{Bogoliubov2005}, that the
+Bogoliubov has also shown, in~\cite{Bogoliubov2005}, that the
 correlation functions
@@
 Furthermore, we define vector field \(\bm{F}(z) = (F_1, \dots, F_N)\),
-where the components are given by
+where its components are given by
 \begin{equation}
@@
 \end{equation}
-As before, we expand \(F_j(z)\) as the geometric sum
+As before, we expand \(F_j(z)\), \(j\neq N\), as the geometric sum
 \begin{equation}
   F_j(z) = \frac{1 - (y_j z)^{M + N}}{1 - y_j z} = \sum_{n=0}^{N + M - 1} (y_j z)^n
-  \equiv \sum_{n=0}^{M + N -1} f_{j, n} z^n\; .
-\end{equation}
+  \equiv \sum_{n=0}^{M + N -1} f_{j, n} z^n \; , \quad  f_{j, n} = y_j^n\; .
+\end{equation}
+We can also express the function \(F_N(z)\) in this form by setting
+\(f_{N,n} = \delta_{n,(M - N - 1 - 2m)/2}\) and requiring \((M - N -
+1)\) to be an even integer.
 
@@
 \begin{equation}
-  \det_{jk} F_j(x_k) = \det_{jk} \left(  \sum_{n=0}^{M + N -1} f_{j, n} x_k^n \right) \; .
-\end{equation}
-By utilizing the Cauchy-Binet formula, we can write
+  \det_{jk} F_j(x_k) = \det_{jk} \left(  \sum_{n=0}^{M + N -1} f_{j, n} x_k^n \right) \; ,
+\end{equation}
+and utilizing the Cauchy-Binet formula, we get
 \begin{equation}
@@
 
-We now employ the definition of the Schur polynomials and utilize the
+We now use the definition of the Schur polynomials and the
 Jacobi-Trudi expression of Plücker coordinates, as detailed
-in~\cite{Alexandrov:2012tr}. Ultimately, this leads to the following
-expression:
+in~\cite{Alexandrov:2012tr}, leading to the following expression:
 \begin{equation}
@@
 \end{equation}
-where \(c_\lambda \equiv \det_{jk}(y_j^{\ell_k})\),  \(\ell_k = \lambda_k - k +N\).
-Therefore, we have that this expression is a KP tau function.
+where \(c_\lambda \equiv \det_{jk}(y_j^{\ell_k})\), \(\ell_k =
+\lambda_k - k +N\).  Putting all these facts together, we conclude
+that this expression is also a KP tau function.
+
+This expression underscores the non-trivial nature of these tau
+functions within the model. However, it also implies the existence of
+other intriguing examples awaiting exploration. Let us now briefly
+investigate other cases.
 
@@
 norm~(\ref{eq:scalar}) times a finite sum over skew Schur
-functions. This expression illustrates that the existence of these tau
-functions is not a trivial aspect of this model.
+functions.
 
@@
 
-\subsection*{Acknowledgments}
+\subsubsection*{Acknowledgments}
 This work is supported by Fapesp through grant \textbf{2022/06599-0}.

### 2024-05-01 draft.tex
finish draft

@@
 \begin{equation}
-  t_p = \frac{1}{p}\sum_{j=1}^N x_j^p\qquad 
-  t'_{-p} = \frac{1}{p}\sum_{j=1}^N y_j^p\; .
-\end{equation}
-One can write the inner product in terms of these coordinates, that is
+  t_q = \frac{1}{q}\sum_{j=1}^N x_j^p\qquad 
+  t'_{-q} = \frac{1}{q}\sum_{j=1}^N y_j^p\; , 
+\end{equation}
+where \(p_q(\bm{x}) = q t_q\) are power sums. One can write the inner
+product in terms of these coordinates, that is
 \begin{equation}
@@
 \begin{equation}
- \mathcal{I}(N,M|\bm{t}) 
+ \lim_{M,N\to \infty}\mathcal{I}(N,M|\bm{t}) 
  = \sum_{\lambda} s_\lambda(\bm{y})  s_\lambda(\bm{t}) 
@@
 \subsubsection{Skew Schur polynomials expansion}
-From the coordinate expansion of the Bethe
-vectors~\cite{Bogoliubov2005, Tsilevich:2006}, it is easy to show that
-\begin{equation}
-\prod_{j=1}^N \mathbb{B}(x_j)\ket{\lambda}  = \sum_{\mu \supset \lambda } s_{\mu/\lambda}(\bm{x})\ket{\mu}\qquad 
- \bra{\lambda} \prod_{j=1}^N \mathbb{C}(x_j) = \sum_{\mu \supset \lambda } s_{\mu/\lambda}(\bm{x})\bra{\mu}
+It has also been demonstrated in~\cite{Bogoliubov2005, Tsilevich:2006}
+that the Bethe states exhibit a coordinate expansion
+\begin{equation}
+  \prod_{j=1}^N \mathbb{B}(x_j)\ket{\lambda}  =
+  \sum_{\substack{\mu \supset \lambda \\ \mu \subseteq [N,M]}} s_{\mu/\lambda}(\bm{x})\ket{\mu}\qquad 
+  \bra{\lambda} \prod_{j=1}^N \mathbb{C}(x_j) = \sum_{\substack{\mu \supset \lambda \\ \mu \subseteq [N,M]}} 
+  s_{\mu/\lambda}(\bm{x})\bra{\mu}\; ,
 \end{equation}
 where \(s_{\mu/\lambda}\) are skew Schur polynomials~\cite{Macdonald:1998}.
+
 Therefore, we can write the correlation function~(\ref{eq:correlation}) as 
@@
   A_m(N, M|\bm{x}, \bm{y}\setminus \{y_N\})
-  & = \sum_\lambda \sum_{\mu \supset (m)} s_{\lambda}(\bm{x}) s_{\mu/(m)}(\bm{y}\setminus\{y_N\})\\
-  & = \sum_\mu s_{\mu/(m)}(\bm{y}\setminus\{y_N\}) s_{\mu}(\bm{x})\; .
+  & = \sum_\lambda \sum_{\substack{\mu \supset (m)\\ \mu \subseteq [N-1, M]}}
+  s_{\lambda}(\bm{x}) s_{\mu/(m)}(\bm{y}\setminus\{y_N\})\\
+  & = \sum_{\substack{\mu \supset (m)\\ \mu \subseteq [N-1, M]}}
+  s_{\mu/(m)}(\bm{y}\setminus\{y_N\}) s_{\mu}(\bm{x})\; .
 \end{split}
@@
   \prod_{k=1}^{N'} \mathbb{B}(y_k)\ket{\lambda_2} \\
-  & = \sum_{\mu} s_{\mu/\lambda_1}(\bm{x}) s_{\mu/\lambda_2}(\bm{y}) \; .
+  & = \sum_{\mu \subseteq [\min(N, N'), M]} s_{\mu/\lambda_1}(\bm{x}) s_{\mu/\lambda_2}(\bm{y}) \; .
 \end{split}
@@
 we have also used that the skew Schur polynomial for any Young diagram
-\(\mu\) that does not contain \(\lambda_1\) and \(\lambda_2\)
-vanishes. Using the elementary property of skew Schur
-polynomials~\cite{Macdonald:1998}
+\(\mu\) that does not contain \(\lambda_1\) and/or \(\lambda_2\)
+vanishes.
+
+In the limit \(M, N, N' \to \infty\), we can use the elementary
+properties of skew Schur polynomials~\cite{Macdonald:1998}
 \begin{equation}
@@
 \end{equation}
-and the Cauchy's identity, we have 
+and the Cauchy's identity, we have that  
 \begin{equation}
 \begin{split}
-  A_{\lambda_1 \lambda_2}(N', N, M|\bm{x}, \bm{y}) & =
- \sum_\mu s_{\mu}(\bm{x}) s_{\mu}(\bm{y})
+  \lim_{M, N, N'\to \infty} A_{\lambda_1 \lambda_2}(N', N, M|\bm{x}, \bm{y})
+  & = \sum_\mu s_{\mu}(\bm{x}) s_{\mu}(\bm{y})
+ \sum_\nu s_{\lambda_1/\nu}(\bm{x}) s_{\lambda_2/\nu}(\bm{y})\\ 
+  & = \sum_\mu s_{\mu}(\bm{x}) s_{\mu}(\bm{y})
  \sum_\nu s_{\lambda_1/\nu}(\bm{x}) s_{\lambda_2/\nu}(\bm{y})\; .
@@
 \end{equation}
-Therefore, this correlation function is nothing but the off-shell
-norm~(\ref{eq:scalar}) times a finite sum over skew Schur
-functions.
+Therefore, we write the finite case as
+\begin{equation}
+  A_{\lambda_1 \lambda_2}(N', N, M|\bm{x}, \bm{y})
+  = \mathcal{I}(\min(N',N),M, \bm{x}, \bm{y})
+  \sum_\nu s_{\lambda_1/\nu}(\bm{x}) s_{\lambda_2/\nu}(\bm{y})\; .
+\end{equation}
+We observe that this correlation function can be expressed as the
+product of the off-shell norm~(\ref{eq:scalar}) and a finite sum over
+skew Schur functions.
 
@@
 functions, may have interesting Schur polynomial expansions. Consider
-the state calculated in~\cite{Bogoliubov2005}.
+the state calculated in~\cite{Bogoliubov2005}
 \begin{equation}
@@
 Based on the discussion we have had so far, and on the general mapping
-between the Phase model and free fermions, one can grasp the general
-form of tau functions in the context of this phase model.
+between the phase model and free fermions, one can grasp the general
+form of tau functions in the context of this integrable chain.
 
@@
 consider the vacuum state \(\ket{\bm{0}}\), often referred to as a
-``Dirac sea'', defined by the conditions \(\psi_m\ket{\bm{0}} =
+``Fermi sea'', defined by the conditions \(\psi_m\ket{\bm{0}} =
 \psi_n^\star\ket{\bm{0}} =0 \) for \(m<0\) and \(n \geq 0\), where
@@
 \end{equation}
-The group \(GL(\infty)\) is defined through the \(\exp:
-\mathfrak{gl}(\infty) \to GL(\infty)\) as usual.
-
-Note that these elements have finitely many non-zero entries,
-including diagonal terms, upper triangular terms, and lower triangular
-terms. These are equivalent to the number operators \(\mathcal{N}\),
-annihilation operators \(\phi\), and creation operators
-\(\phi^\dagger\), respectively. The central charge is equivalent to
-the vacuum projection \(\pi = \ket{\bm{0}}\bra{\bm{0}}\). Therefore,
+The group \(GL(\infty)\) is defined through the exponential map
+\(\bm{\exp}: \mathfrak{gl}(\infty) \to GL(\infty)\) as usual.
+
+Note that these elements have only finitely many non-zero entries: the
+diagonal terms represent number operators \(\mathcal{N}\), the upper
+triangular terms represent annihilation operators \(\phi\), and the
+lower triangular terms represent creation operators
+\(\phi^\dagger\). The central charge corresponds to the vacuum
+projection \(\pi = \ket{\bm{0}}\bra{\bm{0}}\).  Therefore,
 \begin{equation}
@@
    \mathcal{G} = \exp \left[\sum_{i=1}^M \left( \sum_{a=1}^3
-   c_{i, a} T_i^{a}  + c \pi_i \right)\right]\; ,
+   c_{i, a} T_i^{a}  + c \pi_i \right)\right] \; \in  \mathrm{Aut}(\mathcal{F})\; ,
 \end{equation}
@@
 \begin{equation}
-  \begin{split}
-    \mathbb{B}(x) \mapsto \Gamma_-(x) & = \exp \left( \sum_{n\geq 1} \frac{1}{n}x^n J_{-n}\right) \\
-    \mathbb{C}(x) \mapsto \Gamma_+(x) & = \exp \left( \sum_{n\geq 1} \frac{1}{n}x^n J_{n}\right) \; ,
-  \end{split}
+    \mathbb{B}(x) \mapsto \Gamma_-(x)  = \exp \left( \sum_{n\geq 1} \frac{1}{n}x^n J_{-n}\right) \qquad 
+    \mathbb{C}(x) \mapsto \Gamma_+(x)  = \exp \left( \sum_{n\geq 1} \frac{1}{n}x^n J_{n}\right) \; ,
 \end{equation}
@@
 \sum_{j\in \mathbb{Z}} \bm{\colon} \psi_j \psi_{j+n}^\ast
-\bm{\colon}\). These set of operators generate a Heisenberg subalgebra
+\bm{\colon}\). This set of operators generate a Heisenberg subalgebra
 \(\widehat{\mathfrak{gl}}(1) \subset \mathfrak{gl}(\infty)\)
@@
 \end{equation}
-where we use that \(\bracket{\lambda}{\mu} = b_\lambda(Q)
+where we use that \(\bracket{\lambda}{\mu}_Q = b_\lambda(Q)
 \delta_{\lambda, \mu}\) and the completeness relation
@@
 \end{equation}
-Here, the polynomials \(\mathcal{S}_\lambda\), known as the big-Schur
-functions, are defined by a Jacobi-Trudi formula:
+Here, the polynomials \(\mathcal{S}_\lambda\), which we will refer to as
+the big-Schur functions, are defined by a Jacobi-Trudi formula:
 \begin{equation}
@@
 \end{equation}
-and the coefficients \(q_m\) are obtained from the expression:
+where the coefficients \(q_m\) are obtained from the expression:
 \begin{equation}
  \sum_{m} q_m(\bm{y}, Q) z^m =
- \prod_{j} \frac{1-Q y_j z}{1 - y_j z}\; .
-\end{equation}
-where \(z\) is a formal variable. It has been
+ \prod_{j} \frac{1-Q y_j z}{1 - y_j z}\; ,
+\end{equation}
+and \(z\) is a formal variable. It has been
 argued~\cite{Foda:2008hn} in that if we interpret the big-Schur
@@
 expressed as a quotient of two Toda tau functions or as a KP tau
-function with coefficients given by the big Schur
-polynomials. However, further exploration of these results is
-warranted.
+function with coefficients given by the big Schur polynomials.
+Nevertheless, further investigation of these results is necessary.
 
@@
 This expression also reveals something interesting. We can formulate
-this problem in terms of partitions defined in the phase
+this problem in terms of partition statess defined in the phase
 model. Alternatively, we can utilize the conventional vertex operator
@@
 \subsection{Supersymmetric Schur polynomials expansion}
-With this result, we now argue that although~(\ref{eq:scalar-hl}) is
-not a Toda tau function with respect to the coordinates \(\{\bm{t};
-\bm{t}'\}\), we can define a new set of coordinates
-\(\{\bm{t};\bm{T}\}\) that render the inner product a trivial Toda tau
-function.
+Based on this result, we argue that although Equation
+\eqref{eq:scalar-hl} is not a Toda tau function with respect to the
+coordinates \( \{\bm{t}; \bm{t}'\} \), we can define a new set of
+coordinates \( \{\bm{t};\bm{T}\} \) such that the scalar product
+becomes a trivial Toda tau function.
 
@@
 \end{equation}
-where \(t^{(Q)}_n = \frac{Q^n}{n} \sum_j y_j^n = Q^n t'_n\). Additionally
+where \(t^{(Q)}_n = \frac{Q^n}{n} \sum_j y_j^n = Q^n t'_{-n}\). Additionally
 \begin{equation}
@@
     \prod_{j=1}^N \exp \left( \sum_{n>0} \frac{y_j}{n} z^n \right)=
-    \exp \left(\sum_{n>0} t_n z^n \right) \; . 
+    \exp \left(\sum_{n>0} t'_{-n} z^n \right) \; . 
 \end{equation}
@@
  \prod_j \frac{1 - Q y_j z}{1 - y_k z}
- = \exp \left( \sum_{n\geq 1} (t'_n - t_n^{(Q)})z^n \right)
+ = \exp \left( \sum_{n\geq 1} (t'_{-n} - t_n^{(Q)})z^n \right)
  \equiv \exp \left( \sum_{n\geq 1} T_n z^n \right)\; ,
 \end{equation}
-with \(T_n = (1 - Q^n) t'_n\). Then, we conclude that \(q_m(\bm{y}, Q)\)
+with \(T_n = (1 - Q^n) t'_{-n}\). Then, we conclude that \(q_m(\bm{y}, Q)\)
 are homogeneous polynomials with respect the Miwa coordinates
@@
 coordinates.  In this sense, we would have a curve in the infinite
-Grassmannian instead of a point.
+Grassmannian instead of a point. Evidently, it is not a tau functions
+on any known integrable hierarchy, but it might suggest some new
+generalizations that are worth investigating.
 \end{remark}
@@
 Combining these observations, we conclude that the left-hand side of
-equation (\ref{eq:tau-hl}) serves as a tau function with respect to
-the coordinates \(\{\bm{t}; \bm{T}\}\). Additionally, utilizing the
+equation \(\ref{eq:tau-hl}\) is a tau function with respect to
+the coordinates \(\{\bm{t}; \bm{T}\}\). Furthermore, by employing the
 vertex operator formalism, we can express the Schur polynomials as
@@
   s_{\lambda}(\bm{t}') = \bra{\lambda} e^{J_-(\bm{t}')}\ket{\bm{0}}\qquad
-  s_{\mu}(\bm{t}) = \bra{\bm{0}} e^{J_+(\bm{t})}\ket{\mu}\; ,
-\end{equation}
-we have
+  s_{\mu}(\bm{t}) = \bra{\bm{0}} e^{J_+(\bm{t})}\ket{\mu}\; . 
+\end{equation}
+Hence
 \begin{equation}
@@
 \end{equation}
-It is important to emphasize once again that this expression does not
-constitute a tau function with respect to \(\{\bm{t},
-\bm{t}'\}\). Despite this crucial distinction, we proceed under the
-assumption that the coefficients can be expressed as
+It is important to reiterate that this expression does not constitute
+a tau function with respect to \( \{\bm{t}, \bm{t}'\} \). Despite this
+crucial distinction, we proceed under the assumption that the
+coefficients can be expressed as
 \begin{equation}
@@
 \end{equation}
-where \(t'_n = \frac{1}{n}\sum_i y_i^n\).
+where \(t'_{-n} = \frac{1}{n}\sum_i y_i^n\).
 
 To derive the expression for this group element, recall that \(T_m =
-t'_m - Q^m t'_m\). With this in mind, we can express the scalar
+t'_{-m} - Q^m t'_{-m}\). With this in mind, we can express the scalar
 product as follows:
@@
 some early findings in this field and has also unveiled new research
-avenues, which we aim to explore in future studies.  Let us now delve
-into some of these promising directions.
-
-In Section 3, we delved into various facets of the phase model. We
+avenues, which we aim to explore in future studies. Let us now discuss
+some of these promising directions.
+
+In Section 3, we delved into some aspects of the phase model. We
 presented the scalar product of two off-shell Bethe states as an
@@
 Coulomb particles. This discovery opens up promising avenues for
-further inquiry. Specifically, it would be interesting to delve deeper
-into this subject and explore whether the matrix model description
-offers valuable insights into the phase model and its Bethe
-roots. Such an investigation promises to illuminate the underlying
-dynamics of the phase model and its relationship with classical
-integrable systems.
+further inquiry. Specifically, it would be intriguing to further
+explore this subject and investigate whether the matrix model
+description provides valuable insights into the phase model and its
+Bethe roots. Such an investigation promises to illuminate the
+underlying dynamics of the phase model and its relationship with
+classical integrable systems.
 
@@
 research is to delve into the properties of the hypergeometric tau
-functions uncovered in Section 3.3.1. Investigating these functions
-promises to yield valuable insights, and comparing them with existing
-literature will further enhance our understanding of their
+functions uncovered in Section 3.3.1.  Exploring these functions is
+expected to provide significant understanding, and comparing them with
+existing literature will further enrich our comprehension of their
 characteristics and implications within the context of integrable
@@
 q-boson model. Initially, we derived a determinant formula for the
-q-boson, followed by a discussion on its expression as a quotient of
-two phase model tau functions. This result establishes a connection
-between the q-boson and phase model scalar product. Furthermore, we
-explored three different expansions for these scalar products. Notably, we
+q-boson scalar products, followed by a discussion on their expression
+in terms of the phase model data. This result establishes a connection
+between the q-boson and phase model quantities. Furthermore, we
+explored different expansions for these scalar products. Notably, we
 observed that they can be expanded in terms of Big Schur and
-supersymmetric tau functions. Consequently, we introduced a new set of
-coordinates and demonstrated that it can be expressed as a Toda tau
-function in terms of this new coordinate set.
+supersymmetric tau functions. Hence, we introduced a new set of
+coordinates, demonstrating that scalar products can be precisely
+expressed as Toda tau functions within this transformed coordinate
+system.
 
@@
 hierarchy, it becomes imperative to understand how we can capture the
-structure of the former using the results we've derived. Given that
-the q-boson model can be seen as a quantization of the Ablowitz-Ladik
-equation, it stands to reason that we should observe some resemblance of
-this classical problem within the framework of the q-boson
-system. Understanding this relationship promises valuable insights
-into the interplay between classical and quantum integrable systems.
+structure of the AL hierarchy using the results we have derived.
+Considering that the q-boson model effectively quantizes the
+Ablowitz-Ladik equation, it follows that we should detect some
+resemblance of this classical problem within the q-boson system.  This
+understanding holds the promise of shedding significant light on the
+interplay between classical and quantum integrable systems.
 

### 2024-06-22 referee-answer.txt
answer to referee

Dear Editor,

Below are my responses to the referee's assessment:

Dear Referee,

I agree that my results are somewhat buried in the manuscript. Let me
start by saying that I was very careful not to claim credit for any
previous derivations. Some discussions in my current paper have been
addressed before, but not clearly enough. Therefore, this general
exposition also highlights my work and the field in
general. Additionally, there is pedagogical value in this.

Now, let me explain my main new results.

Section 3:

This section has more overlap with existing literature, particularly
with a paper by Wheeler that I cite in my work. In his paper, Wheeler
shows that the inner product of two Bethe states are KP hierarchy tau
functions. He finds the coefficients of the expansion and demonstrates
that these coefficients satisfy Plücker relations. My Section 3.1
improves on this result. I argue that besides the KP hierarchy, these
scalar products are also Toda hierarchy tau functions, as they can be
expanded as a Cauchy identity, with the coefficients being Schur
functions (see equation (28)). If we fix one set of coefficients in
this expression, Wheeler's results are recovered, as argued in
equation (40).

Section 3.2 is new and considers correlation functions. I wanted to
determine when the correlation functions satisfy integrable hierarchy
equations. Using expressions derived by Bogoliubov (see eq. (42.a)), I
rewrote them in the appropriate form to check if the correlation
functions are KP/Toda tau functions. For example, (42.b) cannot be
written as a Toda tau function, but with some adjustments, it can be
written as (51), which is a KP tau function. From the expansion (52),
also derived by Bogoliubov, we find (56) and (57), showing that these
expressions generally do not satisfy integrable hierarchy equations.

Section 3.3 presents my main results. Here, I argue that the general
form of tau functions associated with this phase model must be (67),
where G is given by (63). I also mention that from this general
expression, we can consider interesting cases, such as (68), which
fall into the class of hypergeometric tau functions. These problems
are currently under investigation in my follow-up paper.

Section 4:

This section attempts to replicate the analysis done in Section
3. There aren't many known results in the literature regarding the
relation between these objects and KP/Toda tau functions. I use known
formulas to see if they are tau functions. Wheeler has shown that
scalar products can be viewed as KP tau functions, as before.

I then write these formulas in terms of simpler orthogonal
polynomials. Firstly, I show that these scalar products cannot be
naively written as Toda tau functions (see (84)). However, using some
known expansions (see (85)), we can expand the inner products in
different forms. In one of them, we define a new set of coordinates
(equation (102)), making the scalar product of Q-Bosons Toda tau
functions with respect to these new coordinates. This is my main
result in this section. Additionally, I propose a conjecture worth
investigating in the future, as mentioned in Remark 2, regarding the
existence of 'functional' Plücker coordinates in the infinite
Grassmannian.

My Remark 3 also presents a new result. I emphasize that I do not wish
to take undue credit. Reference [NR19] uses an intricate construction
to show that the Big Schur functions are KP tau functions. Although
significant, my results trivialize their analysis. Since a simple
coordinate transformation makes the Big Schur functions ordinary Schur
functions (see equation (101)), this fact is trivial. I have included
this as a remark rather than stating it bluntly.

Having said that, let me return to my previous assessment of other
references. I am very interested in this field but have noticed that
papers on these topics are scattered. For example, in section 3.1.1,
the relation between the phase model and the ensemble of 2D Coulomb
particles seems obvious, but there are no references. I hope this
paper will draw attention to this particular problem. As I mentioned,
I am currently exploring this topic, but it is more challenging than I
previously thought.

Overall, I do not view this work as a full collection of new results
but as setting the stage for a research field that is somewhat
inactive. I use previous results to organize and draw attention to
open problems. In this process, I prove some new results, which,
though not major, serve as a starting point to investigate those open
problems.

I have added comments to my conclusions and improved my introduction.

I hope these clarifications are helpful.

Best regards,
Thiago

### 2024-06-23 referee-answer.txt
improve text

@@
 
-Below are my responses to the referee's assessment:
+Below is my response to the referee's assessment.
 
@@
 
-I then write these formulas in terms of simpler orthogonal
-polynomials. Firstly, I show that these scalar products cannot be
-naively written as Toda tau functions (see (84)). However, using some
-known expansions (see (85)), we can expand the inner products in
-different forms. In one of them, we define a new set of coordinates
-(equation (102)), making the scalar product of Q-Bosons Toda tau
-functions with respect to these new coordinates. This is my main
-result in this section. Additionally, I propose a conjecture worth
-investigating in the future, as mentioned in Remark 2, regarding the
-existence of 'functional' Plücker coordinates in the infinite
-Grassmannian.
+Observe that the determinant formula 83 is new. From this expression
+we can conclude that these objects cannot be written as 'simple' Toda
+tau functions.  I then write these formulas in terms of simpler
+orthogonal polynomials. Using some known expansions (see (85)), we can
+expand the inner products in different forms. In one of them, we
+define a new set of coordinates (equation (102)), making the scalar
+product of Q-Bosons Toda tau functions with respect to these new
+coordinates. This is my main result in this section. Additionally, I
+propose a conjecture worth investigating in the future, as mentioned
+in Remark 2, regarding the existence of 'functional' Plücker
+coordinates in the infinite Grassmannian.
 
@@
 
-I have added comments to my conclusions and improved my introduction.
+I have added these comments to my conclusions and I have also improved
+my introduction.
 

### 2024-06-23 draft.tex
improve text

@@
 
-The exploration of connections between quantum and classical
-exactly solvable models constitutes a significant endeavor aimed at
-elucidating the overarching structure of integrable systems.  This
-research avenue has yielded fruitful insights, as evidenced
+The study of connections between quantum and classical
+exactly solvable models is an important research program aimed at
+elucidating the underlying structure of integrable systems. This
+research program has yielded fruitful insights, as evidenced
 by~\cite{Its:1992bj, Foda:2009zz, Alexandrov:2011aa, Araujo:2021ghu},
@@
 within correlation functions of the q-boson system. This quantum
-integrable system describes q-deformed bosons confined to
-a one-dimensional chain~\cite{Bogoliubov:1992, Bogoliubov:1997soj,
-  Bogoliubov2005}. Notably, this model bears close relation to the AL
-(Ablowitz-Ladik) model, which is an integrable discretization of the
-nonlinear Schrödinger equation.
-
-In~\cite{Bogoliubov2005}, the author demonstrates that the \(q = 0\)
+integrable system describes q-deformed bosons confined to a
+one-dimensional chain~\cite{Bogoliubov:1992, Bogoliubov:1997soj,
+  Bogoliubov2005}.  Interestingly, this model is closely related to
+the AL (Ablowitz-Ladik) model, which is an integrable discretization
+of the nonlinear Schrödinger equation.
+
+In~\cite{Bogoliubov2005}, the author shows that the \(q = 0\)
 limit of the q-boson, known as the \emph{phase model}, is associated
@@
 This paper builds upon these developments and aims to extend some of
-these results. Firstly, we investigate the connections between
-the scalar product and tau functions, exploring them in a wider
-context, particularly in the case of the Toda
+these results. Firstly, we investigate the connections between the
+scalar product and tau functions, exploring them in a wider context,
+particularly in the case of the Toda
 hierarchy~\cite{Takasaki:2018wsv}. It is important to note that both
@@
 functions, as well as in terms of Kostka-Foulkes polynomials.
-Finally, in Section 5, we conclude with an overview of open problems.
+Finally, in Section 5, we conclude with a discussion of our results
+and an overview of open problems.
 
@@
 %*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*
-\section{Conclusions and Perspectives}
+\section{Discussion, Conclusions and Perspectives}
 %*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*
@@
 
-In Section 3, we delved into some aspects of the phase model. We
-presented the scalar product of two off-shell Bethe states as an
-elementary example of the Toda system tau function. Additionally, we
-revealed an intriguing alternative portrayal of these scalar products
-using a matrix integral framework, which corresponds to an ensemble of
-Coulomb particles. This discovery opens up promising avenues for
-further inquiry. Specifically, it would be intriguing to further
-explore this subject and investigate whether the matrix model
+In Section 3, we explored various aspects of the phase model,
+presenting the scalar product of two off-shell Bethe states as an
+elementary example of a Toda system tau function. This section
+overlaps with existing literature, particularly a paper by Wheeler and
+Foda~\cite{Wheeler:2010vmq, Foda:2008hn}. In their work, they
+demonstrate that the inner product of two Bethe states are KP
+hierarchy tau functions. Our current work builds on this by arguing
+that these scalar products are also Toda hierarchy tau functions, as
+they can be expanded as a Cauchy identity, with the coefficients being
+Schur functions. When one set of coefficients is fixed in this
+expression, Wheeler's results are recovered.
+
+We also extend this analysis to other correlation functions,
+determining when these functions satisfy integrable hierarchy
+equations. Using expressions derived by Bogoliubov and
+collaborators~\cite{Bogoliubov:1997soj}, we reformulated them to check
+if the correlation functions are KP/Toda tau functions. Although these
+objects cannot generally be written as Toda tau functions, with some
+adjustments, they can be expressed as KP tau functions.  We have
+elucidated how various correlation functions align with this
+framework, revealing a remarkably rich structure.
+
+The main results are discussed in Section 3.3, where we examine the
+general form of tau functions associated with this phase model.
+Notably, we explored a mapping between the phase model
+data and the vertex operator representation of free bosons. An
+intriguing avenue for future research is to delve into the properties
+of the hypergeometric tau functions uncovered in Section 3.3.1.
+Exploring these functions is expected to provide significant
+understanding, and comparing them with existing literature will
+further enrich our comprehension of their characteristics and
+implications within the context of integrable systems. This problem is
+currently under investigation.
+
+Additionally, we revealed an intriguing alternative portrayal of these
+scalar products using a matrix integral framework, which corresponds
+to an ensemble of Coulomb particles. This discovery opens up promising
+avenues for further inquiry. Specifically, it would be intriguing to
+further explore this subject and investigate whether the matrix model
 description provides valuable insights into the phase model and its
@@
 
-We have elucidated how various correlation functions align with this
-framework, revealing a remarkably rich structure. Notably, we explored
-a mapping between the phase model data and the vertex operator
-representation of free bosons. An intriguing avenue for future
-research is to delve into the properties of the hypergeometric tau
-functions uncovered in Section 3.3.1.  Exploring these functions is
-expected to provide significant understanding, and comparing them with
-existing literature will further enrich our comprehension of their
-characteristics and implications within the context of integrable
-systems.
-
 In Section 4, we addressed the same problem within the context of the
-q-boson model. Initially, we derived a determinant formula for the
-q-boson scalar products, followed by a discussion on their expression
-in terms of the phase model data. This result establishes a connection
-between the q-boson and phase model quantities. Furthermore, we
-explored different expansions for these scalar products. Notably, we
-observed that they can be expanded in terms of Big Schur and
-supersymmetric tau functions. Hence, we introduced a new set of
-coordinates, demonstrating that scalar products can be precisely
-expressed as Toda tau functions within this transformed coordinate
-system.
+q-boson model, attempting to replicate the analysis done in Section
+3. There are few known results in the literature regarding the
+relationship between these objects and KP/Toda tau functions, so we
+used established formulas to investigate if the scalar products are
+also tau functions.
+
+Initially, we derived a determinant formula for the q-boson scalar
+products and discussed their expression in terms of the phase model
+data. This result establishes a connection between the q-boson and
+phase model quantities. Furthermore, we explored different expansions
+for these scalar products. Notably, we observed that they can be
+expanded in terms of Big Schur and supersymmetric tau
+functions. Consequently, we introduced a new set of coordinates,
+demonstrating that scalar products can be precisely expressed as Toda
+tau functions within this transformed coordinate system.
+
+We also wrote these formulas in terms of simpler orthogonal
+polynomials. Firstly, we showed that these scalar products cannot be
+naively written as Toda tau functions. However, using some known
+expansions, we expanded the inner products in different forms. In one
+of these expansions, we defined a new set of coordinates, making the
+scalar product of q-bosons Toda tau functions with respect to these
+new coordinates. This is the main result of this section.
 

### 2024-06-24 draft.tex
finish new version

@@
 Define two sets of Miwa coordinates \(\bm{t} = (t_1, t_2, \dots)\) and
-\(\bm{t}' = (t'_{-1}, t'_{-2}, \dots)\) by
+\(\bm{t}' = (t'_{-1}, t'_{-2}, \dots)\) as
 \begin{equation}
@@
 This expression also reveals something interesting. We can formulate
-this problem in terms of partition statess defined in the phase
+this problem in terms of partition states defined in the phase
 model. Alternatively, we can utilize the conventional vertex operator
@@
 
-We also wrote these formulas in terms of simpler orthogonal
-polynomials. Firstly, we showed that these scalar products cannot be
+We also write these formulas in terms of simpler orthogonal
+polynomials. Firstly, we show that these scalar products cannot be
 naively written as Toda tau functions. However, using some known
-expansions, we expanded the inner products in different forms. In one
-of these expansions, we defined a new set of coordinates, making the
+expansions, we expand the inner products in different forms. In one
+of these expansions, we define a new set of coordinates, making the
 scalar product of q-bosons Toda tau functions with respect to these

### 2024-07-29 draft.tex
final version npb

@@
 to explore the combinatorial properties of this spin chain. As a
-consequence, it has been shown that the scalar product of off-shell
+consequence, it has been shown that the scalar products of off-shell
 Bethe states are restricted tau functions of the KP
@@
   H = - \frac{1}{2} \sum_{n =0}^M \left(\phi_n^\dagger \phi_{n+1}
-  + \phi_n \phi_{n+1}^\dagger \right) + \bar{\mathcal{N}}\; .
+  + \phi_n \phi_{n+1}^\dagger \right) + \bar{\mathcal{N}}\; ,
 \end{equation}
@@
 \(\lambda = (1^{n_1} 2^{n_2} \cdots M^{n_M})\). It is useful to define
-a proportionaly factor relating these two objects
+a proportionality factor relating these two objects
 \begin{equation}
@@
   \ket{\Psi(y_1, \dots, y_N; Q)} = \prod_{j=1}^N \mathbb{B}(y_j, Q) \ket{\bm{0}}\qquad 
-  \bra{\Psi(y_1, \dots, y_N; Q)} = \bra{\bm{0}}\prod_{j=1}^N \mathbb{C}(y_j, Q) \; .
+  \bra{\Psi(y_1, \dots, y_N; Q)} = \bra{\bm{0}}\prod_{j=1}^N \mathbb{C}(y_j, Q) \; ,
 \end{equation}
@@
 \frac{\det_{ik}(x_i^{\lambda_k - k + N})}{\Delta(\bm{x})} \frac{ \det_{ik}(y_j^{\lambda_k - k + N})}{\Delta(\bm{y})}
-= \sum_{\lambda\subseteq [N,M]} s_\lambda(\bm{x}) s_\lambda(\bm{y}) \; .
+= \sum_{\lambda\subseteq [N,M]} s_\lambda(\bm{x}) s_\lambda(\bm{y}) \; ,
 \end{equation}
@@
 \end{equation}
-that is a free energy, from the viewpoit of the matrix integral
+that is a free energy, from the viewpoint of the matrix integral
 partition function.
@@
   & = \sum_{\substack{\mu \supset (m)\\ \mu \subseteq [N-1, M]}}
-  s_{\mu/(m)}(\bm{y}\setminus\{y_N\}) s_{\mu}(\bm{x})\; .
+  s_{\mu/(m)}(\bm{y}\setminus\{y_N\}) s_{\mu}(\bm{x})\; ,
 \end{split}
@@
   \prod_{k=1}^{N'} \mathbb{B}(y_k)\ket{\lambda_2} \\
-  & = \sum_{\mu \subseteq [\min(N, N'), M]} s_{\mu/\lambda_1}(\bm{x}) s_{\mu/\lambda_2}(\bm{y}) \; .
+  & = \sum_{\mu \subseteq [\min(N, N'), M]} s_{\mu/\lambda_1}(\bm{x}) s_{\mu/\lambda_2}(\bm{y}) \; ,
 \end{split}
@@
 are related to the vertex operators
-\(\Gamma_-(x)\) and \(\Gamma_+(x)\), respectivelly, as
+\(\Gamma_-(x)\) and \(\Gamma_+(x)\), respectively, as
 \begin{equation}
@@
 \sum_{j\in \mathbb{Z}} \bm{\colon} \psi_j \psi_{j+n}^\ast
-\bm{\colon}\). This set of operators generate a Heisenberg subalgebra
+\bm{\colon}\). This set of operators generates a Heisenberg subalgebra
 \(\widehat{\mathfrak{gl}}(1) \subset \mathfrak{gl}(\infty)\)
@@
   e^{\sum_i c_i \mathcal{N}_i} \prod_j \mathbb{B}(y_j) \ket{\bm{0}}\\
-  & = \sum_{\mu, \nu \subseteq [N,M]} c_{\mu\nu} s_\mu(\bm{x}) s_\nu(\bm{x})
+  & = \sum_{\mu, \nu \subseteq [N,M]} c_{\mu\nu} s_\mu(\bm{x}) s_\nu(\bm{y})
 \end{split}
@@
   \tau(\bm{x}, \bm{y}) 
-  = \sum_{\mu, \nu \subseteq [N,M]} c_{\mu} s_\mu(\bm{x}) s_\mu(\bm{x})\; ,
+  = \sum_{\mu, \nu \subseteq [N,M]} c_{\mu} s_\mu(\bm{x}) s_\mu(\bm{y})\; .
 \end{equation}
@@
 \sum_{\lambda} b_\lambda(Q) P_{\lambda}(\bm{x}, Q) P_{\lambda}(\bm{y}, Q)
-= \prod_{j, k=1}^\infty \frac{1-Q x_j y_k}{1 - x_j y_k}\; ,
+= \prod_{j, k=1}^\infty \frac{1-Q x_j y_k}{1 - x_j y_k}\; .
 \end{equation}
@@
   \prod_{i,j}\frac{1}{1 - x_i y_j}  \equiv \mathcal{I}(N,M|\bm{x}, \bm{y}) = 
-  \frac{\det_{j,k}H(\bm{x},\bm{y})}{\Delta(\bm{x}) \Delta(\bm{y})}\; .
+  \frac{\det_{j,k}H(\bm{x},\bm{y})}{\Delta(\bm{x}) \Delta(\bm{y})}\; ,
 \end{equation}
@@
 big Schur polynomials are also KP tau functions. This direct
-conclusion serves as an alternative proof of this fact, originally derived in
+conclusion serves as an alternative proof of this fact, originally derived
 in~\emph{\cite{Necoechea:2019wbg}} using the KP bilinear identity.
@@
   \lim_{N,M\to\infty} \mathcal{I}_Q(N, M| \bm{t}, \bm{t}'(\bm{y})) = 
-  \bra{\bm{0}} e^{J_+(\bm{t})} G_Q(\bm{y}) e^{J_-(\bm{t}')}\ket{\bm{0}}\; .
+  \bra{\bm{0}} e^{J_+(\bm{t})} G_Q(\bm{y}) e^{J_-(\bm{t}')}\ket{\bm{0}}\; ,
 \end{equation}
@@
 Foda~\cite{Wheeler:2010vmq, Foda:2008hn}. In their work, they
-demonstrate that the inner product of two Bethe states are KP
+demonstrate that the inner products of two Bethe states are KP
 hierarchy tau functions. Our current work builds on this by arguing
